### 2016-01-16 paper/paperSource.tex
some paper sources drafts

\documentclass[10pt,journal,final,letterpaper,twocolumn]{IEEEtran}
%\documentclass{ssdbm}
\usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
\usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
\usepackage{epstopdf}
%\usepackage{fixltx2e}

\restylefloat{table}

% This was from the NEW paper
%----------------------------------------------------------
\RequirePackage{fix-cm}

\DeclareMathOperator{\var}{var}
\newcounter{qcounter}

%\smartqed  % flush right qed marks, e.g. at end of proof

%----------------------------------------------------------

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\newcommand{\SEC}[1] {Section \ref{#1}}
\newcommand{\Eq}[1] {Equation \ref{#1}}
\newcommand{\eq}[1] {Eq. (\ref{#1})}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

% paper title
\title{Parallelization of Push-based System for Molecular Simulation Data Analysis}

% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs

\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$% <-this % stops a space

%\thanks{* These authors contributed equally to this work.}
\thanks{$\dagger$ Author to whom all correspondence should be sent.}
\thanks{Vladimir Grupcev and Yi-Cheng Tu are with the Department of Computer
Science and Engineering, University of South Florida, 4202 E. Fowler
Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
vgrupcev@mail.usf.edu, ytu@cse.usf.edu}
\thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
University of South Florida, 4202 E. Fowler Ave., ISA 2019, Tampa,
FL 33620, U.S.A. Emails: jcfogart@mail.usf.edu, pandit@usf.edu} }

%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}

% use only for invited papers
%\specialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}

Thanks to the advancement of the modern computer simulation systems
and state-of-the-art instrumentation, many scientific fields
generate, and require manipulation of large amounts of data. Known
scientific data analysis systems, as well as traditional DBMSs,
follow a pull-based type of architectural design, where the executed
queries mandate the data needed on their part. This design, while
suitable for traditional transaction-based workloads where number of
queries retrieve small parts of data located at various places of
the database, is ill-fitted for applications involving complex
analysis on most of the data. Such design involves random I/O,
considerably affecting the data throughput in the system. All of
this gives rise of redundant and high I/O traffic in the process of
scientific data analysis in pull-based type systems. In this paper,
we design and implement a push-based type system which allows
high-throughput data analysis in the process of scientific
discovery. Our design improves on the aforementioned shortcomings in
two ways: i) it uses a sequential scan-based I/O framework that
loads the data into the main memory, and then ii) the system pushes
the loaded data to a number of pre-programmed queries. These queries
are being executed on the same data stream, with various data
selections. This way, essentially, the system lowers the unnecessary
I/O overhead imposed by the randomized, index-based scan and that of
a multiple data reads if each query were to be fed separately.
Considering the amount of MS data as well as the number of queries
being executed, we believe our system provides substantial
improvement over the current MS data analyzing systems. The
efficiency of the proposed system is backed by the results of
extensive experiments using MS generated data. The running times of
our system are compared to those of the GROMACS system. The
comparison shows the advantage and the potential of using such
push-based system for MS system analysis.

\end{abstract}

\begin{IEEEkeywords}
Push-based system, molecular simulation, scientific databases,
spatial distance histogram.
\end{IEEEkeywords}
%\vspace{-5mm}

\begin{comment}
Many scientific applications share several features that bring in
light the shortcomings of the current data analysis systems. First,
the scientific analysis often involves executing a number of
analytical queries commonly processed on a large portion, if not all
of the generated data. Second, many of these fields use the same
popular, and small number of data analytics primitives as the basic
blocks on which they build their discovery. And three, the
scientific data, once stored is never modified. New data only
appends to the existing dataset, making the data perfect contender
for streaming.

And while the use of an index-based scan might seem better option to
a sequential scan, this is the case only when the query accesses
small portion of the data.
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}\label{sc:intro}
The amount of data generated by scientific applications running on
advanced computer simulation systems as well as the new improved
experimental devices is growing rapidly. This big volume of data is
imposing a significant stress on the data analysis software. Even
though the existing data analysis systems are configured to deal
with large volume of data, they are not optimized for high
throughput data analysis. Often times, the data is available only
for certain period of time (e.g., streams) so the analyzing system
should react to it in a very efficient way. On top of the big amount
of data, another challenge that is being imposed to the data
analysis systems is that the data is usually being accessed through
high-level analytical queries which computation is a lot more
complex than the computation of simple aggregates. Some of these
queries are usually the bottleneck of the data analysis systems,
because they take a lot of time (often many days) to be executed and
the systems are not designed to handle multiple queries on the same
data stream at the same time, thus decreasing the overall efficiency
of the data analysis.

One such example, where fast data analysis is crucial involves
online social
media~\cite{nature_bigdata08,nature_bigdata12,science_social10}. The
social media is known to be used for real-time feedback collection,
which is usually associated to a particular topic or
product~\cite{jcs_twitter11}. In order for a system to be able to
perform analytical examination of the data produced in such
streaming media, the system should have the capability of fast data
access. The reason, the millions of data records(tweets) produced
every second~\cite{tkde_big14}. Moreover, these tweets may have
different geographical origin, introducing different languages and
forms and often times containing unsolicited messages, errors,
malicious content, etc. Therefore, some low level data uniformity
and cleaning on top of the data access and management issues should
be considered and possibly incorporated in the process of analytical
investigation in order to achieve relevant result.

The aforementioned big data imposed issues are also frequently
present in many other application fields, e.g., scientific data
analysis. In many scientific and engineering fields, the main
computational method for analyzing physical/chemical attributes of
natural systems is the particle simulation. Such simulations, when
performed in the field of structural and molecular biology, are
commonly called Molecular Simulations. Molecular simulations (MS)
(sometimes called Molecular Dynamics (MD) simulations) are computer
simulations of complex biological, physical or chemical structures.
These simulations, based on some basic theoretical models are widely
utilized as a basic research mechanism for studying the behavior of
the natural systems. The basic units of such systems/simulations are
natural particles (such as atoms, molecules, stars, etc.) and they
interact among each other for a certain period of time following
postulated classical forces. The number of these basic units taking
part in the simulations is big, usually in the range of hundreds of
thousands to millions. For instance, Fig.~\ref{fg:snap-white} shows
a snapshot of a collagen fiber, consisting of $890,000$ particles.
And oftentimes, the simulations produce datasets consisting of more
than one snapshot (frame) of the system's current state at various
time instants. Each of these frames contain all the particles,
together with all their measurements, such as spatial coordinates,
mass, charge, velocity, forces, etc. And usually, a big number of
such frames (in the tens of thousands) are being produced and stored
in the course of a typical simulation. Quantities measured during
the simulations are analyzed to test the theoretical
model~\cite{Frenkel:api01,Landau:cup05}. In short, the MS is proven
and powerful tool for understanding the inner-workings of a
biological system, by supplying a model description of the
biophysical and biochemical processes that are being unfold at a
nanoscopic scale.

In order to achieve objective discovery and to explain the workings
of the scientific systems, scientist must analyze the data produced
by the MS. These analysis oftentimes comprise of computation of very
complex quantities that show statistical properties of the data.
Such queries are of great importance to scientist because they are
the basic assembly blocks for a series of critical quantities needed
to outline the scientific systems \cite{Frenkel:api01}. The methods
by which the data is being accessed as well as these queries are
executed can either increase or decrease the efficiency of an data
analysis system. The known data analysis systems (e.g., Gromacs
\cite{Gromacs-online}) uses pull based type design in which the data
is being fed to the queries only by demand. When a query needs the
data, it requests it, the data is pulled through the system and
delivered to the query. This type of design, we believe, introduces
two types of problematic issues: 1. overhead in I/O traffic by not
allowing for multiple queries to be executed on the same stream of
data at the same time, and 2. cpu/data latency incurred when a
request for the data is sent to the system (the time it takes from a
request to the time the query can actually use the data).

We believe that by using a push-based type system we can remedy such
issues. In the push-based design the queries do not request the
data. Rather, the data is pushed onto the queries automatically and
is being processed by the active queries. This alleviates the I/O
traffic overhead as well as the cpu/data latency incurred for each
query that would have been introduced by a pull-based design. Such a
system design is especially suitable for many scientific
applications which share the following features. First, the
scientific analysis often involves executing a number of analytical
queries commonly processed on a large portion, if not all of the
generated data. Second, many of these fields use the same popular,
and small number of data analytics primitives as the basic blocks on
which they build their discovery. And three, the scientific data,
once stored is never modified. New data only appends to the existing
dataset, making the data perfect contender for streaming.

In this paper, we describe the idea, design and implementation of
such a push-based system. We would like to point out that the work
presented in this paper is an application of such push-based
strategy that has been previously considered in the database
community in the works presented in~\cite{DataPath,Volcano,Qpipe},
as well as the idea of sharing same tuple of data or data scans
across many queries previously introduced
in~\cite{Candea,PredictablePerformance,CooperativeScans}.

To summarize, data-intensive applications often require compelling
amount of storage space and intensive processing capability, but
also need fast data access and high throughput data analysis.
Therefore, the need of a system that will be optimized to access
data fast, with high throughput, as well as efficiently execute the
analytical queries with a possibility of running multiple queries on
the same data stream is of a great importance to the scientific
community.

\begin{figure}
 \centerline{ \includegraphics[width=1\columnwidth]{images/snap-white.eps} }
 \caption{A snapshot of simulated collagen fiber structure}
 \label{fg:snap-white}
\end{figure}

\subsection{Problem Statement}
As mentioned earlier, in order to describe the scientific system,
scientist must look into/analyze the critical statistical properties
of the data produced by the simulation. These properties are usually
computed through series of queries that are being executed on the
whole, or some selection of the MS data. Most of the systems that
are widely used today across many scientific fields for analysis of
the MS data are part of a software systems that also run the
simulation. Such examples include GROMACS~\cite{GROMACS4},
VMD~\cite{VMD}, MDAnalysis~\cite{MDAnalysis}, Wordom~\cite{wordom},
MD-TRACKS~\cite{MDtracks}, SimulaidOne~\cite{Simulaid},
Charmm~\cite{CHARMM}. Once the simulation is run, the system
produces a flat file with all the measurements of each particle.
These systems take each query as a user's input and apply them to
the data uploaded from these flat files (called trajectories in
GROMACS). The biggest problem/performance issue that these system
impose is the fact that for every new query issued by the user, the
system has to load a significant part of the dataset into main
memory before executing the query. Such pull-based design involves
random I/Os that considerably affecting the data throughput in the
system. And while the use of an index-based scan might seem better
option to a sequential scan, this is the case only when the query
accesses small portion of the data. All of this gives rise of
redundant and high I/O traffic in the process of scientific data
analysis in pull-based type systems. This adds greatly to the
overall performance time of the analysis of the MS system. Having in
mind the volume of data produced by the MS in a single frame (in
gigabytes), and we add the number of frames simulated (tens of
thousands) the total volume of data to be analyzed is of huge
magnitude. So to have to load big chunks of this data from disk to
the main memory every time a query is being executed degrades the
overall system's performance greatly. In this paper we design and
implement a push-based type system which allows high-throughput data
analysis in the process of scientific discovery.

\subsection{Our Approach}
Considering the fact that the running time of the most queries used
in the scientific world to analyze the MS is minuscule in comparison
to the loading time of the MS data, it is safe to say that the
scientific world would benefit from a system that can remedy the
aforementioned problem. Our idea is to build a system that would
load the MS data from hard disk in the main memory only once and
then execute as many queries as needed/wanted on that data, without
the need of reloading the same data over and over again. Considering
the volume of data in a single MS frame as well as the number of
frames produced during a single simulation, this type of system can
save a lot of time for MS analysis. The system would contain many,
most often used queries that can also be run as user's input. These
queries would be pre-programmed as separate modules in our system,
but they would be able to take certain attributes, like data
selection for instance, as user's input. Once the system is run, all
(or a selection) of these queries would be executed against the MS
data or certain selection. The system would act as a type of
push-based system, essentially pushing the loaded data to all the
queries. Once we have the data in the memory, the reading can loop
as many times as the queries need it. This is useful for some
non-streaming functions/queries that cannot be computed in a single
run of the data. Such one query is the SDH. But we have also come up
with a way to resolve this in a way, by creating a quad-tree based
database for this query. This quad-tree (we called Density Map (DM))
is being populated while the data is being loaded into main memory.
After the DM is populated, we can compute the SDH with much faster
algorithm than the brute-force method used everywhere else.

Another improvement that our system provides is the following: by
thorough observation and analysis we have discovered that many of
the quantities (functions/queries) that are being computed in order
to analyze the MS system share some of the same basic parts. In
order to take advantage of that, our systems precomputes these basic
parts and has them handy (stored in memory) whenever one of the more
complex functions needs them. So, for instance, when the total mass
of the system is needed, our system just pulls the precomputed value
out and serves the query that needs it immediately. And if more than
one query uses the same basic part, that translates into time
saving, however minuscule it might be.

\subsection{Contribution and roadmap of the paper} We have designed
and implemented a push-based system that can be used as a tool for
analysis of MS systems. We have tested the system on real MS
datasets. The experimental results evidently show the superiority of
the proposed system over one of the most widely used MS analysis
system across many scientific fields, i.e., GROMACS. The efficiency
improvement ranges anywhere from $3$ to $100$ times improvement (of
the overall running time) per set of selected queries run on a
different selection of the MS data. If we take in account the number
of such queries that the scientist run every day, we believe our
system would be of great benefit to MS community. In addition to the
higher efficiency that our system achieves, we also believe that our
idea will initiate a new breed of database related, push-based
systems that could be used to analyze the MS systems in a more
efficient way. Having a push-based like system, we were also able to
improve the computation of the more complex quantities. Namely, the
computation of some such complex quantities involves computation of
much less complex functions (sub-parts), and many such sub-parts are
being shared among some of the complex quantities. So, by computing
these more basic, sub-quantities in advance, we were able to achieve
however minuscule time gain when computing the more complex queries.

The major technical contributions presented here are:
\begin{itemize}
\item Design the network (tree like) of the most commonly used queries in MS (physics);
\item Build the system: design and build the modules, representing the quantities to be
computed in an efficient manner, following the already built
network;
\item Develop a scientific simulation database benchmark that can be used for evaluating similar systems and products.
\end{itemize}

The remainder of this paper is arranged as follows: in Section
\ref{sc:relatedwork} we give a survey of the systems used in the
field of MS systems analysis. We continue the paper with Section
\ref{sc:querynetwork} in which we show the design of the query
network build from the most widely used quantities in MS system
analysis. In Section \ref{sc:system} we describe our push-based
system for MS data analysis. Then, in Section \ref{sc:experiments}
we present the benchmark designed to test our system as well as the
results attained through comprehensive experiments run on real MS
generated data. At the end, we conclude this paper with Section
\ref{sc:conclusion} in which we give an overview of our possible
future endeavors in the field of MS data analysis.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}\label{sc:relatedwork}
The idea of data streaming has been broadly used in many fields. The
main usage however is aimed at processing live data generated
online. There is an ocean of references for data stream management,
but we believe the presentation in~\cite{DataStreamManagement}
encapsulates the majority of the ideas, problems and solutions. In
the past decade, however, the database community started to follow
the data stream idea to process stored data. Processes can take
advantage of the streaming data at any time the data is being pushed
through the system. This gives rise to push-based type design for
data management systems. The idea of such push-based design was
previously considered in projects such as DataPath~\cite{DataPath},
Volcano~\cite{Volcano} and QPipe~\cite{Qpipe} among others. These
works show ideas in which the data-driven dataflow is compared to
the demand-driven dataflow showing the need for the later. They also
talk about maximizing the data and work sharing among queries at
runtime. Essentially, we incorporate such ideas in our design of the
push-based system.

On the other side, the scientific community has steadily progressed
from processing massive data files towards employing database
systems for the storage, acquisition, and analysis of large-scale
scientific data~\cite{SDSS_SIGMOD02,QBISM_ICDE94}. The widely used
and popular relational database systems are conventionally designed
and optimized to better manage the data produced by the business
type applications. But such conventional database systems (DBMS) are
not well equipped to deal with the type and quantity of scientific
data, such as the data produced by the molecular simulations. In the
recent past, the DBMS community has made some attempts into the
design and construction of database systems optimized for handling
scientific data. Such examples include the BDBMS
project~\cite{BDBMS_CIDR07} that deals with annotation and
provenance of the sequence data in biosciences. and the PeriScope
project~\cite{PeriScope} is designed to efficiently handle
declarative queries against bio sequences. On top of the
aforementioned examples, there are also ideas for new DBMS
frameworks aimed at the management of scientific
data~\cite{NewArch-VLDB07,ROI-CIDR07,SciDB-SIGMOD10}. One of those
systems is the SciDB~\cite{SciDB-SIGMOD10,SciDB} and it is closest
to the idea presented in this paper. SciDB is data management and
analytics system that is primarily used is in application domains
involving very big scale array data. This system, like the one
presented in this paper is designed around a multi-dimensional array
data-model and it uses arrays to store the data. SciDB stores
petabytes of data on a number of machines and runs its queries on
those machines. It is made for high performance, high-availability,
fault tolerance, and scalability. However, to the best of our
knowledge, it too follows the pull based design where its queries
demand the data they need. As seen earlier in this paper, this type
of design can impose I/O overhead and decrease the data throughput
when doing the analysis. Aside the mentioned issue, the design and
build of such DBMS optimized for scientific data management come
with a additional challenges. Such challenges as well as their
probable resolution are outlined in~\cite{Sci_SIGMODR05}. Recently,
there have been some efforts aimed at designing and building MS data
managements systems on top of relational databases. Such efforts are
presented through projects like BioSimGrid~\cite{BioSimGrid} and
SimDB~\cite{SimDB} that were developed especially for molecular
simulations. However, to the best of our knowledge, such systems
still lack the efficiency needed for MS data management as well as
efficient query processing strategies.

Generally, the data produced in the process of molecular simulation
is being stored in large, plain files with no structure whatsoever.
Queries, which are implemented in a stand alone programs within
simulation/analysis systems, are executed onto such files producing
the quantities that scientist use to analyze the molecular system.
Such simulation and/or analysis packages include: Gromacs, VMD,
MDAnalysis, Wordom, MD-TRACKS, SimulaidOne, Charmm among the others.
But to the best of our knowledge, all of these systems work on a
similar basis: they take a user defined query and execute it onto
the MS simulated data. In order for the query to be executed, the
data has to be loaded into the main memory. Then the result is
either produced onto the display or written to a file. When the next
user query comes, the system again loads significant part of the
dataset into the main memory and executes the query. We believe that
there is a room for improvement of such systems, given the fact that
many of the user defined queries executed during system's analysis
are fairly static. In other words, there is a number of queries that
a user would always want to execute on a given simulation data.
Furthermore, the selections of MS data onto which such queries might
be executed, are also fairly constant (i.e., oftentimes the user
selects the same group of atoms (e.g., all hydrogen atoms) to
calculate given quantity, like center of mass for instance). So, by
precoding many such queries and running them automatically once the
system has loaded the data into memory, we believe we can save a lot
of time that otherwise would have been spent in loading the same
data into main memory anytime a query is executed. On top of the
automated query execution, our system can take user's query as input
as well. With this, we believe our system is an improvement over the
MS analysis systems that are used today.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Network of Queries}\label{sc:querynetwork}

\textbf{\emph{MS Queries}}. In order to study some important
statistical features of an MS system, scientist need to ``extract''
various statistical quantities out of the data produced by the
simulation. To achieve this, queries are executed against the data.
Most of the queries used in the analysis of MS systems are
analytical in nature. Essentially, these analytical queries are
mathematical functions that translate a selection of atoms (atoms'
measurements) to a scalar, vector, a matrix, or a data
cube~\cite{SimDB}. Once the simulation is done, the analysis carried
out will depend on the structure being studied as well as the
features of the system that need exploring. In other words, not all
system's quantities need to be computed every time the system is
being analyzed. Some of the more popular queries, including density
(atom counts), first-order statistics (mean), second-order
statistics (variance), and histograms among others, can be seen in
Table~\ref{tb:queries}. The queries shown in this table are the ones
that we have also incorporated in our system. Just to clarify some
of the notation in Table~\ref{tb:queries}: we assume that the MS
system comprises of $n$ particles and $r_i$, $m_i$, $c_i$ and $q_i$
denote coordinates (vector form), mass, charge, and number of
electrons of a particle $i$, respectively.

\begin{table}[h]
\resizebox{\columnwidth}{!}{\begin{minipage}{\columnwidth}
\renewcommand*{\arraystretch}{1.5}
\tabcolsep=0.12cm
\begin{tabular}{|c| c|}
\hline %inserts horizontal line
Function Name & Equation/Description \\[0.5ex] \hline
Moment of Inertia & $\begin{array} {lcl} I & = & \sum\limits_{i=1}^n m_ir_i \end{array}$ \\[0.5ex]
\hline
Moment of Inertia on z axis & $\begin{array} {lcl} I_z & = & \sum\limits_{i=1}^n m_ir_{zi} \end{array}$ \\[0.5ex]
\hline
Sum of masses & $\begin{array} {lcl} M & = & \sum\limits_{i=1}^n m_i \end{array}$ \\[0.5ex]
\hline
Center of mass & $\begin{array} {lcl} CoM & = & \frac{I}{M} \end{array}$ \\[0.5ex]
\hline
Radius of Gyration & $\begin{array} {lcl} RG & = & \sqrt{\frac{I_z}{M}} \end{array}$ \\[0.5ex]
\hline
Dipole Moment& $\begin{array} {lcl} D & = & \sum\limits_{i=1}^n q_ir_i \end{array}$ \\[0.5ex]
\hline
Dipole Histogram & $\begin{array} {lcl} D_z & = & \sum\limits_{i=1}^n \frac{D}{z} \end{array}$ \\[0.5ex]
\hline
Electron Density & $\begin{array} {lcl} ED & = & \frac{\sum\limits_{i=1}^n (e_i-q_i)}{dz \cdot x \cdot y} \end{array}$ \\[0.5ex]
\hline
Heat Capacity & $\begin{array} {lcl} HC & = & \frac{3000 \cdot \sqrt{T} \cdot boltz}{2 \cdot \sqrt{T}-n \cdot df \cdot VarT} \end{array}$ \\[0.5ex]
\hline
Isothermal Compressibility & $\begin{array} {lcl} I & = & \frac{VarV}{V_{avg} \cdot boltz \cdot T \cdot PresFac} \end{array}$ \\[0.5ex]
\hline
Mean Square Displacement& $\begin{array} {lcl} msd & = & \langle (r_{t+\Delta_t}-r_t)^2 \rangle \end{array}$ \\[0.5ex]
\hline
Diffusion Constant & $\begin{array} {lcl} D_t & = & \frac{6 \cdot msd(t)}{t} \end{array}$ \\[0.5ex]
\hline Velocity Autocorrelation & $\begin{array} {lcl} V_{acor} & =&
\langle (V_{t+\Delta_t}\cdot V_t) \rangle \end{array}$ \\[0.5ex]
\hline
Force Autocorrelation & $\begin{array} {lcl} F_{acor} & = & \langle (F_{t+\Delta_t} \cdot F_t) \rangle \end{array}$ \\[0.5ex]
\hline
Density Function & Histogram of atom counts \\[0.5ex]
\hline
SDH & Histogram of all distances \\[0.5ex]
\hline
RDF & $\begin{array} {lcl} rdf(r) & = & \frac{SDH(r)}{4 \cdot \pi \cdot r^2 \cdot \sigma_r \cdot \rho} \end{array}$ \\[0.5ex]
\hline
\end{tabular}
\caption[Table caption text]{Popular analytical queries in MS }
\label{tb:queries}
\end{minipage} }
\end{table}

There are two types of queries/functions among the ones used to
analyze an MS system. The first type are one-body functions. Such
functions usually are algebraic functions~\cite{SDSS_SIGMOD02} and
only involve quantities(attributes) from a single atom at any given
time in the process of computation. Each atom (atom's attributes) is
being processed a constant number of times, thus the total running
time of such functions/queries is $O(n)$. This type of functions is
very suitable for the idea of push-based system, or an online system
in which the data is being read once and acted upon. In other words,
in a single run of the incoming data, all such queries will produce
useful final results. Except the SDH and the RDF (i.e., the last two
in the table), all other queries shown in Table~\ref{tb:queries}
fall into this category of one-body functions. Most of these
functions are defined on a single frame of the MS data. Only the
autocorrelation functions are defined on two distinct frames.

The second type of functions are multi-body functions and are
holistic in nature. The computation of such functions involve more
than one atom's attributes and cannot produce final result in a
single run of the MS data (i.e., if traditional methods are used for
their computation). Such queries include the Radial Distribution
Function (RDF)~\cite{bamdad06, Frenkel:api01, AstroAnalysis02} as
well as some quantities associated with chemical
shifts~\cite{ChemShift_BCB98}. Generally, such functions are
computed through histograms. For instance, the RDF is obtained from
a histogram of all pairwise atom distances (this is the Spatial
Distance Histogram or SDH). The traditional, straightforward (often
the brute-force) way of computing these holistic functions is a very
time consuming process. On top of that, these methods cannot produce
the final result in a single run of the MS data, making such
functions unsuitable for our idea of a push-based system. However,
in our previous work, we have designed a data structure together
with an algorithm that opens up the possibility for such queries to
be executed in a push-based type environment. Further details on
this are given in Section~\ref{sc:system}.

Fig.~\ref{fg:querynetwork} represents an idea to show how such query
processing system can be improved. The idea behind it is that some
of the queries share same sub-routines. Having all the queries made
as separate modules, this sub-routines can be computed once the data
is being pushed through the system and then be used anytime a more
complex query needs it. Having in mind the amount of data in a
single frame that the queries (sub-routines as well) need to go
through, and the fact that in a single MS there are thousands of
frames, we believe this can be immense improvement in terms of total
running time.

\begin{figure*}
 \centerline{ \includegraphics[width=1\textwidth]{images/msstructure.eps} }
 \caption{MS Modules Structure}
 \label{fg:querynetwork}
\end{figure*}

\section{Building the system}\label{sc:system}

\subsection{MS Data retrieval and in memory organization}

A typical MS system generates and stores the data in a number of
trajectory files, usually including multiple frames (snapshots of
the simulated system taken at certain time intervals). Such MS
generated data oftentimes goes through a simple lossless compression
and, depending on the simulation software it may be stored in a
binary format. Such trajectory file format is one of most often used
MS file format (e.g., GROMACS, PDB). But such format is
unrecognizable to our system. So our system has to do three things
before it starts executing the queries: 1) Read the MS data from a
trajectory file, 2) Translate the MS data to a form recognizable to
our system, and 3) Load the data to memory.

\emph{\textbf{Data read - transform.}} As mentioned above, in order
for our system to be able to read the MS generated data, the data
needs to be transformed. The reason for this is following: the MS
data is stored in multiple files and possibly in different formats
as well. One such file holds the global data (identifying the system
and the simulation). Another file holds each frame's data. The frame
data contain general information about the frame, but the main part
is a sequential list of each atom's info, including atom's mass,
position, charge, number of electrons, velocities, forces, etc.
Another file (topology file) holds the molecule/residue info,
essentially identifying what atom belongs to which molecule.

So, in order to extract the data from these files, we have created a
sort of "extractor/transformer" of the attributes needed for the
execution of the queries in our system. This transformer,
essentially, is a separate piece of code that does three things: 1)
it reads the MS generated data that the MS system stores in one of
the often used MS file format (e.g., GROMACS); 2) it translates the
data into a format that our system can read (taking only the
information our system needs); and 3) it stores the data in a file
that has basic structure to it. So, in the end, the data transformer
produces a data file that our system takes as input. This code
serves as a connection between an MS system (e.g., GROMACS) and our
system. With this, our system can essentially be used as an add-on
to GROMACS or other simulation systems and help improve the
efficiency of the data analysis.

\textbf{\emph{Data organization in main memory.}}\footnote{This
paragraph talks only about the data organization used by one-body
queries. For two-body queries (e.g., SDH), the data organization is
discussed later on.} Once the data is in a format our system can
read, the data is being loaded into the main memory one frame at a
time. The organization of the in-memory particle's data is in the
form of a simple, two dimensional array where a single row
represents an atom in the system with all its attributes (e.g.,
coordinates, mass, charge, residue info, etc.). We also keep (in a
one dimensional array) crucial system's information for each frame,
like temperature, energy, pressure, etc. We have used such
structures because they are very suitable for simulating a
push-based type of system: a simple sequential read of the array
gives that on-line type of data stream. So, as the system reads the
array, it pushes the data onto the query-modules. As mentioned
earlier, the one-body (algebraic) queries will produce a final
result at the end of the first sequential read. However, the
two-body (holistic) functions, like SDH and RDF, cannot do this in a
single read of the data.\footnote{However, we have previously
deigned and created a data structure and an algorithm that can take
the advantage of a single data read and produce final results for
SDH computation~\cite{ytu:icde09,EDBT12}. We have incorporate this
into our system presented in this paper.} Therefore, each frame's
data array can be continuously read (in a loop manner) as many times
as a query needs it.
\subsection{Query modules}

As mentioned earlier, there are two types of functions/queries used
for analysis of MS systems: algebraic or one-body, and holistic or
two-body queries (these in general can be multi-body, but in this
paper we only deal with a two-body functions).

\subsubsection{One-body queries} Most of the query modules in
Table~\ref{tb:queries} (except the SDH and RDF) are not that
involved, only containing computations of fairly simple, one-body
functions. These queries were coded as separate modules in our
system. Each of these modules take few attributes as input (e.g.,
atom selection, frames selection(for the autocorrelation functions),
number of atoms, etc.). The system pushes the data as it becomes
available onto these modules. The queries are being executed on the
selection and are put in a ``ready'' mode, awaiting the next frame's
data. First, the more basic queries, like total mass, are being
computed. The results of such queries are temporary stored (in main
memory) and are available for use anytime a more complex query needs
them.

\subsubsection{Two-body queries} In general, queries involving two-body functions are a
bit more complex and cannot provide the final result in a single
data read if a straightforward method is used for their computation.
However, in the proposed system we have incorporated a data
structure and an algorithm for the SDH (also RDF) from our previous
work that is suitable for push-based type of system. In this
subsection we give a brief description of the data structure (DM)
and the algorithm (DM-SDH) designed in~\cite{ytu:icde09, EDBT12} and
implemented in the system proposed in this paper. For more detailed
information, please refer to our previously published work on this
topic~\cite{ytu:icde09, EDBT12}

\emph{\textbf{The data structure.}} The simulation data space is
represented by a conceptual data structure we named Density Map
(DM). The density map splits the simulation space into a grid of
equal size regions (or cells). The cells are cubes in 3D and squares
in 2D\footnote{In this paper, we focus only on the 2D data to
elaborate and illustrate the proposed ideas.} Resolution of a
density map is the reciprocal of the cell size in that density map.
In order to generate higher resolution density map, we split each
cell of the current resolution's grid into four smaller cells of
equal size. This design allows us to use a region
quad-tree~\cite{orenstein:ipl82} to organize density maps of the
same data but with different resolutions. So, essentially, a node in
the quad-tree represents a single cell from the DM. Therefore, a
density map of a certain resolution basically is the set of all
nodes of one level of the tree. Each of the tree nodes records the
cell's location in the density map (coordinates of corner points) as
well as the number of particles in each cell. We name the afore
describe tree the Density-Map tree (DM-tree).

\begin{figure}
 \centerline{ \includegraphics[width=0.65\columnwidth]{images/resolve.eps} }
 \caption{Computing minimum (i.e., length of solid lines) and
 maximum distance (i.e., length of dashed lines) range between two
cells}
 \label{fg:resolve}
\end{figure}

\emph{\textbf{The algorithm.}} The essential part of the DM-SDH
algorithm is a procedure named {\sc ResolveTwoCells}. The input to
this procedure are two cells from the density map (e.g., $A$ and $B$
in Fig.~\ref{fg:resolve}). It computes, in constant time, the
minimum and maximum distance between the two cells. A pair of cells
is ~\emph{resolvable} if both the min and max distance between them
fall into the same SDH bucket $i$. If that is the case, the distance
count of that bucket is being increase by $n_{A}n_{B}$ ($n_{A}$ and
$n_{B}$ are the number of particles in cell $A$ and $B$,
respectively). Otherwise, the cells are non-resolvable and we
either:
\begin{itemize}
\item[(1)] Go to the next density map with higher resolution and resolve all children of $A$
with those of $B$, or
\item[(2)] If leaf-level has been reached: compute every distance between particles of $A$ and $B$ and update the
histogram accordingly.
\end{itemize}

In order to generate the complete SDH, the {\sc ResolveTwoCells}
procedure is executed for all pairs of cells for a given density map
$DM_k$ and the algorithm would recursively call the procedure
(action (1) above) until leaf-level has been reached (action (2)
above).

On top of the aforementioned DM-SDH algorithm, we have also
incorporated two approximate SDH algorithms (ADM-SDH), introduced
and described in \cite{EDBT12}, and \cite{TKDE12}. These approximate
algorithms are substantially faster than the the brute-force
algorithm and also than the DM-SDH algorithm as they take advantage
of some heuristic. For more details on the ADM-SDH algorithms please
see the aforementioned work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Working of the system}

In this subsection we give a brief overview of how the system works
at runtime. Please note that the first, preliminary part is only
executed once, i.e., the data transformation from MS data files to a
file that our system can read.

Here are the steps taken through out the analysis:
\begin{itemize}
\item[(1)] Execute the data transformer
\begin{itemize}
\item[(a)] Read the MS data from trajectory files
\item[(b)] Extract the info needed for our system
\item[(c)] Save the read data to a file recognizable to the system
\end{itemize}
\item[(2)] Load the data into main memory (one frame at a time)
\begin{itemize}
\item[(a)] Load data into a double array (for one-body queries)
\item[(b)] Load data into the quad-tree structure (for two-body queries:SDH, RDF)
\end{itemize}
\item[(3)] Push the data to all queries
\item[(4)] A query, if available, acts upon the pushed data (first executing
the lower level, sub-queries)
\item[(5)] Store intermediate results (results
of sub-queries)
\item[(6)] Repeat steps 3-5 if needed.
\item[(7)] Output results
\item[(8)] Go to step 2 and load the next frame (if needed).
\end{itemize}

Fig.~\ref{fg:system} depicts the flow of the system.

\begin{figure}
 \centerline{ \includegraphics[width=.8\columnwidth]{images/system.eps} }
 \caption{Push-based system flow}
 \label{fg:system}
\end{figure}

Step 2, loading the data into main memory is different for SDH query
compared to the one for the one-body queries. The reason for that,
as mentioned earlier, is the different data structure used to store
the data in memory. While we use double array to store the data for
the one-body queries, we use quad-tree like data structure to store
the data needed to compute the SDH. The loading to the double array
is straightforward. However, to load the quad-tree structure, we
need to use some of the info from the data itself. Namely, the
coordinates of the atoms are used to determine in which tree node an
atom belongs. That way we build the so called density map (DM),
i.e., the different regions with a certain number of atoms in them
(including all the atom's attributes). So basically, to solve the
SDH problem in a push-based manner, we convert the problem into
populating a data structure in push-based manner. This data
structure will then be used as an input to our DM-SDH algorithm
that, although not completely in ``on-the-fly" way, is a great
improvement over the naive methods used in much of today's MS
analysis systems (i.e., GROMACS, PDB, CHARMM, etc.).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
 \centerline{ \includegraphics[width=.99\columnwidth]{images/workload_set_up1.eps} }
 \caption{Workload setup}
 \label{fg:workload_set_up}
\end{figure}

\section{Experimental results}\label{sc:experiments}

The system was implemented in C++ programming language and tested on
real molecular simulation data sets. The experiments were carried
out on an Apple MacPro machine with 8GB of physical memory and two
Quad-Core Intel Xeon 3GHz processors. The MacPro was running OS X
Mavericks $10.9.3$ operating system. We have compared the results
obtained by our system to those obtained by running the analysis
through GROMACS (v. $4.5.7$). Both systems were analyzing the same
data sets.

\emph{\textbf{Data sets.}} In our experiments, six data sets from
different simulations were used. All simulations were done on a
POPC~\footnote{POPC is a chemical compound composed of a
diacylglycerol and phospholipid. Its full name is
1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine and it is one of
the most important lipids in bio-physical molecular simulation.}
lipid bilayer, but were all set to produce data of different sizes
(i.e., different number of particles in the simulation). Namely, we
have tested the system on simulations with $52,400$; $209,600$;
$838,400$; $2.5M$; $4.4M$; and $8.8M$ atoms. Also, since the
simulations were run separately, they produced six different MS
systems with distinct characteristics (distinct structure, atom's
positioning, etc.). From all of the generated data sets we have
randomly selected sets of $10$, $100$, and $1000$ consecutive frames
for the purpose of our experiments. This gave us $18$ different
datasets on which we tested our system.

\emph{\textbf{Query work load.}} Two types of query workload were
used: 1) one involving one-body queries only 2) one including
two-body queries (SDH and RDF) as well. The reason for this is that
GROMACS, the system we used to compare our system to, only has a
naive method of solving the RDF (SDH) problem (like almost all MS
analysis systems). In our system we have incorporated SDH (RDF)
algorithms that are far more superior to the naive method, and
comparing the systems like that would not have been fair (we
believe).

\emph{\textbf{One-body queries only.}} The following set of one-body
queries were included in the test workload: mean square displacement
(msd), radius of gyration, dipole moment, center of mass, velocity
autocorrelation, electron density, mass density, and charge density.
This set of queries were pointed to us, by a group in the physics
field with extensive MS background, as one of the most commonly used
in the field of collagen bilayer MS system analysis. A workload
group contains all $8$ queries executed on one of the $12$
selections, making $12$ groups. Such groups are executed on six
different size data sets, with $10$, $100$ and $1000$ frames. This
workload is then repeated $5$ more times, by executing each of the
queries in the groups $5$, $10$, $25$, $50$, and $100$ times,
essentially just magnifying the workload intensity. In total, we
have $12 x 6 x 3 x 6 = 1,296$ different workload setups to test the
system on. Fig.~\ref{fg:workload_set_up} shows the organization of
the workload setup.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Benchmark}

Through extensive collaboration with a research group from the
Physics department at USF, we have come up with a benchmark that can
be used for testing the efficiency of an analysis system for
molecular simulations. The benchmark consist of three essential
parts: 1. Simulation data produced by an MS, 2. Queries that are to
be executed onto that data in order to produce some information of
interest, and 3.Benchmark parameters that control the size of the
benchmark.

\subsubsection{Benchmark Data}
The data used in the benchmark was real molecular simulation data,
produced through the GROMACS MS system. The initial, pre-simulation
data file consisted of $200$ POPC and $12000$ solvent molecules, or
$12200$ molecules in total. This type of system was used because it
is sufficiently diverse, containing enough distinct POPC and solvent
molecules (e.g., each POPC molecule includes approximately $52$
different atoms) and yet simple enough to be easily transformed into
another system of different size. By using the $genconf$ function in
GROMACS, we produced pre-simulation files of different sizes
(essentially by changing the system's size (box)). $Six$ different
sized pre-simulation files were created. A molecular simulation was
then run on these $6$ files, each producing an MS system of certain
size (volume/number of particles). All of the simulations were set
up to produce $1000$ frames (snapshots in time of the systems), each
frame containing the same number of particles as the base one. The
produced files contained: $52,400$; $209,600$; $838,400$; $2.5M$;
$4.4M$; and $8.8M$ atoms per frame. So, for example, the file with
$52,000$ atoms holds $52,000,000$ records in total ($1000$ frames,
each containing $52,000$ records). As mentioned earlier, this
simulation data comes mostly in binary formats and in trajectory
files having a lot of unneeded overhead. Therefore, it was
transformed to a data arrays files containing only crucial
information of the particles and the system. The size of the files
ranged from $135MB$ for $52,000$ atoms to $24GB$ for $8.8$ million
atoms (this is for data with $100$ frames).

%Fig.~\ref{fg:dataformat} represents the organization/structure of
%these files.

\subsubsection{Benchmark Queries}
The queries selected to be included in this benchmark were derived
through a thorough observation of the way an MS system is being
analyzed. They were found to be the base of the analysis of many MS
systems. In other words, no mater how small or big the analysis was,
these queries were included in that analysis. As mentioned earlier,
they are of two types: one-body (and algebraic) and two-boy (and
holistic). Table~\ref{tb:queries} shows these queries.

\subsubsection{Benchmark Parameters} There are several parameters
that can be used to control the overall size of the system. We divide the parameters into two groups:\\
Data size parameters:
\begin{itemize}
\item[-]Select different sized dataset
\item[-]Number of frames
\item[-]Data selection (within the selected dataset) onto which the queries are being executed
\end{itemize}
Workload size parameters:
\begin{itemize}
\item[-]Number of queries to be executed
\item[-]Number of times each query is executed
\end{itemize}
By changing these parameters, we can produce a versatile testing
benchmark for MS analysis systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames52K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results2-10frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames52K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results2-100frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames209K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results3-10frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames209K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results3-100frames}
\end{figure}

\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

%\begin{figure}
% \centerline{ \includegraphics[width=.8\columnwidth]{images/speedup10frames4.2M.eps} }
% \caption{Showing speedup for different workload}
% \label{fg:results5-10frames}
%\end{figure}

\subsection{Results} We have run extensive experiments over all the
different setups of workload mentioned previously in this section.
However, in this paper we present only the workload setups of $4$
data size sets: $838,400$, $2.5M$, $4.2M$, and $8.8M$ atoms because
we believe they convey enough information about the efficiency of
our system compared to that of the Gromacs system. The running times
of our push-based system were compared to those of the Gromacs
system. The first set of figures, namely
Figures~\ref{fg:speedup838K-select-level}-\ref{fg:speedup8.8M-select-level},
represent the speedup that our system obtains over the Gromacs
system with various atoms selection levels. We define the selection
levels based on the number of comparisons we have to make in order
to extract the needed group(selection) of atoms. For example, if we
want to do analysis on all molecules containing oxygen, or hydrogen,
or carbon we would go over each molecule and compare its components
to the selection list. The bigger the selection list, the higher the
select level in our system. For better visualization, we note three
different selection levels: high (at least $10$ comparisons made),
medium (between $1$ and $10$ comparisons made), and low select level
(with one or less comparisons made). As seen in the figures, for
high selection level, the speedup is smaller compared to that
achieved in low select levels. The reason for this, we believe is in
that the amount of time our system spends extracting the atoms group
increases with the level of selection. Even though our system still
shows considerable speedup over Gromacs in high level selections, we
do believe there is room for improvement in our system and that is
our immediate future work we are planning on doing. These figures
also show the relation of the speedup to the workload intensity,
i.e., the higher the workload intensity the higher the speedup.

The connection between the workload intensity and the speedup is
better represented in the next set of figures,
Figures~\ref{fg:speedup838K}-\ref{fg:speedup8.8M}. They show the
speedup our system achieves over the Gromacs system on a varying
workload intensity. Each of those figures show the speedup with
different dataset sizes (e.g., $838,000$, $2,567,600$ atoms, etc.),
including $10$, $100$, and $1000$ data frames. The speedup is
calculated simply as a ratio between the running time of our system
on a certain set of workload and that of the Gromacs system on the
same workload. These figures show that the speedup over varying
workload intensity achieved by our system ranges anywhere from about
$10$ to $1000$ times, depending on the size of the dataset, number
of frames and the selection of the atoms.

Figure~\ref{fg:speedup-average-over-all-workload}, shows the speedup
our system achieves over all workload intensity (average workload
intensity) with varying dataset sizes. It is clear that, again our
system has better performance than the Gromacs system. The speedup
presented in this set of figures ranges anywhere from about $15$ to
$650$ times.

The last figure, Figure~\ref{fg:speedup-average-workload-select},
shows the speedup our system achieves over all workload intensity
and all select levels with varying dataset sizes. This figure, in a
way, summarizes the previous two sets of figures, bringing together
the workload and the different selections through the average. It is
clear that, again our system has better performance than the Gromacs
system. The speedup ranges anywhere from $50$ to $250$.

All four sets of figures show that such push-based design has clear
advantages over the pull-based type of design incorporated in the
Gromacs system.

\section{Conclusions and Future Work}\label{sc:conclusion}
The objective of our work is to design and implement improved data
analysis system that can be used in the field of molecular
simulation system's analysis. In this paper, we introduce the idea
for such system. We build our system on a push-based type design,
where data from data arrays is being pushed onto available queries
in the system. These queries are being executed on the pushed data
and produce intermediate / final result that would be used as part
of the data analysis. We are able to achieve an improvement over
existing, pull-based type designs because of the I/O overhead such
designs introduce when dealing with large volumes of scientific
data. Also, our queries are able to be executed on the same stream
of data, making it suitable solution for streaming circumstances. We
designed a benchmark that can be used to test data analysis systems.
This benchmark comprises of three parts: 1)benchmark data,
2)benchmark queries, and 3) benchmark parameters. We use this
benchmark to compare our system to one of the most frequently used
MS analysis systems, Gromacs. The efficiency and speedup achieved by
our system is supported by extensive experiments and their results.
The results show that our push-based design achieves up to about
$1000$ times speedup in comparison to a pull-based design, i.e.,
Gromacs.

One direction of our future work will be to further improve our
push-based design. Through the extensive experiments we have learned
that our design can be improved when the atom selection clause
involves many conditions. This improvement may be in the direction
of improving the algorithmic part, but it can also be in the
direction of improving the data presentation/organization we have
used in the system.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\noindent{\bf Acknowledgements:} The project described was supported
by an Award (R01GM086707) from the National Institute Of General
Medical Sciences (NIGMS) at the National Institutes of Health (NIH).
The authors would like to thank Anand Kumar who has contributed his
time and knowledge toward completion of the work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{IEEEtran}
\bibliography{push-based-MS}
\begin{comment}
\begin{IEEEbiographynophoto}{Vladimir Grupcev}
received a BS degree in Applied Mathematics \& Computer Science from
University of Ss. Cyril and Methodius, Macedonia in 2005; Master's
degree in Mathematics from University of South Florida in 2007. He
is currently a PhD student in the department of Computer Science \&
Engineering at the University of South Florida. His area of interest
includes scientific data management and high performance computing.
\end{IEEEbiographynophoto}
\vspace{-10mm}
\begin{IEEEbiographynophoto}{Yi-Cheng Tu}
received a Bachelor's degree in horticulture from Beijing
Agricultural University, China, and MS and PhD degrees in computer
science from Purdue University (2003; 2007). He is currently an
assistant professor in the department of Computer Science \&
Engineering at the University of South Florida. His research is in
energy-efficient database systems, scientific data management, high
performance computing and data stream management systems.
\end{IEEEbiographynophoto}
\vspace{-10mm}
\begin{IEEEbiographynophoto}{Sagar Pandit}
Dr. Sagar Pandit is an Assistant Professor of Physics at the
University of South Florida. He received his Ph.D. from the
Department of Physics, University of Pune, in 1999 in the area of
Dynamical systems. Later, his research shifted to Biological
Physics, specifically membrane physics. His current research
interests include  biomembranes, mathematical modeling of complex
biological and social systems, dynamical systems, and computational
approaches toward addressing problems in these fields.
\end{IEEEbiographynophoto}
\end{comment}

% if you will not have a photo
%\begin{biographynophoto}{John Doe}
%Biography text here.
%\end{biographynophoto}

% insert where needed to balance the two columns on the last page

\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendices

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

### 2016-01-20 paper/paperSource.tex
some changes to the paper (abstract added)

@@
 
-Thanks to the advancement of the modern computer simulation systems
-and state-of-the-art instrumentation, many scientific fields
-generate, and require manipulation of large amounts of data. Known
-scientific data analysis systems, as well as traditional DBMSs,
-follow a pull-based type of architectural design, where the executed
-queries mandate the data needed on their part. This design, while
-suitable for traditional transaction-based workloads where number of
-queries retrieve small parts of data located at various places of
-the database, is ill-fitted for applications involving complex
-analysis on most of the data. Such design involves random I/O,
-considerably affecting the data throughput in the system. All of
-this gives rise of redundant and high I/O traffic in the process of
-scientific data analysis in pull-based type systems. In this paper,
-we design and implement a push-based type system which allows
-high-throughput data analysis in the process of scientific
-discovery. Our design improves on the aforementioned shortcomings in
-two ways: i) it uses a sequential scan-based I/O framework that
-loads the data into the main memory, and then ii) the system pushes
-the loaded data to a number of pre-programmed queries. These queries
-are being executed on the same data stream, with various data
-selections. This way, essentially, the system lowers the unnecessary
-I/O overhead imposed by the randomized, index-based scan and that of
-a multiple data reads if each query were to be fed separately.
-Considering the amount of MS data as well as the number of queries
-being executed, we believe our system provides substantial
-improvement over the current MS data analyzing systems. The
-efficiency of the proposed system is backed by the results of
-extensive experiments using MS generated data. The running times of
-our system are compared to those of the GROMACS system. The
-comparison shows the advantage and the potential of using such
-push-based system for MS system analysis.
+Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
+Traditional database management systems generally follow principle of pulling the needed data, processing it, and then returning the results.
+This approach is then optimized by means of caching, storing in different structures, or doing some sacrifices on precision of the results to make it faster.
+When it comes to the point of doing various queries that require analysis  of the whole data,  this design has the following disadvantages: considerable overhead on random I/O while reading from the simulation output files and low throughput of the data that consequently results in long latency, and, if there was any indexing to optimize selections, overhead of storing those becomes too big, too.
+
+Thankfully, there is a relatively new approach to this problem -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and it's advantages are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filter the data and collect all the needed information.
+It has a considerable advantage in analysis of molecular simulation data, because it normally involves all the data sets to be processed by the queries.
+
+In this paper, we propose improved version of Push-based System for Molecular Simulation Data Analysis.
+Its major difference with the previous design is usage of GPU for the actual processing part of the data flow.
+In the old approach there were some custom data structures such as quad-tree for calculation of histograms to make the processing faster, but that might involve loss in data and some expectations from the type of data, too.
+In the new approach due to high performance of GPU processing and its nature, custom data structures were not needed in order to make it faster and still keep precision.
 
@@
 Push-based system, molecular simulation, scientific databases,
-spatial distance histogram.
+spatial distance histogram, GPU, parallel processing.
 \end{IEEEkeywords}

### 2016-01-21 paper/paperSource.tex
paper changed

@@
-\documentclass[10pt,journal,final,letterpaper,twocolumn]{IEEEtran}
+\documentclass[10pt,journal,final,letterpaper,twocolumn,conference]{IEEEtran}
 %\documentclass{ssdbm}
@@
 \usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
-\usepackage{epstopdf}
+\usepackage{epstopdf,filecontents,lipsum}
 %\usepackage{fixltx2e}
@@
 \thanks{$\dagger$ Author to whom all correspondence should be sent.}
-\thanks{Vladimir Grupcev and Yi-Cheng Tu are with the Department of Computer
-Science and Engineering, University of South Florida, 4202 E. Fowler
-Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
-vgrupcev@mail.usf.edu, ytu@cse.usf.edu}
+\thanks{Iliiazbek Akhmedov and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E. Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
+akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
 \thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
@@
 \section{Introduction}\label{sc:intro}
+
+In various sciences simulation systems take big place and often times they may be very important for results.
+One of such sciences, which is primarily related to this paper, is physics.
+In this case, from computer engineering point of view, simulations generally have the following flow:
+
+\vspace{5mm}
+
+{\small
+\begin{enumerate}
+	\item initial physical properties are given to simulation software as arguments which are interpreted and provided in a language defined by the software
+	\item then it runs for given amount of time that can generally last from milliseconds to months
+	\item finally, the file generated during the simulation is analyzed to extract useful information
+\end{enumerate}
+}
+
+\vspace{3mm}
+
 The amount of data generated by scientific applications running on
@@
 \bibliographystyle{IEEEtran}
-\bibliography{push-based-MS}
-\begin{comment}
-\begin{IEEEbiographynophoto}{Vladimir Grupcev}
-received a BS degree in Applied Mathematics \& Computer Science from
-University of Ss. Cyril and Methodius, Macedonia in 2005; Master's
-degree in Mathematics from University of South Florida in 2007. He
-is currently a PhD student in the department of Computer Science \&
-Engineering at the University of South Florida. His area of interest
-includes scientific data management and high performance computing.
-\end{IEEEbiographynophoto}
-\vspace{-10mm}
-\begin{IEEEbiographynophoto}{Yi-Cheng Tu}
-received a Bachelor's degree in horticulture from Beijing
-Agricultural University, China, and MS and PhD degrees in computer
-science from Purdue University (2003; 2007). He is currently an
-assistant professor in the department of Computer Science \&
-Engineering at the University of South Florida. His research is in
-energy-efficient database systems, scientific data management, high
-performance computing and data stream management systems.
-\end{IEEEbiographynophoto}
-\vspace{-10mm}
-\begin{IEEEbiographynophoto}{Sagar Pandit}
-Dr. Sagar Pandit is an Assistant Professor of Physics at the
-University of South Florida. He received his Ph.D. from the
-Department of Physics, University of Pune, in 1999 in the area of
-Dynamical systems. Later, his research shifted to Biological
-Physics, specifically membrane physics. His current research
-interests include  biomembranes, mathematical modeling of complex
-biological and social systems, dynamical systems, and computational
-approaches toward addressing problems in these fields.
-\end{IEEEbiographynophoto}
-\end{comment}
-
-% if you will not have a photo
-%\begin{biographynophoto}{John Doe}
-%Biography text here.
-%\end{biographynophoto}
-
-% insert where needed to balance the two columns on the last page
+\bibliography{references.bib}
 

### 2016-01-21 paper/paperSource.tex
Paper

@@
-\documentclass[10pt,journal,final,letterpaper,twocolumn,conference]{IEEEtran}
+\documentclass[10pt,journal,final,letterpaper,twocolumn]{IEEEtran}
 %\documentclass{ssdbm}
@@
 \usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
-\usepackage{epstopdf,filecontents,lipsum}
 %\usepackage{fixltx2e}
@@
 One of such sciences, which is primarily related to this paper, is physics.
-In this case, from computer engineering point of view, simulations generally have the following flow:
+In this case, from computer engineering point of view, simulations generally have the following flow \cite{mainPaper}:
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{references.bib}
+\bibliography{references}
 

### 2016-01-24 paper/paperSource.tex
paper changes...

@@
 % paper title
-\title{Parallelization of Push-based System for Molecular Simulation Data Analysis}
+\title{Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU}
 
@@
 Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
-Traditional database management systems generally follow principle of pulling the needed data, processing it, and then returning the results.
+Traditional database management systems follow principle of pulling the needed data, processing it, and then returning the results.
 This approach is then optimized by means of caching, storing in different structures, or doing some sacrifices on precision of the results to make it faster.
@@
 
-Thankfully, there is a relatively new approach to this problem -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and it's advantages are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filter the data and collect all the needed information.
+There is a new approach to this problem presented in the previous paper -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and its primary steps are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filter the data and collect all the needed information which increases efficiency and data throughput.
 It has a considerable advantage in analysis of molecular simulation data, because it normally involves all the data sets to be processed by the queries.
@@
 Its major difference with the previous design is usage of GPU for the actual processing part of the data flow.
-In the old approach there were some custom data structures such as quad-tree for calculation of histograms to make the processing faster, but that might involve loss in data and some expectations from the type of data, too.
-In the new approach due to high performance of GPU processing and its nature, custom data structures were not needed in order to make it faster and still keep precision.
+Using the same scan-based I/O framework the data is pushed through the network of queries which are processed by GPU, and due to the nature of science simulation data, this gives a big advantage for processing it faster and easier.
+In the old approach there were some custom data structures such as quad-tree for calculation of histograms to make the processing faster and those involved loss of data and some expectations from the data nature, too.
+In the new approach due to high performance of GPU processing and its nature, custom data structures were not even needed much though it didn't bear any loss in precision and performance.
 
@@
 Push-based system, molecular simulation, scientific databases,
-spatial distance histogram, GPU, parallel processing.
+spatial distance histogram, GPU, parallel processing, CUDA.
 \end{IEEEkeywords}
@@
 
-\begin{comment}
-Many scientific applications share several features that bring in
-light the shortcomings of the current data analysis systems. First,
-the scientific analysis often involves executing a number of
-analytical queries commonly processed on a large portion, if not all
-of the generated data. Second, many of these fields use the same
-popular, and small number of data analytics primitives as the basic
-blocks on which they build their discovery. And three, the
-scientific data, once stored is never modified. New data only
-appends to the existing dataset, making the data perfect contender
-for streaming.
-
-And while the use of an index-based scan might seem better option to
-a sequential scan, this is the case only when the query accesses
-small portion of the data.
-\end{comment}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-In various sciences simulation systems take big place and often times they may be very important for results.
+In various sciences simulation systems take big place and often times they may be the clue for results.
 One of such sciences, which is primarily related to this paper, is physics.
-In this case, from computer engineering point of view, simulations generally have the following flow \cite{mainPaper}:
+In this case, from computer engineering point of view, simulations have the following flow:
 
@@
 \begin{enumerate}
-	\item initial physical properties are given to simulation software as arguments which are interpreted and provided in a language defined by the software
-	\item then it runs for given amount of time that can generally last from milliseconds to months
+	\item initial physical properties are given to simulation software as arguments which are interpreted and provided in a language defined by the simulation software
+	\item then it runs for given amount of time that can generally last from milliseconds to months or even more
 	\item finally, the file generated during the simulation is analyzed to extract useful information
@@
 }
-
 \vspace{3mm}
-
-The amount of data generated by scientific applications running on
-advanced computer simulation systems as well as the new improved
-experimental devices is growing rapidly. This big volume of data is
-imposing a significant stress on the data analysis software. Even
-though the existing data analysis systems are configured to deal
-with large volume of data, they are not optimized for high
-throughput data analysis. Often times, the data is available only
-for certain period of time (e.g., streams) so the analyzing system
-should react to it in a very efficient way. On top of the big amount
-of data, another challenge that is being imposed to the data
-analysis systems is that the data is usually being accessed through
-high-level analytical queries which computation is a lot more
-complex than the computation of simple aggregates. Some of these
-queries are usually the bottleneck of the data analysis systems,
-because they take a lot of time (often many days) to be executed and
-the systems are not designed to handle multiple queries on the same
-data stream at the same time, thus decreasing the overall efficiency
-of the data analysis.
-
-One such example, where fast data analysis is crucial involves
-online social
-media~\cite{nature_bigdata08,nature_bigdata12,science_social10}. The
-social media is known to be used for real-time feedback collection,
-which is usually associated to a particular topic or
-product~\cite{jcs_twitter11}. In order for a system to be able to
-perform analytical examination of the data produced in such
-streaming media, the system should have the capability of fast data
-access. The reason, the millions of data records(tweets) produced
-every second~\cite{tkde_big14}. Moreover, these tweets may have
+The purpose of this paper is primarily focused on the third step, which basically involves the entire simulation data to be processed by the analysis software.
+
+Big data processing is becoming one of the key issues with the amount of data being generated by modern systems.
+Eventually, this data is required to be processed on the fly, thus, analysis software should be able to handle massive data in a very short period of time.
+When working with huge volume of data, there are non-trivial issues arise.
+For example, it can take days and weeks to analyze big enough data sets, because the data cannot be simply loaded into memory, since it can get up to terabytes, thus, it has extra overhead because of widely used random access disk I/O framework to read from disk chunks by chunks.
+Besides it, analysis of the data can get even more complicated, since going through certain parts of the data once would not be enough, which leads to low throughput and efficiency of loaded data.
+One such example, the data analysis approach has polynomial complexity just for reading the data in order to come up with result, and since the data can't simply be saved in memory, it raises the overhead of disk I/O, too.
+Pull-based architectures in data processing engines are inefficient, since having a set of specific queries, in order to compute them all, it is needed to fetch data, filter it, and apply needed formulas.
+It is inefficient, since for every query the same chunk of data needs to be pulled into the memory at least the number of queries times or more in case of more complex queries.
+
+As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks. .
+"In order for a system to be able to perform analytical examination of the data produced in such streaming media, the system should have the capability of fast data access.
+The reason, the millions of data records(tweets) produced
+every second. Moreover, these tweets may have
 different geographical origin, introducing different languages and
@@
 be considered and possibly incorporated in the process of analytical
-investigation in order to achieve relevant result.
-
-The aforementioned big data imposed issues are also frequently
-present in many other application fields, e.g., scientific data
-analysis. In many scientific and engineering fields, the main
-computational method for analyzing physical/chemical attributes of
-natural systems is the particle simulation. Such simulations, when
-performed in the field of structural and molecular biology, are
-commonly called Molecular Simulations. Molecular simulations (MS)
-(sometimes called Molecular Dynamics (MD) simulations) are computer
-simulations of complex biological, physical or chemical structures.
-These simulations, based on some basic theoretical models are widely
-utilized as a basic research mechanism for studying the behavior of
-the natural systems. The basic units of such systems/simulations are
-natural particles (such as atoms, molecules, stars, etc.) and they
-interact among each other for a certain period of time following
-postulated classical forces. The number of these basic units taking
-part in the simulations is big, usually in the range of hundreds of
-thousands to millions. For instance, Fig.~\ref{fg:snap-white} shows
-a snapshot of a collagen fiber, consisting of $890,000$ particles.
-And oftentimes, the simulations produce datasets consisting of more
-than one snapshot (frame) of the system's current state at various
-time instants. Each of these frames contain all the particles,
-together with all their measurements, such as spatial coordinates,
-mass, charge, velocity, forces, etc. And usually, a big number of
-such frames (in the tens of thousands) are being produced and stored
-in the course of a typical simulation. Quantities measured during
-the simulations are analyzed to test the theoretical
+investigation in order to achieve relevant result. "\cite{mainPaper, nature_bigdata08,nature_bigdata12,science_social10,jcs_twitter11}
+
+The primary focus and problem in this paper is scientific data analysis.
+Particles simulation is one of the most popular methods of analyzing certain chemical reactions, physical processes, or other behavior of different materials.
+Molecular simulations (Molecular Dynamics) are applied in different fields and represent a method of analyzing physical movements of particles, atoms, and molecules in a fixed space with a given period of time, apparently with a possibility of giving initial state for each item that is involved in the process and can affect the system.
+This system is an N-body simulation.
+The number of atoms in simulations vary in hundred of thousands, particularly, we may observe two simulation systems of a collagen fiber structure and dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system consisting of 890,000 and 402,400 atoms respectively on Figure \ref{fg:collagen_dipal}.
+Simulation data represents number of records of physical properties such as mass, charge, velocity, coordinates, and forces for each item aggregated as frames, where each frame represents a snapshot of time, placed with a fixed time interval which may also vary depending on the simulation itself and simulation precision requirement.
+"Quantities measured during the simulations are analyzed to test the theoretical
 model~\cite{Frenkel:api01,Landau:cup05}. In short, the MS is proven
@@
 biophysical and biochemical processes that are being unfold at a
-nanoscopic scale.
+nanoscopic scale." \cite{mainPaper}
+
+Scientist gives the properties to simulation software (for example, Gromacs), runs the simulations, and finally get the output file.
+The output file must then be analyzed to produce certain results which may help him come up with certain consensus on original theoretical model that resulted in the molecular simulation system\cite{Frenkel:api01}.
+Gromacs is simulation software tool that helps scientist to run the actual simulation.
+It is a molecular dynamics package primarily designed for biomolecular systems such as proteins and lipids.
+\cite{Gromacs-online}.
+Besides the fact that it helps to generate the output files for the simulations, apparently it also helps to analyze the data itself, but the original problem is that it is not as optimized as it can be in order to analyze the data.
+Gromacs follows approach of pull-pased design, which means that for any given query (e.g.
+total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In order to achieve objective discovery and to explain the workings
+of the scientific systems, scientist must analyze the data produced
+by the MS. These analysis oftentimes comprise of computation of very
+complex quantities that show statistical properties of the data.
+Such queries are of great importance to scientist because they are
+the basic assembly blocks for a series of critical quantities needed
+to outline the scientific systems \cite{Frenkel:api01}. The methods
+by which the data is being accessed as well as these queries are
+executed can either increase or decrease the efficiency of an data
+analysis system. The known data analysis systems (e.g., Gromacs
+\cite{Gromacs-online}) uses pull based type design in which the data
+is being fed to the queries only by demand. When a query needs the
+data, it requests it, the data is pulled through the system and
+delivered to the query. This type of design, we believe, introduces
+two types of problematic issues: 1. overhead in I/O traffic by not
+allowing for multiple queries to be executed on the same stream of
+data at the same time, and 2. cpu/data latency incurred when a
+request for the data is sent to the system (the time it takes from a
+request to the time the query can actually use the data).
+
+We believe that by using a push-based type system we can remedy such
+issues. In the push-based design the queries do not request the
+data. Rather, the data is pushed onto the queries automatically and
+is being processed by the active queries. This alleviates the I/O
+traffic overhead as well as the cpu/data latency incurred for each
+query that would have been introduced by a pull-based design. Such a
+system design is especially suitable for many scientific
+applications which share the following features. First, the
+scientific analysis often involves executing a number of analytical
+queries commonly processed on a large portion, if not all of the
+generated data. Second, many of these fields use the same popular,
+and small number of data analytics primitives as the basic blocks on
+which they build their discovery. And three, the scientific data,
+once stored is never modified. New data only appends to the existing
+dataset, making the data perfect contender for streaming.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+In this paper, we describe the idea, design and implementation of
+such a push-based system. We would like to point out that the work
+presented in this paper is an application of such push-based
+strategy that has been previously considered in the database
+community in the works presented in~\cite{DataPath,Volcano,Qpipe},
+as well as the idea of sharing same tuple of data or data scans
+across many queries previously introduced
+in~\cite{Candea,PredictablePerformance,CooperativeScans}.
+
+To summarize, data-intensive applications often require compelling
+amount of storage space and intensive processing capability, but
+also need fast data access and high throughput data analysis.
+Therefore, the need of a system that will be optimized to access
+data fast, with high throughput, as well as efficiently execute the
+analytical queries with a possibility of running multiple queries on
+the same data stream is of a great importance to the scientific
+community.
+
+\begin{figure}
+ \centerline{ \includegraphics[width=1\columnwidth]{images/sample_snapshot_of_simulation.eps} }
+ \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle} }
+ \label{fg:collagen_dipal}
+\end{figure}
 
@@
 \bibliographystyle{IEEEtran}
-\bibliography{references}
+\bibliography{references.bib}
 

### 2016-01-24 paper/paperSource.tex
paper changes...

@@
 
-\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$% <-this % stops a space
+\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$, Vladimir Grupcev, Joseph Fogarty, Sagar Pandit % <-this % stops a space
 
@@
 \thanks{$\dagger$ Author to whom all correspondence should be sent.}
-\thanks{Iliiazbek Akhmedov and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E. Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A. Emails:
+\thanks{Iliiazbek Akhmedov, Vladimir Grupcev, and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E.
+Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A.
+Emails:
 akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
@@
 Gromacs follows approach of pull-pased design, which means that for any given query (e.g.
-total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In order to achieve objective discovery and to explain the workings
-of the scientific systems, scientist must analyze the data produced
-by the MS. These analysis oftentimes comprise of computation of very
-complex quantities that show statistical properties of the data.
-Such queries are of great importance to scientist because they are
-the basic assembly blocks for a series of critical quantities needed
-to outline the scientific systems \cite{Frenkel:api01}. The methods
-by which the data is being accessed as well as these queries are
-executed can either increase or decrease the efficiency of an data
-analysis system. The known data analysis systems (e.g., Gromacs
-\cite{Gromacs-online}) uses pull based type design in which the data
-is being fed to the queries only by demand. When a query needs the
-data, it requests it, the data is pulled through the system and
-delivered to the query. This type of design, we believe, introduces
-two types of problematic issues: 1. overhead in I/O traffic by not
-allowing for multiple queries to be executed on the same stream of
-data at the same time, and 2. cpu/data latency incurred when a
-request for the data is sent to the system (the time it takes from a
-request to the time the query can actually use the data).
-
-We believe that by using a push-based type system we can remedy such
-issues. In the push-based design the queries do not request the
-data. Rather, the data is pushed onto the queries automatically and
-is being processed by the active queries. This alleviates the I/O
-traffic overhead as well as the cpu/data latency incurred for each
-query that would have been introduced by a pull-based design. Such a
-system design is especially suitable for many scientific
-applications which share the following features. First, the
-scientific analysis often involves executing a number of analytical
-queries commonly processed on a large portion, if not all of the
-generated data. Second, many of these fields use the same popular,
-and small number of data analytics primitives as the basic blocks on
-which they build their discovery. And three, the scientific data,
-once stored is never modified. New data only appends to the existing
-dataset, making the data perfect contender for streaming.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In this paper, we describe the idea, design and implementation of
-such a push-based system. We would like to point out that the work
-presented in this paper is an application of such push-based
-strategy that has been previously considered in the database
-community in the works presented in~\cite{DataPath,Volcano,Qpipe},
-as well as the idea of sharing same tuple of data or data scans
-across many queries previously introduced
-in~\cite{Candea,PredictablePerformance,CooperativeScans}.
-
-To summarize, data-intensive applications often require compelling
-amount of storage space and intensive processing capability, but
-also need fast data access and high throughput data analysis.
-Therefore, the need of a system that will be optimized to access
-data fast, with high throughput, as well as efficiently execute the
-analytical queries with a possibility of running multiple queries on
-the same data stream is of a great importance to the scientific
-community.
+total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result just for a single query.
+As it has been proposed in the previous paper, in order to remedy such issues, the push-based design does exactly the opposite, where instead of loading the data on demand for each query, the queries are batched into a network, then the entire dataset is loaded chunk by chunk pushing it through the network which has its internal relationships and dependencies amonth the queries.
+In this case, since scientific simulation data is run once with specific physical properties, it is never modified, thus, on continuation, it will only append, which means that the processing can also be run on appended frames.
+\cite{mainPaper}
+This type of approach has already been revised by other systems \cite {DataPath,Volcano,Qpipe} as of reusing loaded data through queries produced \cite {Candea,PredictablePerformance,CooperativeScans}.
+
+In this paper, we are incorporating parallelization into the processing part of the proposed design by means of CUDA programming language for GPU.
+Since storage of the simulation data is very expensive, it might come to the point of analyzing the data on the fly (meaning running the simulation and analyzing it at the same time in a streaming manner), which leads to a problem of optimizing the processing part of the design proposed in the previous paper, because time spent on generating data should be tried to conceal the time spent on the processing part by means of overlapping or simply running it in a quick manner.
 
@@
  \label{fg:collagen_dipal}
-\end{figure}
-
-In order to achieve objective discovery and to explain the workings
-of the scientific systems, scientist must analyze the data produced
-by the MS. These analysis oftentimes comprise of computation of very
-complex quantities that show statistical properties of the data.
-Such queries are of great importance to scientist because they are
-the basic assembly blocks for a series of critical quantities needed
-to outline the scientific systems \cite{Frenkel:api01}. The methods
-by which the data is being accessed as well as these queries are
-executed can either increase or decrease the efficiency of an data
-analysis system. The known data analysis systems (e.g., Gromacs
-\cite{Gromacs-online}) uses pull based type design in which the data
-is being fed to the queries only by demand. When a query needs the
-data, it requests it, the data is pulled through the system and
-delivered to the query. This type of design, we believe, introduces
-two types of problematic issues: 1. overhead in I/O traffic by not
-allowing for multiple queries to be executed on the same stream of
-data at the same time, and 2. cpu/data latency incurred when a
-request for the data is sent to the system (the time it takes from a
-request to the time the query can actually use the data).
-
-We believe that by using a push-based type system we can remedy such
-issues. In the push-based design the queries do not request the
-data. Rather, the data is pushed onto the queries automatically and
-is being processed by the active queries. This alleviates the I/O
-traffic overhead as well as the cpu/data latency incurred for each
-query that would have been introduced by a pull-based design. Such a
-system design is especially suitable for many scientific
-applications which share the following features. First, the
-scientific analysis often involves executing a number of analytical
-queries commonly processed on a large portion, if not all of the
-generated data. Second, many of these fields use the same popular,
-and small number of data analytics primitives as the basic blocks on
-which they build their discovery. And three, the scientific data,
-once stored is never modified. New data only appends to the existing
-dataset, making the data perfect contender for streaming.
-
-In this paper, we describe the idea, design and implementation of
-such a push-based system. We would like to point out that the work
-presented in this paper is an application of such push-based
-strategy that has been previously considered in the database
-community in the works presented in~\cite{DataPath,Volcano,Qpipe},
-as well as the idea of sharing same tuple of data or data scans
-across many queries previously introduced
-in~\cite{Candea,PredictablePerformance,CooperativeScans}.
-
-To summarize, data-intensive applications often require compelling
-amount of storage space and intensive processing capability, but
-also need fast data access and high throughput data analysis.
-Therefore, the need of a system that will be optimized to access
-data fast, with high throughput, as well as efficiently execute the
-analytical queries with a possibility of running multiple queries on
-the same data stream is of a great importance to the scientific
-community.
-
-\begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/snap-white.eps} }
- \caption{A snapshot of simulated collagen fiber structure}
- \label{fg:snap-white}
 \end{figure}

### 2016-01-24 paper/paperSource.tex
paper changes...

@@
 Besides the fact that it helps to generate the output files for the simulations, apparently it also helps to analyze the data itself, but the original problem is that it is not as optimized as it can be in order to analyze the data.
-Gromacs follows approach of pull-pased design, which means that for any given query (e.g.
+Gromacs follows approach of pull-based design, which means that for any given query (e.g.
 total mass or total charge, which are very similar type of 1-body queries without sophisticated selection) it will pull data separately and generate addition overhead wasting disk read I/O in order to come up with the result just for a single query.
@@
 \subsection{Problem Statement}
-As mentioned earlier, in order to describe the scientific system,
-scientist must look into/analyze the critical statistical properties
-of the data produced by the simulation. These properties are usually
-computed through series of queries that are being executed on the
-whole, or some selection of the MS data. Most of the systems that
-are widely used today across many scientific fields for analysis of
-the MS data are part of a software systems that also run the
-simulation. Such examples include GROMACS~\cite{GROMACS4},
-VMD~\cite{VMD}, MDAnalysis~\cite{MDAnalysis}, Wordom~\cite{wordom},
-MD-TRACKS~\cite{MDtracks}, SimulaidOne~\cite{Simulaid},
-Charmm~\cite{CHARMM}. Once the simulation is run, the system
-produces a flat file with all the measurements of each particle.
-These systems take each query as a user's input and apply them to
-the data uploaded from these flat files (called trajectories in
-GROMACS). The biggest problem/performance issue that these system
-impose is the fact that for every new query issued by the user, the
-system has to load a significant part of the dataset into main
-memory before executing the query. Such pull-based design involves
-random I/Os that considerably affecting the data throughput in the
-system. And while the use of an index-based scan might seem better
-option to a sequential scan, this is the case only when the query
-accesses small portion of the data. All of this gives rise of
-redundant and high I/O traffic in the process of scientific data
-analysis in pull-based type systems. This adds greatly to the
-overall performance time of the analysis of the MS system. Having in
-mind the volume of data produced by the MS in a single frame (in
-gigabytes), and we add the number of frames simulated (tens of
-thousands) the total volume of data to be analyzed is of huge
-magnitude. So to have to load big chunks of this data from disk to
-the main memory every time a query is being executed degrades the
-overall system's performance greatly. In this paper we design and
-implement a push-based type system which allows high-throughput data
-analysis in the process of scientific discovery.
-
-\subsection{Our Approach}
+Simulation software systems, in general, follow the same methodology of running and storing simulation data.
+The simulation software system examples are: Gromacs \cite{GROMACS4}, VMD\cite{VMD}, MDAnalysis~\cite{MDAnalysis}, Wordom~\cite{wordom}, MD-TRACKS~\cite{MDtracks}, SimulaidOne~\cite{Simulaid}, Charmm~\cite{CHARMM}.
+In the type of simulations brought up as examples above, the flow of the data is the following.
+Once the simulation is run, the output files are contained as trajectory files with descriptors (they contain information about space dimensions, number of atoms and frames, etc.) that can be easily transposed into simple flat files containing the physical properties atom by atom, frame by frame, which are consequently read and processed by the proposed push-based system.
+Since we have certain amount of queries needed to be run on given simulation data, generating high I/O traffic followed by design of pull-based system is not considerable.
+The approach proposed in the previous paper is very good in terms of performance in comparison with the original pull-based system\cite{mainPaper}.
+The problem is still that some of the queries processed by those means are still improvable, especially taking into consideration the fact that in the used previous works for calculation of 2-body functions, which take the biggest time for processing~\cite{ytu:icde09, EDBT12}, we might have some error bounds, which might be unacceptable for certain simulation analysis where sacrifice on loss of data is unbearable.
+Besides it, although we might have huge performance boost on specific data structure as density map, we still have certain expectation from data nature (such as its uniformity), thus, it makes sense to have desire to simplify the processing and still having pure computation based on any values, considering that the queries should be available to be pre-programmed by user in a separate module of the code.
+
+\begin{figure}
+ \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4-eps-converted-to.pdf} }
+ \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainArticle} }
+ \label{fg:sample_estimation_old_paper}
+\end{figure}
+
+\subsection{Our approach with improvement}
+
+Reading the data from files is a huge overhead, thus, speedup of push-based system in general is pretty significant over pull-based system in our given case. 
+It's been demonstrated in the previous paper with different amount of atoms, frames, and workloads \cite{mainPaper}.
+Since we have to load the data every time there is a different query versus push-based system loads a chunk of data once, it is very noticable how the framework \cite{mainPaper} contributes to efficiency of analysis of simulation data.
+You can observe some estimation samples in Figure \ref{fg:sample_estimation_old_paper}
+
+On the other hand, some of the queries may be very slow due to their nature on a sequential type of computation,  especially, 2-body functions.
+Since the nature of the data that comes with simulation is basically physical properties of atoms, thus, there is a lot of computation that involves independent primitive mathematical operations, which makes it a perfect problem for parallelism.
+
 Considering the fact that the running time of the most queries used

### 2016-01-26 paper/paperSource.tex
paper changes

@@
  \centerline{ \includegraphics[width=1\columnwidth]{images/sample_snapshot_of_simulation.eps} }
- \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle} }
+ \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle}}
  \label{fg:collagen_dipal}
@@
  \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4-eps-converted-to.pdf} }
- \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainArticle} }
+ \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainPaper}}
  \label{fg:sample_estimation_old_paper}
@@
 
-Reading the data from files is a huge overhead, thus, speedup of push-based system in general is pretty significant over pull-based system in our given case. 
-It's been demonstrated in the previous paper with different amount of atoms, frames, and workloads \cite{mainPaper}.
-Since we have to load the data every time there is a different query versus push-based system loads a chunk of data once, it is very noticable how the framework \cite{mainPaper} contributes to efficiency of analysis of simulation data.
-You can observe some estimation samples in Figure \ref{fg:sample_estimation_old_paper}
+Reading the data from files is a huge overhead, thus, speedup of push-based system in general is significant over pull-based system in our given case. 
+It's been demonstrated in the previous paper with different amount of atoms, frames, and workloads.
+Since we have to load the data every time there is a different query versus push-based system loads a chunk of data once, it is quite noticeable how the framework contributes to efficiency of analysis of simulation data.
+You can observe some estimation samples in Figure \ref{fg:sample_estimation_old_paper}.
+Even though the memory loading and pushing steps are the key features of the proposed design, nevertheless, having general knowledge about the network of queries, in this paper, we will try to focus on optimizing the actual execution of them.
 
-On the other hand, some of the queries may be very slow due to their nature on a sequential type of computation,  especially, 2-body functions.
-Since the nature of the data that comes with simulation is basically physical properties of atoms, thus, there is a lot of computation that involves independent primitive mathematical operations, which makes it a perfect problem for parallelism.
+Some of the queries may be very slow due to their nature on a sequential type of computation,  especially, 2-body functions.
+Since the nature of the data that comes with simulation is basically physical properties of atoms, there is a lot of computation that involves independent primitive mathematical operations, which makes it a perfect problem for parallelism.
 
-Considering the fact that the running time of the most queries used
-in the scientific world to analyze the MS is minuscule in comparison
-to the loading time of the MS data, it is safe to say that the
-scientific world would benefit from a system that can remedy the
-aforementioned problem. Our idea is to build a system that would
-load the MS data from hard disk in the main memory only once and
-then execute as many queries as needed/wanted on that data, without
-the need of reloading the same data over and over again. Considering
-the volume of data in a single MS frame as well as the number of
-frames produced during a single simulation, this type of system can
-save a lot of time for MS analysis. The system would contain many,
-most often used queries that can also be run as user's input. These
-queries would be pre-programmed as separate modules in our system,
-but they would be able to take certain attributes, like data
-selection for instance, as user's input. Once the system is run, all
-(or a selection) of these queries would be executed against the MS
-data or certain selection. The system would act as a type of
-push-based system, essentially pushing the loaded data to all the
-queries. Once we have the data in the memory, the reading can loop
-as many times as the queries need it. This is useful for some
-non-streaming functions/queries that cannot be computed in a single
-run of the data. Such one query is the SDH. But we have also come up
-with a way to resolve this in a way, by creating a quad-tree based
-database for this query. This quad-tree (we called Density Map (DM))
-is being populated while the data is being loaded into main memory.
-After the DM is populated, we can compute the SDH with much faster
-algorithm than the brute-force method used everywhere else.
+Although the original idea of push-based system for molecular simulation data analysis was focused on optimizing throughput and usage of loaded data, there were some extra approaches specifically for 2-body functions.
+For example, Density Map for Spacial Distance Histogram was used in order to avoid additional memory allocation and latency reduction with proven error bounds.
+SDH is a quite intensive and computational problem, especially with increasing amount of atoms.
 
-Another improvement that our system provides is the following: by
-thorough observation and analysis we have discovered that many of
-the quantities (functions/queries) that are being computed in order
-to analyze the MS system share some of the same basic parts. In
-order to take advantage of that, our systems precomputes these basic
-parts and has them handy (stored in memory) whenever one of the more
-complex functions needs them. So, for instance, when the total mass
-of the system is needed, our system just pulls the precomputed value
-out and serves the query that needs it immediately. And if more than
-one query uses the same basic part, that translates into time
-saving, however minuscule it might be.
+We believe that having this nature of computational problems, the proposed GPU improved version of push-based system will significantly change in terms of performance by incorporating parallelism with CUDA.
 
-\subsection{Contribution and roadmap of the paper} We have designed
-and implemented a push-based system that can be used as a tool for
-analysis of MS systems. We have tested the system on real MS
-datasets. The experimental results evidently show the superiority of
-the proposed system over one of the most widely used MS analysis
-system across many scientific fields, i.e., GROMACS. The efficiency
-improvement ranges anywhere from $3$ to $100$ times improvement (of
-the overall running time) per set of selected queries run on a
-different selection of the MS data. If we take in account the number
-of such queries that the scientist run every day, we believe our
-system would be of great benefit to MS community. In addition to the
-higher efficiency that our system achieves, we also believe that our
-idea will initiate a new breed of database related, push-based
-systems that could be used to analyze the MS systems in a more
-efficient way. Having a push-based like system, we were also able to
-improve the computation of the more complex quantities. Namely, the
-computation of some such complex quantities involves computation of
-much less complex functions (sub-parts), and many such sub-parts are
-being shared among some of the complex quantities. So, by computing
-these more basic, sub-quantities in advance, we were able to achieve
-however minuscule time gain when computing the more complex queries.
+\subsection{Contribution and roadmap of the paper} 
+The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis even faster now.
+Taking an advantage of GPU devices nature and incorporating parallelism and streaming for different types of queries, we have come up with a good speed up over sequential processing, which already had a good speed up in terms of performance in comparison with original pull-based approach.
+Since in this paper we don't expect certain form of data, we have come up with a tool that generates mock data for any number of atoms and frames, which can be used for performance tests that were done consequently, too.
+This mock data has exactly the same information that would come with Gromacs simulation files, and it has exactly the same format as in the previous paper \cite{mainPaper} right before loading it to memory.
+This improved version has been developed on Amazon Elastic GPU nodes that consequently can be replicated and scaled up becoming more of a streaming processing engine, which can be very effective on costs, since it is always easy to scale up the nodes, shut them down and spin them back up as well as on a simple desktop computer with a GPU device.
 

### 2016-01-27 paper/paperSource.tex
paper changes

@@
 This mock data has exactly the same information that would come with Gromacs simulation files, and it has exactly the same format as in the previous paper \cite{mainPaper} right before loading it to memory.
-This improved version has been developed on Amazon Elastic GPU nodes that consequently can be replicated and scaled up becoming more of a streaming processing engine, which can be very effective on costs, since it is always easy to scale up the nodes, shut them down and spin them back up as well as on a simple desktop computer with a GPU device.
+This improved version has been developed on Amazon Elastic GPU \cite{amazonGPU} nodes that consequently can be replicated and scaled up becoming more of a streaming distributed processing engine (which can be very effective on costs, since it is always easy to scale them up, shut them down and spin them back up), as well as on a simple desktop computer with a GPU device.
+In general, major contribution of this paper is the framework of push-based system with an incorporated parallelism based on CUDA programming language.
+Besides it, there was developed a tool which generates mockup data of the same format as of the real MS simulations and represent a python script with arguments given as number of frames and atoms.
 
-The major technical contributions presented here are:
-\begin{itemize}
-\item Design the network (tree like) of the most commonly used queries in MS (physics);
-\item Build the system: design and build the modules, representing the quantities to be
-computed in an efficient manner, following the already built
-network;
-\item Develop a scientific simulation database benchmark that can be used for evaluating similar systems and products.
-\end{itemize}
-
-The remainder of this paper is arranged as follows: in Section
-\ref{sc:relatedwork} we give a survey of the systems used in the
-field of MS systems analysis. We continue the paper with Section
-\ref{sc:querynetwork} in which we show the design of the query
-network build from the most widely used quantities in MS system
-analysis. In Section \ref{sc:system} we describe our push-based
-system for MS data analysis. Then, in Section \ref{sc:experiments}
-we present the benchmark designed to test our system as well as the
-results attained through comprehensive experiments run on real MS
-generated data. At the end, we conclude this paper with Section
-\ref{sc:conclusion} in which we give an overview of our possible
-future endeavors in the field of MS data analysis.
+The structure of the framework which will be described later is developed in such a way that one can easily add new queries based on their complexity or modify existing ones adding appropriate selections and filters.
 

### 2016-02-01 paper/paperSource.tex
paper changes

@@
 %\smartqed  % flush right qed marks, e.g. at end of proof
-
 %----------------------------------------------------------
@@
 \begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4-eps-converted-to.pdf} }
+ \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
  \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainPaper}}
@@
 
+\vspace{3mm}
 \section{Related Work}\label{sc:relatedwork}
-The idea of data streaming has been broadly used in many fields. The
-main usage however is aimed at processing live data generated
-online. There is an ocean of references for data stream management,
-but we believe the presentation in~\cite{DataStreamManagement}
-encapsulates the majority of the ideas, problems and solutions. In
-the past decade, however, the database community started to follow
-the data stream idea to process stored data. Processes can take
-advantage of the streaming data at any time the data is being pushed
-through the system. This gives rise to push-based type design for
-data management systems. The idea of such push-based design was
-previously considered in projects such as DataPath~\cite{DataPath},
-Volcano~\cite{Volcano} and QPipe~\cite{Qpipe} among others. These
-works show ideas in which the data-driven dataflow is compared to
-the demand-driven dataflow showing the need for the later. They also
-talk about maximizing the data and work sharing among queries at
-runtime. Essentially, we incorporate such ideas in our design of the
-push-based system.
+\vspace{3mm}
 
-On the other side, the scientific community has steadily progressed
-from processing massive data files towards employing database
-systems for the storage, acquisition, and analysis of large-scale
-scientific data~\cite{SDSS_SIGMOD02,QBISM_ICDE94}. The widely used
-and popular relational database systems are conventionally designed
-and optimized to better manage the data produced by the business
-type applications. But such conventional database systems (DBMS) are
-not well equipped to deal with the type and quantity of scientific
-data, such as the data produced by the molecular simulations. In the
-recent past, the DBMS community has made some attempts into the
-design and construction of database systems optimized for handling
-scientific data. Such examples include the BDBMS
-project~\cite{BDBMS_CIDR07} that deals with annotation and
-provenance of the sequence data in biosciences. and the PeriScope
-project~\cite{PeriScope} is designed to efficiently handle
-declarative queries against bio sequences. On top of the
-aforementioned examples, there are also ideas for new DBMS
-frameworks aimed at the management of scientific
-data~\cite{NewArch-VLDB07,ROI-CIDR07,SciDB-SIGMOD10}. One of those
-systems is the SciDB~\cite{SciDB-SIGMOD10,SciDB} and it is closest
-to the idea presented in this paper. SciDB is data management and
-analytics system that is primarily used is in application domains
-involving very big scale array data. This system, like the one
-presented in this paper is designed around a multi-dimensional array
-data-model and it uses arrays to store the data. SciDB stores
-petabytes of data on a number of machines and runs its queries on
-those machines. It is made for high performance, high-availability,
-fault tolerance, and scalability. However, to the best of our
-knowledge, it too follows the pull based design where its queries
-demand the data they need. As seen earlier in this paper, this type
-of design can impose I/O overhead and decrease the data throughput
-when doing the analysis. Aside the mentioned issue, the design and
-build of such DBMS optimized for scientific data management come
-with a additional challenges. Such challenges as well as their
-probable resolution are outlined in~\cite{Sci_SIGMODR05}. Recently,
-there have been some efforts aimed at designing and building MS data
-managements systems on top of relational databases. Such efforts are
-presented through projects like BioSimGrid~\cite{BioSimGrid} and
-SimDB~\cite{SimDB} that were developed especially for molecular
-simulations. However, to the best of our knowledge, such systems
-still lack the efficiency needed for MS data management as well as
-efficient query processing strategies.
+Push-based system for MS data analysis is primarily proposing an idea of pushing chunks of data through a network of queries.
+Essentially, this has also been proposed, as mentioned already, by other frameworks \cite{DataPath,Volcano,Qpipe}.
+In other words, the major point is to use common data loaded into memory for execution of concurrent queries, which is supposed to eliminate I/O overhead predominantly.
+The architecture and explanation of DSMS (Data Stream Management System) versus DBMS is well explained in one of the lectures of Morgan and Claypool \cite{DataStreamManagement}.
+It is very important to capture the differences and specialties of DSMS in network architecture, stream models and windows, scheduling, load balancing, approximation, data expiration, etc.
 
-Generally, the data produced in the process of molecular simulation
-is being stored in large, plain files with no structure whatsoever.
-Queries, which are implemented in a stand alone programs within
-simulation/analysis systems, are executed onto such files producing
-the quantities that scientist use to analyze the molecular system.
-Such simulation and/or analysis packages include: Gromacs, VMD,
-MDAnalysis, Wordom, MD-TRACKS, SimulaidOne, Charmm among the others.
-But to the best of our knowledge, all of these systems work on a
-similar basis: they take a user defined query and execute it onto
-the MS simulated data. In order for the query to be executed, the
-data has to be loaded into the main memory. Then the result is
-either produced onto the display or written to a file. When the next
-user query comes, the system again loads significant part of the
-dataset into the main memory and executes the query. We believe that
-there is a room for improvement of such systems, given the fact that
-many of the user defined queries executed during system's analysis
-are fairly static. In other words, there is a number of queries that
-a user would always want to execute on a given simulation data.
-Furthermore, the selections of MS data onto which such queries might
-be executed, are also fairly constant (i.e., oftentimes the user
-selects the same group of atoms (e.g., all hydrogen atoms) to
-calculate given quantity, like center of mass for instance). So, by
-precoding many such queries and running them automatically once the
-system has loaded the data into memory, we believe we can save a lot
-of time that otherwise would have been spent in loading the same
-data into main memory anytime a query is executed. On top of the
-automated query execution, our system can take user's query as input
-as well. With this, we believe our system is an improvement over the
-MS analysis systems that are used today.
+Since dealing with data is very delicate in this kind of engines, there are couple of features and requirements that this architecture is applicable to, in order to avoid inappropriate understanding or application of this approach in systems that were covered in the lectures\cite{DataStreamManagement}.
+The data is volatile and not persistent anymore because of obvious sizes of data.
+Since we are pulling the data as little times as possible taking into consideration that MS data is only appended and not modified, the access is sequential rather than random.
+There are concepts of framing and windowing, and memory is limited in streaming engine, while in DBMS secondary storage is considered to be unlimited.
+Obviously, because of all of these features or requirements to data, the streaming engine frameworks have some limitations like going back into the history of data, since it is volatile.
+In fact, this has also been tried by Borealis Streaming Engine \cite{borealis}.
+On top of the fact that it is a streaming engine, as it was built based on Medusa\cite{medusa} and Aurora\cite{aurora}, it proposes distributed processing at this point having some features to handle errors and look back in the history.
+In our case, it is primarily a single data source, since it is based on the simulation, though in the future the simulation software systems might introduce distributed computing, too.
+It is also important to mention such projects as PeriScope\cite{PeriScope}, SciDB\cite{SciDB}, and other\cite{SDSS_SIGMOD02,QBISM_ICDE94}, since they target scientific data, which is a lot more different, than usual industry customer oriented data.
+
+This version of Push-based System of MS data analysis is primarily about taking an advantage of GPU processing of the data.
+Hadoop is a well-known and well-used framework by many companies nowadays\cite{hadoop}.
+Its purpose is an ability to store and serve large scale data across clusters in a timely fashion.
+Since usage of GPU of massive data (not only related to graphics processing) is a relatively new concept, there is also an improved version of Hadoop MapReduce Framework with GPU\cite{hadoopgpu}.
+Having an ability to scale it up in a distributed computing environment is probably one of the best approaches, but again it might be a bigger overhead in terms of costs, since keeping all nodes up and still being able to solve specific problem sets with hardware stack of a lower capacity raise a will to explore more.
+Having a specific software for simulation like Gromacs, apparently there is also a possibility of taking an advantage of GPUs.
+For example, Gromacs allows to optimize simulation and query tools by adding configurations based on GPU device located on the processing computer.
+Unfortunately, even with this feature, Gromacs doesn't follow push-based approach which means that improvement with GPU with pull data per query makes it negligible.
+
+Gromacs generate output files with physical properties description for each atom in each frame.
+Modern CPUs do great job in terms of caching and computing, but obviously in this particular problem massive monolithic computation is needed, thus, GPU devices are perfect candidates for such a problem.
+To be more exact, for example, a single kernel in a GPU device could be dedicated to a single atom (this relationship might change depending on the query, but this is to understand the scale and relationship of multiprocessors), and since we have scientific data of MS, the data is appended, which means that we will only move forward and not take an advantage of caching on chunks high level (though we might take an advantage of GPU caching features particularly in processing phase) having multiple workloads per chunk.
+Just to make it more clear, workloads in this particular case represent transactions based on number of clients.
+For example, if we were to serve it all in a streaming fashion, 100 clients might demand concurrent queries with different slight selection properties and expect their results, each workload might be dedicated to each client if not aggregated.
+Thus, the performance of GPU is certainly increasing based on workload too, which might benefit considerably.
 

### 2016-02-01 paper/paperSource.tex
automatic triggered changes

@@
 The architecture and explanation of DSMS (Data Stream Management System) versus DBMS is well explained in one of the lectures of Morgan and Claypool \cite{DataStreamManagement}.
-It is very important to capture the differences and specialties of DSMS in network architecture, stream models and windows, scheduling, load balancing, approximation, data expiration, etc.
+It is very important to capture the differences and specialties of DSMS in network architecture, stream models and windows, scheduling, load balancing, approximation, data expiration, etc, because of the fact that data flow design is the key that required all of those changes to common designs.
 
@@
 Gromacs generate output files with physical properties description for each atom in each frame.
-Modern CPUs do great job in terms of caching and computing, but obviously in this particular problem massive monolithic computation is needed, thus, GPU devices are perfect candidates for such a problem.
+Modern CPUs do great job in terms of caching and computing, but obviously in this particular problem massive monolithic computation is needed, thus, GPU devices are perfect candidates for such a problem following SIMD type of computation.
 To be more exact, for example, a single kernel in a GPU device could be dedicated to a single atom (this relationship might change depending on the query, but this is to understand the scale and relationship of multiprocessors), and since we have scientific data of MS, the data is appended, which means that we will only move forward and not take an advantage of caching on chunks high level (though we might take an advantage of GPU caching features particularly in processing phase) having multiple workloads per chunk.

### 2016-02-07 paper/paperSource.tex
automatic triggered changes

@@
 \usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
+\usepackage[justification=centering]{caption}
 %\usepackage{fixltx2e}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}
+\centering
+\centerline{\includegraphics[width=0.5\columnwidth]{images/kernelatom.eps}}
+\caption{ Example of atom mapping with kernels}
+ \label{fg:kernelatom}
+
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Contribution and roadmap of the paper} 
@@
 
+\vspace{3mm}
 \section{Network of Queries}\label{sc:querynetwork}
+\vspace{3mm}
 
-\textbf{\emph{MS Queries}}. In order to study some important
-statistical features of an MS system, scientist need to ``extract''
-various statistical quantities out of the data produced by the
-simulation. To achieve this, queries are executed against the data.
-Most of the queries used in the analysis of MS systems are
-analytical in nature. Essentially, these analytical queries are
-mathematical functions that translate a selection of atoms (atoms'
-measurements) to a scalar, vector, a matrix, or a data
-cube~\cite{SimDB}. Once the simulation is done, the analysis carried
-out will depend on the structure being studied as well as the
-features of the system that need exploring. In other words, not all
-system's quantities need to be computed every time the system is
-being analyzed. Some of the more popular queries, including density
-(atom counts), first-order statistics (mean), second-order
-statistics (variance), and histograms among others, can be seen in
-Table~\ref{tb:queries}. The queries shown in this table are the ones
-that we have also incorporated in our system. Just to clarify some
-of the notation in Table~\ref{tb:queries}: we assume that the MS
-system comprises of $n$ particles and $r_i$, $m_i$, $c_i$ and $q_i$
-denote coordinates (vector form), mass, charge, and number of
-electrons of a particle $i$, respectively.
+Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
+Often times, the queries that need to be run over the data look like selection and some kind of accumulation, if it is not more complicated.
+To be more exact, in this section we will try to summarize common set of queries we will be improving.
+
+In previous work, there has already been developed a network of queries widely used by scientists for MS systems.
+In Table~\ref{tb:queries} you may observe the common set of queries that has been developed.
+Basically, these are the queries that need to be improved with our new approach of parallelism.
+Just to mention, $n$, $r_i$, $m_i$, $c_i$ and $q_i$ denote number of particles, coordinates (vector form), mass, charge, and number of electrons of a particle $i$.
 
@@
 \end{tabular}
-\caption[Table caption text]{Popular analytical queries in MS }
+\caption[Table caption text]{Popular analytical queries in MS\cite{mainPaper} }
 \label{tb:queries}
@@
 
-There are two types of queries/functions among the ones used to
-analyze an MS system. The first type are one-body functions. Such
-functions usually are algebraic functions~\cite{SDSS_SIGMOD02} and
-only involve quantities(attributes) from a single atom at any given
-time in the process of computation. Each atom (atom's attributes) is
-being processed a constant number of times, thus the total running
-time of such functions/queries is $O(n)$. This type of functions is
-very suitable for the idea of push-based system, or an online system
-in which the data is being read once and acted upon. In other words,
-in a single run of the incoming data, all such queries will produce
-useful final results. Except the SDH and the RDF (i.e., the last two
-in the table), all other queries shown in Table~\ref{tb:queries}
-fall into this category of one-body functions. Most of these
-functions are defined on a single frame of the MS data. Only the
-autocorrelation functions are defined on two distinct frames.
+In general, with the given queries we have divided them into two categories: one body functions and multiple body functions.
+The first ones apparently have linear complexity of $O(n)$, while the other ones have bigger complexity.
 
-The second type of functions are multi-body functions and are
-holistic in nature. The computation of such functions involve more
-than one atom's attributes and cannot produce final result in a
-single run of the MS data (i.e., if traditional methods are used for
-their computation). Such queries include the Radial Distribution
-Function (RDF)~\cite{bamdad06, Frenkel:api01, AstroAnalysis02} as
-well as some quantities associated with chemical
-shifts~\cite{ChemShift_BCB98}. Generally, such functions are
-computed through histograms. For instance, the RDF is obtained from
-a histogram of all pairwise atom distances (this is the Spatial
-Distance Histogram or SDH). The traditional, straightforward (often
-the brute-force) way of computing these holistic functions is a very
-time consuming process. On top of that, these methods cannot produce
-the final result in a single run of the MS data, making such
-functions unsuitable for our idea of a push-based system. However,
-in our previous work, we have designed a data structure together
-with an algorithm that opens up the possibility for such queries to
-be executed in a push-based type environment. Further details on
-this are given in Section~\ref{sc:system}.
+\emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
 
-Fig.~\ref{fg:querynetwork} represents an idea to show how such query
-processing system can be improved. The idea behind it is that some
-of the queries share same sub-routines. Having all the queries made
-as separate modules, this sub-routines can be computed once the data
-is being pushed through the system and then be used anytime a more
-complex query needs it. Having in mind the amount of data in a
-single frame that the queries (sub-routines as well) need to go
-through, and the fact that in a single MS there are thousands of
-frames, we believe this can be immense improvement in terms of total
-running time.
+\emph{\textbf{Multiple body functions}} such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
+This is generally combination of two across N data: 	$C\binom{N}{2}$.
+These are the most expensive queries, thus, it is important to focus on their implementation taking advantage of GPU nature and its processing power.
+SDH is a very expensive computation, and we have used some of the existing work of ours\cite{sdhgpu} to incorporate in this design.
 
-\begin{figure*}
- \centerline{ \includegraphics[width=1\textwidth]{images/msstructure.eps} }
- \caption{MS Modules Structure}
- \label{fg:querynetwork}
-\end{figure*}
+The might be some dependencies and preliminary computation possibilities.
+For example, some queries need total mass, and we could compute it before we push it further, but this can be easily done by user based on the complexity and dependency.
+It is explained in more details in further sections.
 
-\section{Building the system}\label{sc:system}
+\vspace{3mm}
+\section{The system design}\label{sc:system}
+\vspace{3mm}
 
-\subsection{MS Data retrieval and in memory organization}
+The proposed design has a pretty straightforward structure which is going to be explained in this section.
+Shortly, the system starts with original input files (or data) and gets consumed in a sequential manner by simple disk IO framework into random access memory by chunks, then, chunk gets uploaded to GPU device and processed, at some point, the results that need to be shown to the user are transported from GPU (where they initially get accumulated) to the main memory, and then might get written as well to some other output files.
 
-A typical MS system generates and stores the data in a number of
-trajectory files, usually including multiple frames (snapshots of
-the simulated system taken at certain time intervals). Such MS
-generated data oftentimes goes through a simple lossless compression
-and, depending on the simulation software it may be stored in a
-binary format. Such trajectory file format is one of most often used
-MS file format (e.g., GROMACS, PDB). But such format is
-unrecognizable to our system. So our system has to do three things
-before it starts executing the queries: 1) Read the MS data from a
-trajectory file, 2) Translate the MS data to a form recognizable to
-our system, and 3) Load the data to memory.
+\begin{figure}
+ \centerline{ \includegraphics[width=0.65\columnwidth]{images/memory-organization.eps} }
+ \caption{Data flow}
+ \label{fg:data-flow}
+\end{figure}
+
+\subsection{Data flow}
+Since the size of the data in the problem is huge, organization of the data flow is crucial in our case.
+As mentioned before, original data is retrieved from simulation software output files and in our case it is Gromacs Software's formatted trajectory files.
+Tool for decompressing those files to a flat format has been developed in the previous work\cite{mainPaper}.
 

### 2016-02-08 paper/paperSource.tex
automatic triggered changes

@@
 The proposed design has a pretty straightforward structure which is going to be explained in this section.
-Shortly, the system starts with original input files (or data) and gets consumed in a sequential manner by simple disk IO framework into random access memory by chunks, then, chunk gets uploaded to GPU device and processed, at some point, the results that need to be shown to the user are transported from GPU (where they initially get accumulated) to the main memory, and then might get written as well to some other output files.
+Shortly, the system starts with original input files (or data) and gets consumed in a sequential manner by simple disk IO framework into random access memory by chunks, then, chunk gets uploaded to GPU device and processed.
+At some point, the results that need to be shown to the user are transported from global memory of GPU (where they initially get accumulated) to the main memory, and then might get written as well to an output file.
+General data flow map is represented in Figure  \ref{fg:data-flow}.
 
@@
  \centerline{ \includegraphics[width=0.65\columnwidth]{images/memory-organization.eps} }
- \caption{Data flow}
+ \caption{Data flow in the improved design of push-based system for MS analysis}
  \label{fg:data-flow}
@@
 
-\subsection{Data flow}
-Since the size of the data in the problem is huge, organization of the data flow is crucial in our case.
-As mentioned before, original data is retrieved from simulation software output files and in our case it is Gromacs Software's formatted trajectory files.
-Tool for decompressing those files to a flat format has been developed in the previous work\cite{mainPaper}.
+\subsection{Data input}
+Since the size of the data in the problem is huge, organization of the data flow is crucial.
+As mentioned before, original data is retrieved from simulation software output files and in our case it is Gromacs Software's trajectory files\footnote{The file types, their structure, and their decompression has been well explained in the previous paper in section "Data orgranization in main memory"}.
+For a simulation, there can be several files of different format and usually they are also compressed without any loss of information.
+Tool for decompressing those files to a flat format for each frame and atom has been developed in the previous work\cite{mainPaper}.
 
-\emph{\textbf{Data read - transform.}} As mentioned above, in order
-for our system to be able to read the MS generated data, the data
-needs to be transformed. The reason for this is following: the MS
-data is stored in multiple files and possibly in different formats
-as well. One such file holds the global data (identifying the system
-and the simulation). Another file holds each frame's data. The frame
-data contain general information about the frame, but the main part
-is a sequential list of each atom's info, including atom's mass,
-position, charge, number of electrons, velocities, forces, etc.
-Another file (topology file) holds the molecule/residue info,
-essentially identifying what atom belongs to which molecule.
+In order to experiment with different input cases, the input files have been generated with a tool that we have built in Python language.
+Thus, we are able to generate mock data files with any number of frames and atoms, and the values are random and can be adjusted with needs of the experiment (for example, for cases when we want to analyze bucket widths in a limited space, the coordinates are generated in range of $0 - 1$, etc.).
+As it has already been mentioned earlier, we don't expect the data to be distributed with a certain behavior, the importance of having this kind of tool to generate mock data is having a variety of random data sets that could reveal any issues related to values provided to the program, on top of that, having the actual file written and then being read has been a part of the experiments to have more realistic use case.
+In the future, the same script for generating the mock data could be used for regression and performance tests in a distributed implementation of this design as a single source.
 
-So, in order to extract the data from these files, we have created a
-sort of "extractor/transformer" of the attributes needed for the
-execution of the queries in our system. This transformer,
-essentially, is a separate piece of code that does three things: 1)
-it reads the MS generated data that the MS system stores in one of
-the often used MS file format (e.g., GROMACS); 2) it translates the
-data into a format that our system can read (taking only the
-information our system needs); and 3) it stores the data in a file
-that has basic structure to it. So, in the end, the data transformer
-produces a data file that our system takes as input. This code
-serves as a connection between an MS system (e.g., GROMACS) and our
-system. With this, our system can essentially be used as an add-on
-to GROMACS or other simulation systems and help improve the
-efficiency of the data analysis.
+\subsection{Data flow architecture}
 
-\textbf{\emph{Data organization in main memory.}}\footnote{This
-paragraph talks only about the data organization used by one-body
-queries. For two-body queries (e.g., SDH), the data organization is
-discussed later on.} Once the data is in a format our system can
-read, the data is being loaded into the main memory one frame at a
-time. The organization of the in-memory particle's data is in the
-form of a simple, two dimensional array where a single row
-represents an atom in the system with all its attributes (e.g.,
-coordinates, mass, charge, residue info, etc.). We also keep (in a
-one dimensional array) crucial system's information for each frame,
-like temperature, energy, pressure, etc. We have used such
-structures because they are very suitable for simulating a
-push-based type of system: a simple sequential read of the array
-gives that on-line type of data stream. So, as the system reads the
-array, it pushes the data onto the query-modules. As mentioned
-earlier, the one-body (algebraic) queries will produce a final
-result at the end of the first sequential read. However, the
-two-body (holistic) functions, like SDH and RDF, cannot do this in a
-single read of the data.\footnote{However, we have previously
-deigned and created a data structure and an algorithm that can take
-the advantage of a single data read and produce final results for
-SDH computation~\cite{ytu:icde09,EDBT12}. We have incorporate this
-into our system presented in this paper.} Therefore, each frame's
-data array can be continuously read (in a loop manner) as many times
-as a query needs it.
+MS data roughly can be represented as a list of atoms with physical properties aggregated as frames.
+Thus, we consider a single frame as a chunk which will be loaded into the memory\footnote{Memory in this design might be interpreted differently, since we have a separate device memory on GPUs, but it will be explained better later} and pushed through our network of query.
+In this case, it is very convenient, because there are almost no dependencies between frames, rather than list of atoms in the same frame, which potentially again could create additional disk I/O overhead.
+
+Having defined the chunk, we load it from flat format file to RAM.
+A single atom data structure has all the attributes such as charge, mass, velocity, type of atom, coordinates, and so on.
+In other words, we have an array of atoms, and most of the attributes (in case of data structure, it is a class field), are simply primitive floats or doubles depending on precision requirements.
+Next, the same chunk of data with the same structure is loaded onto GPU global memory.
+GPU devices, in general, have about 4-5 types of memory, and global memory is the biggest and slowest.
+Consequently, depending on algorithm or query we load parts of data chunk to another types of memory called shared memory.
+It depends on the actual query, because the same part of chunk can be loaded and unloaded in regards with shared memory several times, especially with SDH.
+
 \subsection{Query modules}
-
 As mentioned earlier, there are two types of functions/queries used

### 2016-02-09 paper/paperSource.tex
automatic triggered changes

@@
 
-\subsection{Query modules}
-As mentioned earlier, there are two types of functions/queries used
-for analysis of MS systems: algebraic or one-body, and holistic or
-two-body queries (these in general can be multi-body, but in this
-paper we only deal with a two-body functions).
+\subsection{Network of queries organization}
 
-\subsubsection{One-body queries} Most of the query modules in
-Table~\ref{tb:queries} (except the SDH and RDF) are not that
-involved, only containing computations of fairly simple, one-body
-functions. These queries were coded as separate modules in our
+Since we assume that our framework should be easily used by a scientist doing MS analysis, it is important to keep queries module as a separate piece of code to make it easily extendable.
+In order to understand how queries module is organized and the reason for it, we need to understand how CUDA and GPU work.
+
+\emph{\textbf{GPU execution}}.
+From hardware standpoint, it is important to understand that every GPU device consists of SM (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
+Having this in mind, since this is a pure parallel processing power, the goal is to increase hardware throughput and keep highest utilization to exploit every register given.
+CUDA is an amazing tool that tries to help developers map this hardware capabilities with the actual application layer.
+In CUDA we have grids, blocks, and threads, where grid is 3 dimensional unit where each cell is a block, and the same relationship applies between blocks and threads\footnote{This explanation doesn't include memory setup in GPU devices, but it is covered in further sections}.
+Originally, this mapping was perfect for solving image processing problems or other graphics computation, since dimensions and such division (e.g.
+by pixels to threads) is very inherent.
+As a language feature, in CUDA we have to define such functions called 'kernel' function.
+Essentially, it is the function where we are supposed to map every execution thread with data and perform the actual execution, this approach is widely known as SIMD\footnote{SIMD - Single Instruction Multiple Data}.
+As a convenience we have separated one-body functions with multiple-body functions into separate kernels.
+
+\emph{\textbf{One-body queries kernel}}.
+Most of the query modules in Table~\ref{tb:queries} (except the SDH and RDF) are not that involved, only containing computations of fairly simple, one-body functions.
+These queries were coded as separate modules in our
 system. Each of these modules take few attributes as input (e.g.,

### 2016-02-10 paper/paperSource.tex
automatic triggered changes

@@
 \emph{\textbf{One-body queries kernel}}.
-Most of the query modules in Table~\ref{tb:queries} (except the SDH and RDF) are not that involved, only containing computations of fairly simple, one-body functions.
-These queries were coded as separate modules in our
-system. Each of these modules take few attributes as input (e.g.,
-atom selection, frames selection(for the autocorrelation functions),
-number of atoms, etc.). The system pushes the data as it becomes
-available onto these modules. The queries are being executed on the
-selection and are put in a ``ready'' mode, awaiting the next frame's
-data. First, the more basic queries, like total mass, are being
-computed. The results of such queries are temporary stored (in main
-memory) and are available for use anytime a more complex query needs
-them.
+Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
+Figure \ref{fg:reduction} nicely represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel.
+Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes.
+We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of spacial prefetch, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage.
+In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries.
+We fetch all the atoms from GPU global memory, since we use references to atoms once.
+In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage.
 
-\subsubsection{Two-body queries} In general, queries involving two-body functions are a
+\emph{\textbf{Two-body queries kernel}}.
+In general, queries involving two-body functions are a
 bit more complex and cannot provide the final result in a single
@@
 \begin{figure}
- \centerline{ \includegraphics[width=0.65\columnwidth]{images/resolve.eps} }
- \caption{Computing minimum (i.e., length of solid lines) and
- maximum distance (i.e., length of dashed lines) range between two
-cells}
- \label{fg:resolve}
+ \centerline{ \includegraphics[width=0.65\columnwidth]{images/reduction} }
+ \caption{Array reduction using coalesced memory access\cite{gpureduction}}
+ \label{fg:reduction}
 \end{figure}

### 2016-02-15 paper/paperSource.tex
automatic triggered changes

@@
 by pixels to threads) is very inherent.
-As a language feature, in CUDA we have to define such functions called 'kernel' function.
+As a language feature, in CUDA we have to define such a function called 'kernel' function.
 Essentially, it is the function where we are supposed to map every execution thread with data and perform the actual execution, this approach is widely known as SIMD\footnote{SIMD - Single Instruction Multiple Data}.
@@
 Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes.
-We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of spacial prefetch, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage.
-In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries.
+We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of coalesced memory access, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage of it.
+In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries for a scientist to program.
 We fetch all the atoms from GPU global memory, since we use references to atoms once.
 In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage.
+Finally, memory transportation and data consumption are implemented for user, thus, he only needs to extend the queries module which consequently can be turned into plugins model.
 
 \emph{\textbf{Two-body queries kernel}}.
-In general, queries involving two-body functions are a
-bit more complex and cannot provide the final result in a single
-data read if a straightforward method is used for their computation.
-However, in the proposed system we have incorporated a data
-structure and an algorithm for the SDH (also RDF) from our previous
-work that is suitable for push-based type of system. In this
-subsection we give a brief description of the data structure (DM)
-and the algorithm (DM-SDH) designed in~\cite{ytu:icde09, EDBT12} and
-implemented in the system proposed in this paper. For more detailed
-information, please refer to our previously published work on this
-topic~\cite{ytu:icde09, EDBT12}
-
-\emph{\textbf{The data structure.}} The simulation data space is
-represented by a conceptual data structure we named Density Map
-(DM). The density map splits the simulation space into a grid of
-equal size regions (or cells). The cells are cubes in 3D and squares
-in 2D\footnote{In this paper, we focus only on the 2D data to
-elaborate and illustrate the proposed ideas.} Resolution of a
-density map is the reciprocal of the cell size in that density map.
-In order to generate higher resolution density map, we split each
-cell of the current resolution's grid into four smaller cells of
-equal size. This design allows us to use a region
-quad-tree~\cite{orenstein:ipl82} to organize density maps of the
-same data but with different resolutions. So, essentially, a node in
-the quad-tree represents a single cell from the DM. Therefore, a
-density map of a certain resolution basically is the set of all
-nodes of one level of the tree. Each of the tree nodes records the
-cell's location in the density map (coordinates of corner points) as
-well as the number of particles in each cell. We name the afore
-describe tree the Density-Map tree (DM-tree).
+SDH (or RDF) is the algorithm that requires special treatment in our design.
+The point, as mentioned earlier, is that this function requires more than one iteration of the given set of atoms, but rather all combinations of pairs in a given frame, thus, it is not linear complexity.
+An implementation of this type of query might be very non-intuitive because of the nature of GPUs.
+Another issue is that since to take an advantage on this type of computation, before writing an implementation for it, a user must have a good understanding of GPU architecture which makes the system less user-friendly.
+At this point, it is more important to achieve good performance improvement results and then take care of ease of system extensions.
+As we already mentioned, GPU threads execute in parallel and consume certain amount of work followed by best try of all hardware capacities utilization, but, unfortunately, it is also important to take care of equal workload for every thread independently, so, that utilization is kept high.
+For this reason, we've come up with a solution that should increase performance of the computation which will be described later.
 
@@
  \centerline{ \includegraphics[width=0.65\columnwidth]{images/reduction} }
- \caption{Array reduction using coalesced memory access\cite{gpureduction}}
+ \caption{Array reduction example\cite{gpureduction}}
  \label{fg:reduction}
@@
 
-\emph{\textbf{The algorithm.}} The essential part of the DM-SDH
-algorithm is a procedure named {\sc ResolveTwoCells}. The input to
-this procedure are two cells from the density map (e.g., $A$ and $B$
-in Fig.~\ref{fg:resolve}). It computes, in constant time, the
-minimum and maximum distance between the two cells. A pair of cells
-is ~\emph{resolvable} if both the min and max distance between them
-fall into the same SDH bucket $i$. If that is the case, the distance
-count of that bucket is being increase by $n_{A}n_{B}$ ($n_{A}$ and
-$n_{B}$ are the number of particles in cell $A$ and $B$,
-respectively). Otherwise, the cells are non-resolvable and we
-either:
-\begin{itemize}
-\item[(1)] Go to the next density map with higher resolution and resolve all children of $A$
-with those of $B$, or
-\item[(2)] If leaf-level has been reached: compute every distance between particles of $A$ and $B$ and update the
-histogram accordingly.
-\end{itemize}
+\begin{figure}
+ \centerline{ \includegraphics[width=0.65\columnwidth]{images/overallarchitecture} }
+ \caption{Overall system architecture}
+ \label{fg:overallarchitecture}
+\end{figure}
 
-In order to generate the complete SDH, the {\sc ResolveTwoCells}
-procedure is executed for all pairs of cells for a given density map
-$DM_k$ and the algorithm would recursively call the procedure
-(action (1) above) until leaf-level has been reached (action (2)
-above).
+\emph{\textbf{Architecture overview.}}
+This section is to emphasize general architecture of the system and summarize the work flow as well as the data flow.
+In total, we have flat files with ready flat values for each atom aggregated in frames coming into the program as data input using common disk I/O framework of an operating system.
+Then, once we have loaded a frame (which is considered to be a chunk of the entire data set) onto RAM of our machine, we copy over the entire frame onto the GPU device global memory.
+It is important that, at this point, we never write any values to simulation files or any other data related to simulation, thus, we don't have to care about any synchronization of incoming data.
+After this, we have two predefined kernel functions (which were explained before), where first kernel stands for one-body queries, and the second one is for two-body query SDH.
+Just for architecture clearance and possible future improvements, two kernels are executed in two different streams\footnote{In CUDA stream is another abstract layer of organizing execution threads.
+They are not guaranteed to be executed at the same time or after each other, GPU scheduling architecture is completely different from CPU having a warp to be single scheduling unit at maximum of 32 threads at a time} (You can see them on Figure\ref{fg:overallarchitecture}).
+In each kernel we apply certain heuristics to improve memory access using such architecture features of GPU as shared memory and blocking of kernel threads.
 
-On top of the aforementioned DM-SDH algorithm, we have also
-incorporated two approximate SDH algorithms (ADM-SDH), introduced
-and described in \cite{EDBT12}, and \cite{TKDE12}. These approximate
-algorithms are substantially faster than the the brute-force
-algorithm and also than the DM-SDH algorithm as they take advantage
-of some heuristic. For more details on the ADM-SDH algorithms please
-see the aforementioned work.
+\emph{\textbf{GPU grid and block sizes.}} As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
+The approach we came up with is 1 execution thread per atom.
+For CUDA version 2.0 maximum number of threads per block is 1024.
+Blocks are 3 dimensional entities in a 3 dimensional grid.
+The less abstract unit is, the less overhead we have in communication between them, thus, we try to utilize smallest units more.
+For example, for 1000,000 atoms, we would allocate 1024 threads per block as a static hard coded value, since we usually have always more than 1024 atoms in simulations.
+One dimension of grid can take up to 64,000 blocks, thus, if N is bigger than $1024 * 64,000$, than the rest would be allocated to $grid.y$ and $grid.z$ respectively.
+We have not considered numbers bigger than GPU can allocate, which could be explored in future work.
+It is important to understand that even though there are abstract entities of grids and blocks, it is not real hardware capacities, but they can affect scheduling and performance consequently.
+Decision to go with atom to thread execution is because it is pretty convenient to keep high utilization of hardware and still not lose idea if we are about to incorporate more functions, which makes the system more user-friendly, though later on as a plugin we could add data loaded and transformer between RAM and GPU global memory, because there are a lot of custom ways to aggregate data and take an advantage of coalesced memory access.
 
@@
 
-\subsection{Working of the system}
+\subsection{Computation work flow}
 
@@
 \end{itemize}
-
-Fig.~\ref{fg:system} depicts the flow of the system.
-
-\begin{figure}
- \centerline{ \includegraphics[width=.8\columnwidth]{images/system.eps} }
- \caption{Push-based system flow}
- \label{fg:system}
-\end{figure}
 

### 2016-02-15 paper/paperSource.tex
automatic triggered changes

@@
 It is important to understand that even though there are abstract entities of grids and blocks, it is not real hardware capacities, but they can affect scheduling and performance consequently.
-Decision to go with atom to thread execution is because it is pretty convenient to keep high utilization of hardware and still not lose idea if we are about to incorporate more functions, which makes the system more user-friendly, though later on as a plugin we could add data loaded and transformer between RAM and GPU global memory, because there are a lot of custom ways to aggregate data and take an advantage of coalesced memory access.
+Decision to go with atom to thread execution is because it is pretty convenient to keep high utilization of hardware and still not lose idea if we are about to incorporate more functions, which makes the system more user-friendly, though later on as a plugin we left some room to add data transformer between RAM and GPU global memory, because there are a lot of custom ways to aggregate data and take an advantage of coalesced memory access.
+
+\begin{figure}
+ \centerline{ \includegraphics[width=1.0\columnwidth]{images/threadload} }
+ \caption{SDH thread loading}
+ \label{fg:threadload}
+\end{figure}
 
@@
 
-In this subsection we give a brief overview of how the system works
-at runtime. Please note that the first, preliminary part is only
-executed once, i.e., the data transformation from MS data files to a
-file that our system can read.
+Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
+Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
+\subsubsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
+Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g.
+generic CUDA function atomicAdd).
+We run the first stream with one-body queries and wait till it gets finished.
+The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately.
+Meanwhile, while the computation is running, there is a possibility to use a third stream to load another chunk of data, too.
+First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep in the GPU.
+\subsubsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
+As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity.
+In this particular case it is $O(n^2)$.
+GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence.
+As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it can't it has to wait till other threads finish their jobs.
+Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.
 
@@
 \end{itemize}
-
-Step 2, loading the data into main memory is different for SDH query
-compared to the one for the one-body queries. The reason for that,
-as mentioned earlier, is the different data structure used to store
-the data in memory. While we use double array to store the data for
-the one-body queries, we use quad-tree like data structure to store
-the data needed to compute the SDH. The loading to the double array
-is straightforward. However, to load the quad-tree structure, we
-need to use some of the info from the data itself. Namely, the
-coordinates of the atoms are used to determine in which tree node an
-atom belongs. That way we build the so called density map (DM),
-i.e., the different regions with a certain number of atoms in them
-(including all the atom's attributes). So basically, to solve the
-SDH problem in a push-based manner, we convert the problem into
-populating a data structure in push-based manner. This data
-structure will then be used as an input to our DM-SDH algorithm
-that, although not completely in ``on-the-fly" way, is a great
-improvement over the naive methods used in much of today's MS
-analysis systems (i.e., GROMACS, PDB, CHARMM, etc.).
 

### 2016-02-16 main.cu
automatic triggered changes

@@ -37,6 +37,14 @@ int main(int argc, char *argv[]) {
     * Read name of the file
     */
     std::string fileName = argv[2];
+
+    int workload = 1;
+    if(argc < 3) {
+        workload = atoi(argv[3]);
+    }
+    
+    printf("workload: %d\n", workload);
+
     std::ifstream stream(fileName.c_str());
     std::cout << "Reading file: " << fileName << std::endl;
 

### 2016-02-16 main.cu
automatic triggered changes

@@ -39,7 +39,7 @@ int main(int argc, char *argv[]) {
     std::string fileName = argv[2];
 
     int workload = 1;
-    if(argc < 3) {
+    if(argc > 3) {
         workload = atoi(argv[3]);
     }
     

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -159,7 +159,7 @@ void output_histogram(bucket* hist, int num_buckets){
 }
 
 
-void run_single_kernel(int atomsCnt, atom* atomList) {
+void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
 
     printf("---------GPU-SINGLE-KERNEL---------\n");
 
@@ -170,22 +170,6 @@ void run_single_kernel(int atomsCnt, atom* atomList) {
 
     query_results* res = (query_results*) malloc(sizeof(query_results));
     bucket* histogram = (bucket *)malloc(sizeof(bucket) * num_buckets); 
-    
-    //set default empty values to remove some garbage inside
-    res->mass = 0;
-    res->charge = 0;
-    res->max_x = 0;
-    res->max_y = 0;
-    res->max_z = 0;
-    res->inertiaX = 0;
-    res->inertiaY = 0;
-    res->inertiaZ = 0;
-    res->depoleMoment = 0;
-
-    int i;
-    for(i = 0; i < num_buckets; i++) {
-        histogram[i].d_cnt = 0;
-    }
 
     struct timezone i_dunno;
     struct timeval start_time;
@@ -236,20 +220,39 @@ void run_single_kernel(int atomsCnt, atom* atomList) {
     //int gridSize = ceil(atomsCnt / (float)blockSize) + 1;
     //int stripe = 1024 / ;
 
-    /**
-    * KERNEL CALLS
-    */
-    //mass and charge
-    //----------------------------------1 BODY KERNEL---------------------------------------------------
-    int smem1 = sizeof(float) * block_size.x * 2; //this is not really used for now
-    gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
+    int i, w;
+    for(w = 0; w < workload; w++) {
 
-    //----------------------------------2 BODY KERNEL---------------------------------------------------
-    int smem2 = num_buckets * sizeof(unsigned long long) + block_size.x * sizeof(atom);
-    gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
-    
-    cudaStreamSynchronize(streamComp1);
-    cudaStreamSynchronize(streamComp2);
+        //set default empty values to remove some garbage inside
+        res->mass = 0;
+        res->charge = 0;
+        res->max_x = 0;
[…]

### 2016-02-16 gpu_single.h

@@ -6,6 +6,6 @@
 __global__ void gpu_one_body_functions_kernel(int* g_s_atoms_cnt, atom* g_s_atom_list, query_results* g_s_res);
 __global__ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double PDH_res);
 
-void run_single_kernel(int atomsCnt, atom* atomList);
+void run_single_kernel(int atomsCnt, atom* atomList, int workload);
 
 #endif

### 2016-02-16 main.cu

@@ -42,8 +42,6 @@ int main(int argc, char *argv[]) {
     if(argc > 3) {
         workload = atoi(argv[3]);
     }
-    
-    printf("workload: %d\n", workload);
 
     std::ifstream stream(fileName.c_str());
     std::cout << "Reading file: " << fileName << std::endl;
@@ -81,7 +79,7 @@ int main(int argc, char *argv[]) {
             std::cout << atomCount << " atoms read in previous frame." << std::endl;
 
             if(atomCount > 0) {
-                run_single_kernel(atomsCnt, atomsList);
+                run_single_kernel(atomsCnt, atomsList, workload);
             }
             
             // check for error

### 2016-02-16 main.cu
automatic triggered changes

@@ -65,20 +65,18 @@ int main(int argc, char *argv[]) {
     while(!stream.eof()) {
         //read line from file
         std::getline(stream, line);
-
         std::stringstream lineStream(line);
         
         lineStream >> token;
+
         if(token.compare("HEAD") == 0) {
             //skip the header
-
             std::cout << line << std::endl;
-
             heads++;
-            std::cout << "Frame #" << heads << " processing. " << std::endl;
-            std::cout << atomCount << " atoms read in previous frame." << std::endl;
 
-            if(atomCount > 0) {
+            if(heads > 0 && atomCount > 0) {
+                std::cout << "Frame #" << heads << " processing. " << std::endl;
+                std::cout << atomCount << " atoms read in previous frame." << std::endl;
                 run_single_kernel(atomsCnt, atomsList, workload);
             }
             
@@ -123,8 +121,6 @@ int main(int argc, char *argv[]) {
     
     float elapsed = time_calc(start_time);
     printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
-
-    // run_single_kernel(atoms_cnt, workload);
 
     // check for error
     cudaError_t error = cudaGetLastError();

### 2016-02-16 main.cu
automatic triggered changes

@@ -70,8 +70,6 @@ int main(int argc, char *argv[]) {
         lineStream >> token;
 
         if(token.compare("HEAD") == 0) {
-            //skip the header
-            std::cout << line << std::endl;
             heads++;
 
             if(heads > 0 && atomCount > 0) {

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -161,8 +161,6 @@ void output_histogram(bucket* hist, int num_buckets){
 
 void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
 
-    printf("---------GPU-SINGLE-KERNEL---------\n");
-
     int BOX_SIZE = 175;
     int PDH_res = 1;
 
@@ -214,7 +212,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         grid_size.y = atomsCnt / (block_size.x * maxGridX);
     }
 
-    printf("grid_size.x/y: %d/%d\n", grid_size.x, grid_size.y);
+    printf("Grid Sizes: [%d, %d] \n", grid_size.x, grid_size.y);
 
     //int blockSize = 1024;
     //int gridSize = ceil(atomsCnt / (float)blockSize) + 1;
@@ -268,7 +266,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3f\n", "Inertia Z Axis: ", res->inertiaZ);
     printf("%-40s %.3f\n", "Depole Moment Z Axis: ", res->depoleMoment);
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
-    output_histogram(histogram, num_buckets);
+    //output_histogram(histogram, num_buckets);
 
 
     /**

### 2016-02-16 main.cu

@@ -74,7 +74,7 @@ int main(int argc, char *argv[]) {
 
             if(heads > 0 && atomCount > 0) {
                 std::cout << "Frame #" << heads << " processing. " << std::endl;
-                std::cout << atomCount << " atoms read in previous frame." << std::endl;
+                std::cout << atomCount << " atoms read." << std::endl;
                 run_single_kernel(atomsCnt, atomsList, workload);
             }
             

### 2016-02-16 main.cu
automatic triggered changes

@@ -73,7 +73,7 @@ int main(int argc, char *argv[]) {
             heads++;
 
             if(heads > 0 && atomCount > 0) {
-                std::cout << "Frame #" << heads << " processing. " << std::endl;
+                std::cout << "**********************Frame #" << heads << "*****************" << std::endl;
                 std::cout << atomCount << " atoms read." << std::endl;
                 run_single_kernel(atomsCnt, atomsList, workload);
             }

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -189,9 +189,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     cudaMalloc((void**)&g_s_atomsCnt, sizeof(int));
     cudaMalloc((void**)&d_histogram, num_buckets * sizeof(bucket));
 
-    cudaMemcpy(g_s_res, res, sizeof(query_results), cudaMemcpyHostToDevice);
     cudaMemcpy(g_s_atomsCnt, &atomsCnt, sizeof(int), cudaMemcpyHostToDevice);
-    cudaMemcpy(d_histogram, histogram, num_buckets * sizeof(bucket), cudaMemcpyHostToDevice);
     cudaMemcpy(g_s_atom_list, atomList, sizeof(atom) * atomsCnt, cudaMemcpyHostToDevice);
 
     
@@ -235,6 +233,9 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         for(i = 0; i < num_buckets; i++) {
             histogram[i].d_cnt = 0;
         }
+
+        cudaMemcpy(g_s_res, res, sizeof(query_results), cudaMemcpyHostToDevice);
+        cudaMemcpy(d_histogram, histogram, num_buckets * sizeof(bucket), cudaMemcpyHostToDevice);
 
         /**
         * KERNEL CALLS

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -267,7 +267,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3f\n", "Inertia Z Axis: ", res->inertiaZ);
     printf("%-40s %.3f\n", "Depole Moment Z Axis: ", res->depoleMoment);
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
-    //output_histogram(histogram, num_buckets);
+    output_histogram(histogram, num_buckets);
 
 
     /**

### 2016-02-16 main.cu
automatic triggered changes

@@ -70,8 +70,6 @@ int main(int argc, char *argv[]) {
         lineStream >> token;
 
         if(token.compare("HEAD") == 0) {
-            heads++;
-
             if(heads > 0 && atomCount > 0) {
                 std::cout << "**********************Frame #" << heads << "*****************" << std::endl;
                 std::cout << atomCount << " atoms read." << std::endl;
@@ -112,6 +110,7 @@ int main(int argc, char *argv[]) {
         atomsList[atomCount].z = rand() % 100;
 
         atomCount++;
+        heads++;
     }
 
     printf("\n\n\nHeads: %d\n", heads);

### 2016-02-16 main.cu
automatic triggered changes

@@ -86,6 +86,7 @@ int main(int argc, char *argv[]) {
             }
 
             atomCount = 0;
+            heads++;
             continue;
         }
 
@@ -110,7 +111,6 @@ int main(int argc, char *argv[]) {
         atomsList[atomCount].z = rand() % 100;
 
         atomCount++;
-        heads++;
     }
 
     printf("\n\n\nHeads: %d\n", heads);

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -267,7 +267,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3f\n", "Inertia Z Axis: ", res->inertiaZ);
     printf("%-40s %.3f\n", "Depole Moment Z Axis: ", res->depoleMoment);
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
-    output_histogram(histogram, num_buckets);
+    //output_histogram(histogram, num_buckets);
 
 
     /**

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -269,6 +269,13 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
     //output_histogram(histogram, num_buckets);
 
+    cudaError_t error = cudaGetLastError();
+    if(error != cudaSuccess)
+    {
+        // print the CUDA error message and exit
+        printf("CUDA error: %s\n", cudaGetErrorString(error));
+        exit(-1);
+    }
 
     /**
     * MEM FREE
@@ -276,4 +283,5 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     cudaFree(g_s_atom_list);
     cudaFree(g_s_res);
     cudaFree(g_s_atomsCnt);
+    cudaFree(d_histogram);
 }

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -269,14 +269,6 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
     //output_histogram(histogram, num_buckets);
 
-    cudaError_t error = cudaGetLastError();
-    if(error != cudaSuccess)
-    {
-        // print the CUDA error message and exit
-        printf("CUDA error: %s\n", cudaGetErrorString(error));
-        exit(-1);
-    }
-
     /**
     * MEM FREE
     */
@@ -284,4 +276,12 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     cudaFree(g_s_res);
     cudaFree(g_s_atomsCnt);
     cudaFree(d_histogram);
+
+    cudaError_t error = cudaGetLastError();
+    if(error != cudaSuccess)
+    {
+        // print the CUDA error message and exit
+        printf("CUDA error: %s\n", cudaGetErrorString(error));
+        exit(-1);
+    }
 }

### 2016-02-16 gpu_single.cu
automatic triggered changes

@@ -276,12 +276,4 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     cudaFree(g_s_res);
     cudaFree(g_s_atomsCnt);
     cudaFree(d_histogram);
-
-    cudaError_t error = cudaGetLastError();
-    if(error != cudaSuccess)
-    {
-        // print the CUDA error message and exit
-        printf("CUDA error: %s\n", cudaGetErrorString(error));
-        exit(-1);
-    }
 }

### 2016-02-16 main.cu

@@ -123,6 +123,7 @@ int main(int argc, char *argv[]) {
     cudaError_t error = cudaGetLastError();
     if(error != cudaSuccess)
     {
+        printf("Error! ");
         // print the CUDA error message and exit
         printf("CUDA error: %s\n", cudaGetErrorString(error));
         exit(-1);

### 2016-02-16 main.cu
automatic triggered changes

@@ -119,15 +119,18 @@ int main(int argc, char *argv[]) {
     float elapsed = time_calc(start_time);
     printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
 
+    printf("Before err! ");
+    
     // check for error
     cudaError_t error = cudaGetLastError();
     if(error != cudaSuccess)
     {
-        printf("Error! ");
         // print the CUDA error message and exit
         printf("CUDA error: %s\n", cudaGetErrorString(error));
         exit(-1);
     }
+
+    printf("After err! ");
 
 	return 0;
 }

### 2016-02-16 main.cu
automatic triggered changes

@@ -119,18 +119,18 @@ int main(int argc, char *argv[]) {
     float elapsed = time_calc(start_time);
     printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
 
-    printf("Before err! ");
+    // printf("Before err! ");
     
-    // check for error
-    cudaError_t error = cudaGetLastError();
-    if(error != cudaSuccess)
-    {
-        // print the CUDA error message and exit
-        printf("CUDA error: %s\n", cudaGetErrorString(error));
-        exit(-1);
-    }
+    // // check for error
+    // cudaError_t error = cudaGetLastError();
+    // if(error != cudaSuccess)
+    // {
+    //     // print the CUDA error message and exit
+    //     printf("CUDA error: %s\n", cudaGetErrorString(error));
+    //     exit(-1);
+    // }
 
-    printf("After err! ");
+    // printf("After err! ");
 
 	return 0;
 }

### 2016-02-16 main.cu
automatic triggered changes

@@ -113,14 +113,14 @@ int main(int argc, char *argv[]) {
         atomCount++;
     }
 
-    printf("\n\n\nHeads: %d\n", heads);
-    printf("Atom Count: %d\n", atomCount);
+    //printf("\n\n\nHeads: %d\n", heads);
+    //printf("Atom Count: %d\n", atomCount);
     
-    float elapsed = time_calc(start_time);
-    printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
+    //float elapsed = time_calc(start_time);
+    //printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
 
     // printf("Before err! ");
-    
+
     // // check for error
     // cudaError_t error = cudaGetLastError();
     // if(error != cudaSuccess)

### 2016-02-16 main.cu
automatic triggered changes

@@ -62,9 +62,8 @@ int main(int argc, char *argv[]) {
 
     gettimeofday(&start_time, &i_dunno);
 
-    while(!stream.eof()) {
+    while(std::getline(stream, line)) {
         //read line from file
-        std::getline(stream, line);
         std::stringstream lineStream(line);
         
         lineStream >> token;

### 2016-02-16 main.cu
automatic triggered changes

@@ -112,24 +112,24 @@ int main(int argc, char *argv[]) {
         atomCount++;
     }
 
-    //printf("\n\n\nHeads: %d\n", heads);
-    //printf("Atom Count: %d\n", atomCount);
+    printf("\n\n\nHeads: %d\n", heads);
+    printf("Atom Count: %d\n", atomCount);
     
-    //float elapsed = time_calc(start_time);
-    //printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
+    float elapsed = time_calc(start_time);
+    printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
 
-    // printf("Before err! ");
+    printf("Before err! ");
 
-    // // check for error
-    // cudaError_t error = cudaGetLastError();
-    // if(error != cudaSuccess)
-    // {
-    //     // print the CUDA error message and exit
-    //     printf("CUDA error: %s\n", cudaGetErrorString(error));
-    //     exit(-1);
-    // }
+    // check for error
+    cudaError_t error = cudaGetLastError();
+    if(error != cudaSuccess)
+    {
+        // print the CUDA error message and exit
+        printf("CUDA error: %s\n", cudaGetErrorString(error));
+        exit(-1);
+    }
 
-    // printf("After err! ");
+    printf("After err! ");
 
 	return 0;
 }

### 2016-02-16 main.cu
automatic triggered changes

@@ -69,7 +69,7 @@ int main(int argc, char *argv[]) {
         lineStream >> token;
 
         if(token.compare("HEAD") == 0) {
-            if(heads > 0 && atomCount > 0) {
+            if(atomCount > 0) {
                 std::cout << "**********************Frame #" << heads << "*****************" << std::endl;
                 std::cout << atomCount << " atoms read." << std::endl;
                 run_single_kernel(atomsCnt, atomsList, workload);
@@ -112,24 +112,18 @@ int main(int argc, char *argv[]) {
         atomCount++;
     }
 
-    printf("\n\n\nHeads: %d\n", heads);
-    printf("Atom Count: %d\n", atomCount);
+    //printf("\n\n\nHeads: %d\n", heads);
+    //printf("Atom Count: %d\n", atomCount);
     
     float elapsed = time_calc(start_time);
     printf("%-40s %.3fmillis\n", "Total Running time: ", elapsed);
 
-    printf("Before err! ");
-
     // check for error
     cudaError_t error = cudaGetLastError();
-    if(error != cudaSuccess)
-    {
-        // print the CUDA error message and exit
+    if(error != cudaSuccess) {
         printf("CUDA error: %s\n", cudaGetErrorString(error));
         exit(-1);
     }
-
-    printf("After err! ");
 
 	return 0;
 }

### 2016-02-16 sequential.c

@@ -53,6 +53,12 @@ int main(int argc, char *argv[]) {
     * Read name of the file
     */
     std::string fileName = argv[2];
+
+    int workload = 1;
+    if(argc > 3) {
+        workload = atoi(argv[3]);
+    }
+
     std::ifstream stream(fileName.c_str());
     std::cout << "Reading file: " << fileName << std::endl;
 
@@ -73,10 +79,7 @@ int main(int argc, char *argv[]) {
     query_results* res = (query_results*) malloc(sizeof(query_results));
     bucket* histogram = (bucket *)malloc(sizeof(bucket) * num_buckets); 
 
-    while(!stream.eof()) {
-        //read line from file
-        std::getline(stream, line);
-
+    while(std::getline(stream, line)) {
         std::stringstream lineStream(line);
         
         lineStream >> token;
@@ -85,11 +88,9 @@ int main(int argc, char *argv[]) {
 
             std::cout << line << std::endl;
 
-            heads++;
-            std::cout << "Frame #" << heads << " processing. " << std::endl;
-            std::cout << atomCount << " atoms read in previous frame." << std::endl;
-
             if(atomCount > 0) {
+                std::cout << "**********************Frame #" << heads << "*****************" << std::endl;
+                std::cout << atomCount << " atoms read." << std::endl;
                 //here we run sequential
                 //fix the time
                 gettimeofday(&start_time, &i_dunno);
@@ -105,38 +106,39 @@ int main(int argc, char *argv[]) {
                 res->inertiaZ = 0;
                 res->depoleMoment = 0;
 
-                int i, j;
+                int i, j, w;
                 for(i = 0; i < num_buckets; i++) {
                     histogram[i].d_cnt = 0;
                 }
 
-                //one body
-                for(i = 0; i < atomsCnt; i++) {
-                    res->mass += atomsList[i].mass;
-                    res->charge += atomsList[i].charge;
-                    res->inertiaX += atomsList[i].mass * atomsList[i].x;
-                    res->inertiaY += atomsList[i].mass * atomsList[i].y;
-                    res->inertiaZ += atomsList[i].mass * atomsList[i].z;
-                    res->depoleMoment += atomsList[i].charge * atomsList[i].z;
-                }
+                for(w = 0; w < workload; w++) {
+                    //one body
+                    for(i = 0; i < atomsCnt; i++) {
+                        res->mass += atomsList[i].mass;
+                        res->charge += atomsList[i].charge;
[…]

### 2016-02-19 gpu_single.cu
Shared memory to one body queries

@@ -7,7 +7,7 @@
 __global__
 void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query_results* g_s_res) {
 
-    extern __shared__ int sdata[];
+    extern __shared__ float sdata[];
     
     long index_x = blockIdx.x * blockDim.x + threadIdx.x;
     long index_y = blockIdx.y * blockDim.y + threadIdx.y;
@@ -20,13 +20,6 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
         return;
     }
 
-    //for some reason shared memory is becoming slower
-    // //shared memory structure:
-
-    // //first mass
-    // sdata[tid] = g_s_atom_list[i].mass;
-    // //second charge
-    // sdata[blockDim.x + tid] = g_s_atom_list[i].charge;
 
     // while(i < *g_s_atomsCnt) {
     //     sdata[tid] += g_s_atom_list[i].mass;
@@ -37,20 +30,41 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
     //     __syncthreads();
     // }
 
-    // atomicAdd(&g_s_res->mass, sdata[tid]);
-    // atomicAdd(&g_s_res->charge, sdata[blockDim.x + tid]);
+    float* shared_result = &sdata[0];
 
-    //current atom instance
     atom atomInstance = g_s_atom_list[i];
 
-    atomicAdd(&g_s_res->mass, atomInstance.mass);
-    atomicAdd(&g_s_res->charge, atomInstance.charge);
+    atomicAdd(&shared_result[0], atomInstance.mass);
+    atomicAdd(&shared_result[1], atomInstance.charge);
+    atomicAdd(&shared_result[2], atomInstance.mass * x);
+    atomicAdd(&shared_result[3], atomInstance.mass * y);
+    atomicAdd(&shared_result[4], atomInstance.mass * z);
+    atomicAdd(&shared_result[5], atomInstance.charge * atomInstance.z);
 
-    atomicAdd(&g_s_res->inertiaX, atomInstance.mass * atomInstance.x);
-    atomicAdd(&g_s_res->inertiaY, atomInstance.mass * atomInstance.y);
-    atomicAdd(&g_s_res->inertiaZ, atomInstance.mass * atomInstance.z);
+    __syncthreads();
 
[…]

### 2016-02-19 gpu_single.cu
automatic triggered changes

@@ -36,9 +36,9 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
 
     atomicAdd(&shared_result[0], atomInstance.mass);
     atomicAdd(&shared_result[1], atomInstance.charge);
-    atomicAdd(&shared_result[2], atomInstance.mass * x);
-    atomicAdd(&shared_result[3], atomInstance.mass * y);
-    atomicAdd(&shared_result[4], atomInstance.mass * z);
+    atomicAdd(&shared_result[2], atomInstance.mass * atomInstance.x);
+    atomicAdd(&shared_result[3], atomInstance.mass * atomInstance.y);
+    atomicAdd(&shared_result[4], atomInstance.mass * atomInstance.z);
     atomicAdd(&shared_result[5], atomInstance.charge * atomInstance.z);
 
     __syncthreads();

### 2016-02-19 paper/paperSource.tex
automatic triggered changes

@@
 The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately.
-Meanwhile, while the computation is running, there is a possibility to use a third stream to load another chunk of data, too.
-First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep in the GPU.
+While the computation is running, there is a possibility to use a third stream to load another chunk of data, too.
+First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
 \subsubsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
@@
 GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence.
-As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it can't it has to wait till other threads finish their jobs.
+As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs.
 Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.
 
-Here are the steps taken through out the analysis:
+For SDH we the following strategy
+
 \begin{itemize}
-\item[(1)] Execute the data transformer
-\begin{itemize}
-\item[(a)] Read the MS data from trajectory files
-\item[(b)] Extract the info needed for our system
-\item[(c)] Save the read data to a file recognizable to the system
-\end{itemize}
-\item[(2)] Load the data into main memory (one frame at a time)
-\begin{itemize}
-\item[(a)] Load data into a double array (for one-body queries)
-\item[(b)] Load data into the quad-tree structure (for two-body queries:SDH, RDF)
-\end{itemize}
-\item[(3)] Push the data to all queries
-\item[(4)] A query, if available, acts upon the pushed data (first executing
-the lower level, sub-queries)
-\item[(5)] Store intermediate results (results
-of sub-queries)
-\item[(6)] Repeat steps 3-5 if needed.
-\item[(7)] Output results
-\item[(8)] Go to step 2 and load the next frame (if needed).
+\item[1.] Find current index (for target atom)
+\item[2.] Find range of atoms it has to interact with in an equally loaded fashion
+\item[3.] Load the range of atoms into shared memory
+\item[4.] Iterate through the range of atoms from shared memory and atomically write the distances counts into buckets pre-allocated in share memory
+\item[5.] Synchronize with other threads in the block (not in the entire grid)
+\item[6.] If it is the first thread in block, add all bucket counts from shared memory into global memory
 \end{itemize}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure}
- \centerline{ \includegraphics[width=.99\columnwidth]{images/workload_set_up1.eps} }
- \caption{Workload setup}
- \label{fg:workload_set_up}
-\end{figure}
 
-\section{Experimental results}\label{sc:experiments}
+\section{Experiments Results}\label{sc:experiments}
 
-The system was implemented in C++ programming language and tested on
-real molecular simulation data sets. The experiments were carried
-out on an Apple MacPro machine with 8GB of physical memory and two
-Quad-Core Intel Xeon 3GHz processors. The MacPro was running OS X
-Mavericks $10.9.3$ operating system. We have compared the results
-obtained by our system to those obtained by running the analysis
-through GROMACS (v. $4.5.7$). Both systems were analyzing the same
-data sets.
+The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed C++.
+Experiments were run on operating systems which didn't have any additional workload besides usual installation, and the operating systems particularly were Amazon Linux, which doesn't have a lot of modifications except for some additional features for management with Amazon EC-2 Console.
+We have generated mock data sets with different variations of atoms, frames and ran them with different workloads comparing sequential version of Push-based system with improved version written in CUDA.
+
+\emph{\textbf{Hardware stack.}}
+There are primarily several parts that are really important in this hardware stack.
+Since this is primarily big data processing, we care about processing power and memory.
+RAM on the node that we were running has 16GB.
+The server allocated from Amazon has the following CPU parameters:
+
+\begin{center}
+  \begin{tabular}{ l  l }
+
+Architecture & x86\_64 \\  
+CPU op-mode(s) & 32-bit, 64-bit \\  
+Byte Order & Little Endian \\  
+CPU(s) & 8 \\  
+On-line CPU(s) list & 0-7 \\  
+Thread(s) per core & 2 \\  
+Core(s) per socket & 4 \\  
+CPU family & 6 \\  
+Model & 45 \\  
+Model name & Intel Xeon E5-2670 2.60GHz \\  
+Stepping & 7 \\  
+CPU MHz & 2593.814 \\  
+L1d cache & 32K \\  
+L1i cache & 32K \\  
+L2 cache & 256K \\  
+L3 cache & 20480K \\  
+NUMA node0 CPU(s) & 0-7 \\
+
+  \end{tabular}
+\end{center}
+
+The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
+Just to mentioned, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load.
+Consequently, we also have GPU device with the following parameters:
+
+\begin{center}
+  \begin{tabular}{ l l }
+Name & Grid 34C K520 \\
+Memory & 4095Mib \\ 
+Total Nvidia CUDA Cores & 1536 \\ 
+Memory Size & 8 GB GDDR5 (4 GB/GPU) \\
+Total GPUs & 2 GK104 GPUs \\
+  \end{tabular}
+\end{center}
 

### 2016-02-25 paper/paperSource.tex
automatic triggered changes

@@
 
-\emph{\textbf{Data sets.}} In our experiments, six data sets from
-different simulations were used. All simulations were done on a
-POPC~\footnote{POPC is a chemical compound composed of a
-diacylglycerol and phospholipid. Its full name is
-1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine and it is one of
-the most important lipids in bio-physical molecular simulation.}
-lipid bilayer, but were all set to produce data of different sizes
-(i.e., different number of particles in the simulation). Namely, we
-have tested the system on simulations with $52,400$; $209,600$;
-$838,400$; $2.5M$; $4.4M$; and $8.8M$ atoms. Also, since the
-simulations were run separately, they produced six different MS
-systems with distinct characteristics (distinct structure, atom's
-positioning, etc.). From all of the generated data sets we have
-randomly selected sets of $10$, $100$, and $1000$ consecutive frames
-for the purpose of our experiments. This gave us $18$ different
-datasets on which we tested our system.
+\emph{\textbf{Data sets.}} As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
+It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format.
+We have generated number of mock data sets in range of $1K-8M$ atoms with $10, 100, 1000$ frames each.
+For each data set we run workloads with the values: $1, 5, 25, 50, 100$ .
+Number of atoms and frames are obviously up to the scientist and depend on simulations, but workloads are a bit trickier.
+First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests.
+This possibility will be elaborated in more details in Future Work section.
 
-\emph{\textbf{Query work load.}} Two types of query workload were
-used: 1) one involving one-body queries only 2) one including
-two-body queries (SDH and RDF) as well. The reason for this is that
-GROMACS, the system we used to compare our system to, only has a
-naive method of solving the RDF (SDH) problem (like almost all MS
-analysis systems). In our system we have incorporated SDH (RDF)
-algorithms that are far more superior to the naive method, and
-comparing the systems like that would not have been fair (we
-believe).
+\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
 
-\emph{\textbf{One-body queries only.}} The following set of one-body
+The following set of one-body
 queries were included in the test workload: mean square displacement
@@
 
+\begin{figure}
+ \centerline{ \includegraphics[width=1.0\columnwidth]{images/chart_1000}}
+ \caption{Speedup for 1000 atoms}
+ \label{fg:chart_1000}
+\end{figure}
+
+\begin{figure}
+ \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_10000}}
+ \caption{Speedup for 10000 atoms}
+ \label{fg:chart_10000}
+\end{figure}
+
+\begin{figure}
+ \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_max}}
+ \caption{Speedup for larger number of atoms}
+ \label{fg:chart_max}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Benchmark}
-
 Through extensive collaboration with a research group from the
@@
 benchmark for MS analysis systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{comment}
-
-\begin{figure}
- \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames52K.eps} }
- \caption{Showing speedup for different workload}
- \label{fg:results2-10frames}
-\end{figure}
-
-\begin{figure}
- \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames52K.eps} }
- \caption{Showing speedup for different workload}
- \label{fg:results2-100frames}
-\end{figure}
-
-\begin{figure}
- \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames209K.eps} }
- \caption{Showing speedup for different workload}
- \label{fg:results3-10frames}
-\end{figure}
-
-\begin{figure}
- \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames209K.eps} }
- \caption{Showing speedup for different workload}
- \label{fg:results3-100frames}
-\end{figure}
-
-\end{comment}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
-
-%\begin{figure}
-% \centerline{ \includegraphics[width=.8\columnwidth]{images/speedup10frames4.2M.eps} }
-% \caption{Showing speedup for different workload}
-% \label{fg:results5-10frames}
-%\end{figure}
 

### 2016-02-25 paper/paperSource.tex
automatic triggered changes

@@
 \emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
+More improvements can be done for acceleration of this approach, but in the system we have not gone too far with data converter module, because initial framework has to be simple and user friendly.
 
-The following set of one-body
-queries were included in the test workload: mean square displacement
-(msd), radius of gyration, dipole moment, center of mass, velocity
-autocorrelation, electron density, mass density, and charge density.
-This set of queries were pointed to us, by a group in the physics
-field with extensive MS background, as one of the most commonly used
-in the field of collagen bilayer MS system analysis. A workload
-group contains all $8$ queries executed on one of the $12$
-selections, making $12$ groups. Such groups are executed on six
-different size data sets, with $10$, $100$ and $1000$ frames. This
-workload is then repeated $5$ more times, by executing each of the
-queries in the groups $5$, $10$, $25$, $50$, and $100$ times,
-essentially just magnifying the workload intensity. In total, we
-have $12 x 6 x 3 x 6 = 1,296$ different workload setups to test the
-system on. Fig.~\ref{fg:workload_set_up} shows the organization of
-the workload setup.
+\subsection{Results}
+Results are nicely represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
+You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is the following.
+There is a small overhead we have to consider before computing part.
+In order to let GPU process the data, we have to transfer the data to Global Memory, and upon finish we have to load the results back to RAM.
+In this particular case, for small workload and frames, it is even slower than the sequential processing.
+Afterwards, the bigger workload the more convincing speedup gets, unfortunately, it doesn't get more than $4$ times for 1000 atoms.
+Fortunately, scientists usually work with much larger data sets, thus, this doesn't really represent general speedup for average use case.
+
+\emph{\textbf{10000 atoms.}}
+As you can see, speedup bumps up to almost 30 with 10000 atoms (Figure \ref{fg:chart_10000}).
+Once the overhead on memory transfer is covered by computation speedup, it keeps growing.
+It is important to pay attention to the growth of speedup over growth of workload.
+It is not that significant, because even though CPUs are much slower, they still use smart caching and internal multi-core processing, which is pretty fast, too.
 
@@
 \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Benchmark}
-Through extensive collaboration with a research group from the
-Physics department at USF, we have come up with a benchmark that can
-be used for testing the efficiency of an analysis system for
-molecular simulations. The benchmark consist of three essential
-parts: 1. Simulation data produced by an MS, 2. Queries that are to
-be executed onto that data in order to produce some information of
-interest, and 3.Benchmark parameters that control the size of the
-benchmark.
-
-\subsubsection{Benchmark Data}
-The data used in the benchmark was real molecular simulation data,
-produced through the GROMACS MS system. The initial, pre-simulation
-data file consisted of $200$ POPC and $12000$ solvent molecules, or
-$12200$ molecules in total. This type of system was used because it
-is sufficiently diverse, containing enough distinct POPC and solvent
-molecules (e.g., each POPC molecule includes approximately $52$
-different atoms) and yet simple enough to be easily transformed into
-another system of different size. By using the $genconf$ function in
-GROMACS, we produced pre-simulation files of different sizes
-(essentially by changing the system's size (box)). $Six$ different
-sized pre-simulation files were created. A molecular simulation was
-then run on these $6$ files, each producing an MS system of certain
-size (volume/number of particles). All of the simulations were set
-up to produce $1000$ frames (snapshots in time of the systems), each
-frame containing the same number of particles as the base one. The
-produced files contained: $52,400$; $209,600$; $838,400$; $2.5M$;
-$4.4M$; and $8.8M$ atoms per frame. So, for example, the file with
-$52,000$ atoms holds $52,000,000$ records in total ($1000$ frames,
-each containing $52,000$ records). As mentioned earlier, this
-simulation data comes mostly in binary formats and in trajectory
-files having a lot of unneeded overhead. Therefore, it was
-transformed to a data arrays files containing only crucial
-information of the particles and the system. The size of the files
-ranged from $135MB$ for $52,000$ atoms to $24GB$ for $8.8$ million
-atoms (this is for data with $100$ frames).
-
-%Fig.~\ref{fg:dataformat} represents the organization/structure of
-%these files.
-
-\subsubsection{Benchmark Queries}
-The queries selected to be included in this benchmark were derived
-through a thorough observation of the way an MS system is being
-analyzed. They were found to be the base of the analysis of many MS
-systems. In other words, no mater how small or big the analysis was,
-these queries were included in that analysis. As mentioned earlier,
-they are of two types: one-body (and algebraic) and two-boy (and
-holistic). Table~\ref{tb:queries} shows these queries.
-
-\subsubsection{Benchmark Parameters} There are several parameters
-that can be used to control the overall size of the system. We divide the parameters into two groups:\\
-Data size parameters:
-\begin{itemize}
-\item[-]Select different sized dataset
-\item[-]Number of frames
-\item[-]Data selection (within the selected dataset) onto which the queries are being executed
-\end{itemize}
-Workload size parameters:
-\begin{itemize}
-\item[-]Number of queries to be executed
-\item[-]Number of times each query is executed
-\end{itemize}
-By changing these parameters, we can produce a versatile testing
-benchmark for MS analysis systems.
-
-\subsection{Results} We have run extensive experiments over all the
-different setups of workload mentioned previously in this section.
-However, in this paper we present only the workload setups of $4$
-data size sets: $838,400$, $2.5M$, $4.2M$, and $8.8M$ atoms because
-we believe they convey enough information about the efficiency of
-our system compared to that of the Gromacs system. The running times
-of our push-based system were compared to those of the Gromacs
-system. The first set of figures, namely
-Figures~\ref{fg:speedup838K-select-level}-\ref{fg:speedup8.8M-select-level},
-represent the speedup that our system obtains over the Gromacs
-system with various atoms selection levels. We define the selection
-levels based on the number of comparisons we have to make in order
-to extract the needed group(selection) of atoms. For example, if we
-want to do analysis on all molecules containing oxygen, or hydrogen,
-or carbon we would go over each molecule and compare its components
-to the selection list. The bigger the selection list, the higher the
-select level in our system. For better visualization, we note three
-different selection levels: high (at least $10$ comparisons made),
-medium (between $1$ and $10$ comparisons made), and low select level
-(with one or less comparisons made). As seen in the figures, for
-high selection level, the speedup is smaller compared to that
-achieved in low select levels. The reason for this, we believe is in
-that the amount of time our system spends extracting the atoms group
-increases with the level of selection. Even though our system still
-shows considerable speedup over Gromacs in high level selections, we
-do believe there is room for improvement in our system and that is
-our immediate future work we are planning on doing. These figures
-also show the relation of the speedup to the workload intensity,
-i.e., the higher the workload intensity the higher the speedup.
-
-The connection between the workload intensity and the speedup is
-better represented in the next set of figures,
-Figures~\ref{fg:speedup838K}-\ref{fg:speedup8.8M}. They show the
-speedup our system achieves over the Gromacs system on a varying
-workload intensity. Each of those figures show the speedup with
-different dataset sizes (e.g., $838,000$, $2,567,600$ atoms, etc.),
-including $10$, $100$, and $1000$ data frames. The speedup is
-calculated simply as a ratio between the running time of our system
-on a certain set of workload and that of the Gromacs system on the
-same workload. These figures show that the speedup over varying
-workload intensity achieved by our system ranges anywhere from about
-$10$ to $1000$ times, depending on the size of the dataset, number
-of frames and the selection of the atoms.
-
-Figure~\ref{fg:speedup-average-over-all-workload}, shows the speedup
-our system achieves over all workload intensity (average workload
-intensity) with varying dataset sizes. It is clear that, again our
-system has better performance than the Gromacs system. The speedup
-presented in this set of figures ranges anywhere from about $15$ to
-$650$ times.
-
-The last figure, Figure~\ref{fg:speedup-average-workload-select},
-shows the speedup our system achieves over all workload intensity
-and all select levels with varying dataset sizes. This figure, in a
-way, summarizes the previous two sets of figures, bringing together
-the workload and the different selections through the average. It is
-clear that, again our system has better performance than the Gromacs
-system. The speedup ranges anywhere from $50$ to $250$.
-
-All four sets of figures show that such push-based design has clear
-advantages over the pull-based type of design incorporated in the
-Gromacs system.
 
@@
 
-\noindent{\bf Acknowledgements:} The project described was supported
-by an Award (R01GM086707) from the National Institute Of General
-Medical Sciences (NIGMS) at the National Institutes of Health (NIH).
-The authors would like to thank Anand Kumar who has contributed his
-time and knowledge toward completion of the work.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \bibliographystyle{IEEEtran}

### 2016-03-01 paper/paperSource.tex
automatic triggered changes

@@
 Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
-Figure \ref{fg:reduction} nicely represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel.
+Figure \ref{fg:reduction} represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel.
 Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes.
@@
 \subsection{Results}
-Results are nicely represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
-You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is the following.
-There is a small overhead we have to consider before computing part.
-In order to let GPU process the data, we have to transfer the data to Global Memory, and upon finish we have to load the results back to RAM.
+Results are represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
+You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
+There is a small overhead we have to consider before computing part: in order to let GPU process the data, we have to transfer the data to Global Memory, and upon completion it is needed to load the results back to RAM to show it to the user.
 In this particular case, for small workload and frames, it is even slower than the sequential processing.
@@
 
+Fortunately, scientific data is much bigger than above examples.
+You can see average speedup for bigger number of atoms on Figure \ref{fg:chart_max}.
+The speedup gets upto 70 times which could save huge amount of time for scientists.
+These results are not the best, as it has been already mentioned earlier, since we have not modified general structure of memory being transferred to Global Memory not taking advantage of coalesced memory access to make it easily extendable.
+
 \begin{figure}
@@
 
+It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
+It is because we've used static approach of computing grid dimensions.
+Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
+
 \section{Conclusions and Future Work}\label{sc:conclusion}
-The objective of our work is to design and implement improved data
-analysis system that can be used in the field of molecular
-simulation system's analysis. In this paper, we introduce the idea
-for such system. We build our system on a push-based type design,
-where data from data arrays is being pushed onto available queries
-in the system. These queries are being executed on the pushed data
-and produce intermediate / final result that would be used as part
-of the data analysis. We are able to achieve an improvement over
-existing, pull-based type designs because of the I/O overhead such
-designs introduce when dealing with large volumes of scientific
-data. Also, our queries are able to be executed on the same stream
-of data, making it suitable solution for streaming circumstances. We
-designed a benchmark that can be used to test data analysis systems.
-This benchmark comprises of three parts: 1)benchmark data,
-2)benchmark queries, and 3) benchmark parameters. We use this
-benchmark to compare our system to one of the most frequently used
-MS analysis systems, Gromacs. The efficiency and speedup achieved by
-our system is supported by extensive experiments and their results.
-The results show that our push-based design achieves up to about
-$1000$ times speedup in comparison to a pull-based design, i.e.,
-Gromacs.
-
-One direction of our future work will be to further improve our
-push-based design. Through the extensive experiments we have learned
-that our design can be improved when the atom selection clause
-involves many conditions. This improvement may be in the direction
-of improving the algorithmic part, but it can also be in the
-direction of improving the data presentation/organization we have
-used in the system.\\
+The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
+As you could see from result on charts (Figure \ref{fg:chart_1000}, \ref{fg:chart_max}), the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time.
+The results have been gathered with a single snapshot of the framework tracked in control version system which can be improved with means of data converter module and dynamic grid dimensions computation module.
+In the future, this system has all potentials to be expanded into a distributed system, where frames could be aggregated by nodes and processed in parallel.\\
 

### 2016-04-19 thesis_paper/source.tex
paper stuff

--- 2016-03-01 paper/paperSource.tex
@@
-\documentclass[10pt,journal,final,letterpaper,twocolumn]{IEEEtran}
-%\documentclass{ssdbm}
+\documentclass[11pt,a4paper]{report}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{microtype}
+\usepackage[dvips]{graphicx}
+\usepackage{xcolor}
+\usepackage{times}
+\usepackage{ragged2e}
+\usepackage{titlesec}
 \usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
@@
 \usepackage[justification=centering]{caption}
-%\usepackage{fixltx2e}
 
-\restylefloat{table}
+\usepackage[
+breaklinks=true,colorlinks=true,
+%linkcolor=blue,urlcolor=blue,citecolor=blue,% PDF VIEW
+linkcolor=black,urlcolor=black,citecolor=black,% PRINT
+bookmarks=true,bookmarksopenlevel=2]{hyperref}
 
-% This was from the NEW paper
-%----------------------------------------------------------
-\RequirePackage{fix-cm}
+\usepackage{geometry}
+% PDF VIEW
+\geometry{total={210mm,297mm},
+left=25mm,right=25mm,%
+bindingoffset=0mm, top=25mm,bottom=25mm}
+%PRINT
+\geometry{total={210mm,297mm},
+left=20mm,right=20mm,
+bindingoffset=0mm, top=15mm,bottom=35mm}
 
-\DeclareMathOperator{\var}{var}
-\newcounter{qcounter}
+\linespread{1.5}
 
-%\smartqed  % flush right qed marks, e.g. at end of proof
-%----------------------------------------------------------
+\titleformat{\chapter}
+{\normalfont\LARGE\bfseries}{\thechapter.}{1em}{}
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{definition}{Definition}
-\newtheorem{remark}{Remark}
-
-\newcommand{\SEC}[1] {Section \ref{#1}}
-\newcommand{\Eq}[1] {Equation \ref{#1}}
-\newcommand{\eq}[1] {Eq. (\ref{#1})}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
 
@@
 
-% paper title
-\title{Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+%   TITLEPAGE
+%
+%   due to variety of titlepage schemes it is probably better to make titlepage manually
+%
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+\thispagestyle{empty}
 
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
+{%%%
+\sffamily
+\centering
 
-\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$, Vladimir Grupcev, Joseph Fogarty, Sagar Pandit % <-this % stops a space
+{\Large 
+Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU
+}
 
-%\thanks{* These authors contributed equally to this work.}
-\thanks{$\dagger$ Author to whom all correspondence should be sent.}
-\thanks{Iliiazbek Akhmedov, Vladimir Grupcev, and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E.
-Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A.
-Emails:
-akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
-\thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
-University of South Florida, 4202 E. Fowler Ave., ISA 2019, Tampa,
-FL 33620, U.S.A. Emails: jcfogart@mail.usf.edu, pandit@usf.edu} }
+\vspace{1.5cm}
 
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
+{
+by
+}
+\vspace{.5cm}
 
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
+{ 
+Iliiazbek Akhmedov
+}
 
-% make the title area
-\maketitle
+\vspace{3.5cm}
+
+A thesis submitted in partial fulfillment
+
+of the requirements for the degree of
+
+Master of Science in Computer Science
+
+Department of Computer Science and Engineering
+
+College of Engineering
+
+University of South Florida
+
+\vspace{3.5cm}
+Major Professor:  Yicheng Tu, Ph.D.
+\vspace{\fill}
+
+\vspace{.5cm}
+Date of Approval:
+
+N/A
+\vspace{\fill}
+
+Copyright \textcopyright\space 2016, Iliiazbek Akhmedov
+
+%%%
+}%%%
+
+\clearpage
+
+\pagestyle{plain}
+\centering
+
+\textbf{DEDICATION}
+\vspace{0.5cm}
+
+\begin{flushleft}
+\hspace{3em} I dedicate my thesis work to my family who have always been supporting me throughout my life.
+I also dedicate it to my closest friends who have shared my passion and encouraged me to explore and try my ideas in life.
+\end{flushleft}
+
+\clearpage
+
+\pagestyle{plain}
+\centering
+\textbf{ACKNOWLEDGMENTS}
+\vspace{0.5cm}
+
+\begin{flushleft}
+\hspace{3em} I would like to thank my major professor throughout my master thesis Dr.
+Yicheng Tu for his guidance and knowledge that I received through all the time I worked with him.
+I really appreciate his contribution and very happy that I got a chance to work with him.
+I as well want to thank the committee members for their support and encouragement.
+\end{flushleft}
+
+\clearpage
+
+\tableofcontents
+
+\clearpage
 
 \begin{abstract}
-
+\justify
 Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
@@
 In the new approach due to high performance of GPU processing and its nature, custom data structures were not even needed much though it didn't bear any loss in precision and performance.
-
 \end{abstract}
 
-\begin{IEEEkeywords}
-Push-based system, molecular simulation, scientific databases,
-spatial distance histogram, GPU, parallel processing, CUDA.
-\end{IEEEkeywords}
-%\vspace{-5mm}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Introduction}
 
-\section{Introduction}\label{sc:intro}
-
+\justify
 In various sciences simulation systems take big place and often times they may be the clue for results.
@@
 
-{\small
 \begin{enumerate}
@@
 \end{enumerate}
-}
+
 \vspace{3mm}
@@
 \begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/sample_snapshot_of_simulation.eps} }
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/sample_snapshot_of_simulation.eps} }
  \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle}}
@@
 
-\subsection{Problem Statement}
+\section{Problem Statement}
 Simulation software systems, in general, follow the same methodology of running and storing simulation data.
@@
 
-\subsection{Our approach with improvement}
+\section{Our approach with improvement}
 
@@
 
-\subsection{Contribution and roadmap of the paper} 
+\section{Contribution and roadmap of the paper} 
 The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis even faster now.
@@
 
-\vspace{3mm}
-\section{Related Work}\label{sc:relatedwork}
-\vspace{3mm}
+\chapter{Related Work}\label{sc:relatedwork}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\vspace{3mm}
 \section{Network of Queries}\label{sc:querynetwork}
-\vspace{3mm}
-
 Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
@@
 
-\vspace{3mm}
-\section{The system design}\label{sc:system}
-\vspace{3mm}
-
-The proposed design has a pretty straightforward structure which is going to be explained in this section.
-Shortly, the system starts with original input files (or data) and gets consumed in a sequential manner by simple disk IO framework into random access memory by chunks, then, chunk gets uploaded to GPU device and processed.
-At some point, the results that need to be shown to the user are transported from global memory of GPU (where they initially get accumulated) to the main memory, and then might get written as well to an output file.
-General data flow map is represented in Figure  \ref{fg:data-flow}.
-
-\begin{figure}
- \centerline{ \includegraphics[width=0.65\columnwidth]{images/memory-organization.eps} }
- \caption{Data flow in the improved design of push-based system for MS analysis}
- \label{fg:data-flow}
-\end{figure}
-
-\subsection{Data input}
-Since the size of the data in the problem is huge, organization of the data flow is crucial.
-As mentioned before, original data is retrieved from simulation software output files and in our case it is Gromacs Software's trajectory files\footnote{The file types, their structure, and their decompression has been well explained in the previous paper in section "Data orgranization in main memory"}.
-For a simulation, there can be several files of different format and usually they are also compressed without any loss of information.
-Tool for decompressing those files to a flat format for each frame and atom has been developed in the previous work\cite{mainPaper}.
-
-In order to experiment with different input cases, the input files have been generated with a tool that we have built in Python language.
-Thus, we are able to generate mock data files with any number of frames and atoms, and the values are random and can be adjusted with needs of the experiment (for example, for cases when we want to analyze bucket widths in a limited space, the coordinates are generated in range of $0 - 1$, etc.).
-As it has already been mentioned earlier, we don't expect the data to be distributed with a certain behavior, the importance of having this kind of tool to generate mock data is having a variety of random data sets that could reveal any issues related to values provided to the program, on top of that, having the actual file written and then being read has been a part of the experiments to have more realistic use case.
-In the future, the same script for generating the mock data could be used for regression and performance tests in a distributed implementation of this design as a single source.
-
-\subsection{Data flow architecture}
-
-MS data roughly can be represented as a list of atoms with physical properties aggregated as frames.
-Thus, we consider a single frame as a chunk which will be loaded into the memory\footnote{Memory in this design might be interpreted differently, since we have a separate device memory on GPUs, but it will be explained better later} and pushed through our network of query.
-In this case, it is very convenient, because there are almost no dependencies between frames, rather than list of atoms in the same frame, which potentially again could create additional disk I/O overhead.
-
-Having defined the chunk, we load it from flat format file to RAM.
-A single atom data structure has all the attributes such as charge, mass, velocity, type of atom, coordinates, and so on.
-In other words, we have an array of atoms, and most of the attributes (in case of data structure, it is a class field), are simply primitive floats or doubles depending on precision requirements.
-Next, the same chunk of data with the same structure is loaded onto GPU global memory.
-GPU devices, in general, have about 4-5 types of memory, and global memory is the biggest and slowest.
-Consequently, depending on algorithm or query we load parts of data chunk to another types of memory called shared memory.
-It depends on the actual query, because the same part of chunk can be loaded and unloaded in regards with shared memory several times, especially with SDH.
-
-\subsection{Network of queries organization}
-
-Since we assume that our framework should be easily used by a scientist doing MS analysis, it is important to keep queries module as a separate piece of code to make it easily extendable.
-In order to understand how queries module is organized and the reason for it, we need to understand how CUDA and GPU work.
-
-\emph{\textbf{GPU execution}}.
-From hardware standpoint, it is important to understand that every GPU device consists of SM (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
-Having this in mind, since this is a pure parallel processing power, the goal is to increase hardware throughput and keep highest utilization to exploit every register given.
-CUDA is an amazing tool that tries to help developers map this hardware capabilities with the actual application layer.
-In CUDA we have grids, blocks, and threads, where grid is 3 dimensional unit where each cell is a block, and the same relationship applies between blocks and threads\footnote{This explanation doesn't include memory setup in GPU devices, but it is covered in further sections}.
-Originally, this mapping was perfect for solving image processing problems or other graphics computation, since dimensions and such division (e.g.
-by pixels to threads) is very inherent.
-As a language feature, in CUDA we have to define such a function called 'kernel' function.
-Essentially, it is the function where we are supposed to map every execution thread with data and perform the actual execution, this approach is widely known as SIMD\footnote{SIMD - Single Instruction Multiple Data}.
-As a convenience we have separated one-body functions with multiple-body functions into separate kernels.
-
-\emph{\textbf{One-body queries kernel}}.
-Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
-Figure \ref{fg:reduction} represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel.
-Having an array of atom data structures, we reduce the array accumulating all the needed data from each atom's attributes.
-We don't expect much performance improvement from coalesced memory access, since the data structure for atom can eliminate any gain of coalesced memory access, though for specific types of queries it could be possible to aggregate certain attributes into merged arrays and take an advantage of it.
-In this case, the solution should be simple in order to keep it easily extendable for other simple one-body queries for a scientist to program.
-We fetch all the atoms from GPU global memory, since we use references to atoms once.
-In this case, allocating shared memory and copying it over would complicate the implementation and possibly create additional overhead for one time reference usage.
-Finally, memory transportation and data consumption are implemented for user, thus, he only needs to extend the queries module which consequently can be turned into plugins model.
-
-\emph{\textbf{Two-body queries kernel}}.
-SDH (or RDF) is the algorithm that requires special treatment in our design.
-The point, as mentioned earlier, is that this function requires more than one iteration of the given set of atoms, but rather all combinations of pairs in a given frame, thus, it is not linear complexity.
-An implementation of this type of query might be very non-intuitive because of the nature of GPUs.
-Another issue is that since to take an advantage on this type of computation, before writing an implementation for it, a user must have a good understanding of GPU architecture which makes the system less user-friendly.
-At this point, it is more important to achieve good performance improvement results and then take care of ease of system extensions.
-As we already mentioned, GPU threads execute in parallel and consume certain amount of work followed by best try of all hardware capacities utilization, but, unfortunately, it is also important to take care of equal workload for every thread independently, so, that utilization is kept high.
-For this reason, we've come up with a solution that should increase performance of the computation which will be described later.
-
-\begin{figure}
- \centerline{ \includegraphics[width=0.65\columnwidth]{images/reduction} }
- \caption{Array reduction example\cite{gpureduction}}
- \label{fg:reduction}
-\end{figure}
-
-\begin{figure}
- \centerline{ \includegraphics[width=0.65\columnwidth]{images/overallarchitecture} }
- \caption{Overall system architecture}
- \label{fg:overallarchitecture}
-\end{figure}
-
-\emph{\textbf{Architecture overview.}}
-This section is to emphasize general architecture of the system and summarize the work flow as well as the data flow.
-In total, we have flat files with ready flat values for each atom aggregated in frames coming into the program as data input using common disk I/O framework of an operating system.
-Then, once we have loaded a frame (which is considered to be a chunk of the entire data set) onto RAM of our machine, we copy over the entire frame onto the GPU device global memory.
-It is important that, at this point, we never write any values to simulation files or any other data related to simulation, thus, we don't have to care about any synchronization of incoming data.
-After this, we have two predefined kernel functions (which were explained before), where first kernel stands for one-body queries, and the second one is for two-body query SDH.
-Just for architecture clearance and possible future improvements, two kernels are executed in two different streams\footnote{In CUDA stream is another abstract layer of organizing execution threads.
-They are not guaranteed to be executed at the same time or after each other, GPU scheduling architecture is completely different from CPU having a warp to be single scheduling unit at maximum of 32 threads at a time} (You can see them on Figure\ref{fg:overallarchitecture}).
-In each kernel we apply certain heuristics to improve memory access using such architecture features of GPU as shared memory and blocking of kernel threads.
-
-\emph{\textbf{GPU grid and block sizes.}} As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
-The approach we came up with is 1 execution thread per atom.
-For CUDA version 2.0 maximum number of threads per block is 1024.
-Blocks are 3 dimensional entities in a 3 dimensional grid.
-The less abstract unit is, the less overhead we have in communication between them, thus, we try to utilize smallest units more.
-For example, for 1000,000 atoms, we would allocate 1024 threads per block as a static hard coded value, since we usually have always more than 1024 atoms in simulations.
-One dimension of grid can take up to 64,000 blocks, thus, if N is bigger than $1024 * 64,000$, than the rest would be allocated to $grid.y$ and $grid.z$ respectively.
-We have not considered numbers bigger than GPU can allocate, which could be explored in future work.
-It is important to understand that even though there are abstract entities of grids and blocks, it is not real hardware capacities, but they can affect scheduling and performance consequently.
-Decision to go with atom to thread execution is because it is pretty convenient to keep high utilization of hardware and still not lose idea if we are about to incorporate more functions, which makes the system more user-friendly, though later on as a plugin we left some room to add data transformer between RAM and GPU global memory, because there are a lot of custom ways to aggregate data and take an advantage of coalesced memory access.
-
-\begin{figure}
- \centerline{ \includegraphics[width=1.0\columnwidth]{images/threadload} }
- \caption{SDH thread loading}
- \label{fg:threadload}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Computation work flow}
-
-Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
-Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-\subsubsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
-Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g.
-generic CUDA function atomicAdd).
-We run the first stream with one-body queries and wait till it gets finished.
-The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately.
-While the computation is running, there is a possibility to use a third stream to load another chunk of data, too.
-First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
-\subsubsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
-As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity.
-In this particular case it is $O(n^2)$.
-GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence.
-As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs.
-Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.
-
-For SDH we the following strategy
-
-\begin{itemize}
-\item[1.] Find current index (for target atom)
-\item[2.] Find range of atoms it has to interact with in an equally loaded fashion
-\item[3.] Load the range of atoms into shared memory
-\item[4.] Iterate through the range of atoms from shared memory and atomically write the distances counts into buckets pre-allocated in share memory
-\item[5.] Synchronize with other threads in the block (not in the entire grid)
-\item[6.] If it is the first thread in block, add all bucket counts from shared memory into global memory
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiments Results}\label{sc:experiments}
-
-The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed C++.
-Experiments were run on operating systems which didn't have any additional workload besides usual installation, and the operating systems particularly were Amazon Linux, which doesn't have a lot of modifications except for some additional features for management with Amazon EC-2 Console.
-We have generated mock data sets with different variations of atoms, frames and ran them with different workloads comparing sequential version of Push-based system with improved version written in CUDA.
-
-\emph{\textbf{Hardware stack.}}
-There are primarily several parts that are really important in this hardware stack.
-Since this is primarily big data processing, we care about processing power and memory.
-RAM on the node that we were running has 16GB.
-The server allocated from Amazon has the following CPU parameters:
-
-\begin{center}
-  \begin{tabular}{ l  l }
-
-Architecture & x86\_64 \\  
-CPU op-mode(s) & 32-bit, 64-bit \\  
-Byte Order & Little Endian \\  
-CPU(s) & 8 \\  
-On-line CPU(s) list & 0-7 \\  
-Thread(s) per core & 2 \\  
-Core(s) per socket & 4 \\  
-CPU family & 6 \\  
-Model & 45 \\  
-Model name & Intel Xeon E5-2670 2.60GHz \\  
-Stepping & 7 \\  
-CPU MHz & 2593.814 \\  
-L1d cache & 32K \\  
-L1i cache & 32K \\  
-L2 cache & 256K \\  
-L3 cache & 20480K \\  
-NUMA node0 CPU(s) & 0-7 \\
-
-  \end{tabular}
-\end{center}
-
-The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
-Just to mentioned, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load.
-Consequently, we also have GPU device with the following parameters:
-
-\begin{center}
-  \begin{tabular}{ l l }
-Name & Grid 34C K520 \\
-Memory & 4095Mib \\ 
-Total Nvidia CUDA Cores & 1536 \\ 
-Memory Size & 8 GB GDDR5 (4 GB/GPU) \\
-Total GPUs & 2 GK104 GPUs \\
-  \end{tabular}
-\end{center}
-
-\emph{\textbf{Data sets.}} As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
-It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format.
-We have generated number of mock data sets in range of $1K-8M$ atoms with $10, 100, 1000$ frames each.
-For each data set we run workloads with the values: $1, 5, 25, 50, 100$ .
-Number of atoms and frames are obviously up to the scientist and depend on simulations, but workloads are a bit trickier.
-First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests.
-This possibility will be elaborated in more details in Future Work section.
-
-\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
-More improvements can be done for acceleration of this approach, but in the system we have not gone too far with data converter module, because initial framework has to be simple and user friendly.
-
-\subsection{Results}
-Results are represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
-You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
-There is a small overhead we have to consider before computing part: in order to let GPU process the data, we have to transfer the data to Global Memory, and upon completion it is needed to load the results back to RAM to show it to the user.
-In this particular case, for small workload and frames, it is even slower than the sequential processing.
-Afterwards, the bigger workload the more convincing speedup gets, unfortunately, it doesn't get more than $4$ times for 1000 atoms.
-Fortunately, scientists usually work with much larger data sets, thus, this doesn't really represent general speedup for average use case.
-
-\emph{\textbf{10000 atoms.}}
-As you can see, speedup bumps up to almost 30 with 10000 atoms (Figure \ref{fg:chart_10000}).
-Once the overhead on memory transfer is covered by computation speedup, it keeps growing.
-It is important to pay attention to the growth of speedup over growth of workload.
-It is not that significant, because even though CPUs are much slower, they still use smart caching and internal multi-core processing, which is pretty fast, too.
-
-\begin{figure}
- \centerline{ \includegraphics[width=1.0\columnwidth]{images/chart_1000}}
- \caption{Speedup for 1000 atoms}
- \label{fg:chart_1000}
-\end{figure}
-
-\begin{figure}
- \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_10000}}
- \caption{Speedup for 10000 atoms}
- \label{fg:chart_10000}
-\end{figure}
-
-Fortunately, scientific data is much bigger than above examples.
-You can see average speedup for bigger number of atoms on Figure \ref{fg:chart_max}.
-The speedup gets upto 70 times which could save huge amount of time for scientists.
-These results are not the best, as it has been already mentioned earlier, since we have not modified general structure of memory being transferred to Global Memory not taking advantage of coalesced memory access to make it easily extendable.
-
-\begin{figure}
- \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_max}}
- \caption{Speedup for larger number of atoms}
- \label{fg:chart_max}
-\end{figure}
-
-It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
-It is because we've used static approach of computing grid dimensions.
-Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
-
-\section{Conclusions and Future Work}\label{sc:conclusion}
-The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
-As you could see from result on charts (Figure \ref{fg:chart_1000}, \ref{fg:chart_max}), the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time.
-The results have been gathered with a single snapshot of the framework tracked in control version system which can be improved with means of data converter module and dynamic grid dimensions computation module.
-In the future, this system has all potentials to be expanded into a distributed system, where frames could be aggregated by nodes and processed in parallel.\\
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{IEEEtran}
+\bibliographystyle{unsrt}
 \bibliography{references.bib}
 
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendices
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2016-04-19 thesis_paper/source.tex
paper stuff

@@
 
+\newgeometry{left=20mm,right=20mm,top=35mm,bottom=35mm}
+
 \pagestyle{plain}
@@
 \clearpage
+\restoregeometry
 
@@
 \justify
-Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
+\hspace{3em} Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
 Traditional database management systems follow principle of pulling the needed data, processing it, and then returning the results.
@@
 When working with huge volume of data, there are non-trivial issues arise.
-For example, it can take days and weeks to analyze big enough data sets, because the data cannot be simply loaded into memory, since it can get up to terabytes, thus, it has extra overhead because of widely used random access disk I/O framework to read from disk chunks by chunks.
+For example, it can take days and weeks to analyze big enough data sets, because the data cannot be simply loaded into memory, since it can get up to terabytes, thus, it has extra overhead because of traditional random access disk I/O framework to read from disk chunks by chunks.
 Besides it, analysis of the data can get even more complicated, since going through certain parts of the data once would not be enough, which leads to low throughput and efficiency of loaded data.
@@
 
-As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks. .
+As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks.
 "In order for a system to be able to perform analytical examination of the data produced in such streaming media, the system should have the capability of fast data access.
-The reason, the millions of data records(tweets) produced
-every second. Moreover, these tweets may have
-different geographical origin, introducing different languages and
-forms and often times containing unsolicited messages, errors,
+The reason, the millions of data records (tweets) produced every second.
+Moreover, these tweets may have different geographical origin, introducing different languages and forms and often times containing unsolicited messages, errors,
 malicious content, etc. Therefore, some low level data uniformity
@@
 \begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
  \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainPaper}}

### 2016-05-18 thesis_paper/source.tex
Thesis paper changes

--- 2016-03-01 paper/paperSource.tex
@@
 \usepackage[justification=centering]{caption}
-%\usepackage{fixltx2e}
+\usepackage{url}
 
-\restylefloat{table}
+\usepackage[
+breaklinks=true,colorlinks=true,
+%linkcolor=blue,urlcolor=blue,citecolor=blue,% PDF VIEW
+linkcolor=black,urlcolor=black,citecolor=black,% PRINT
+bookmarks=true,bookmarksopenlevel=2]{hyperref}
 
-% This was from the NEW paper
-%----------------------------------------------------------
-\RequirePackage{fix-cm}
+\usepackage{geometry}
+% PDF VIEW
+\geometry{total={210mm,297mm},
+left=25mm,right=25mm,%
+bindingoffset=0mm, top=25mm,bottom=25mm}
+%PRINT
+\geometry{total={210mm,297mm},
+left=20mm,right=20mm,
+bindingoffset=0mm, top=15mm,bottom=35mm}
 
-\DeclareMathOperator{\var}{var}
-\newcounter{qcounter}
+\linespread{1.5}
 
-%\smartqed  % flush right qed marks, e.g. at end of proof
-%----------------------------------------------------------
+\titleformat{\chapter}
+{\normalfont\LARGE\bfseries}{\thechapter.}{1em}{}
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}{Lemma}
-\newtheorem{definition}{Definition}
-\newtheorem{remark}{Remark}
-
-\newcommand{\SEC}[1] {Section \ref{#1}}
-\newcommand{\Eq}[1] {Equation \ref{#1}}
-\newcommand{\eq}[1] {Eq. (\ref{#1})}
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
 
@@
 
-% paper title
-\title{Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+%   TITLEPAGE
+%
+%   due to variety of titlepage schemes it is probably better to make titlepage manually
+%
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+\thispagestyle{empty}
 
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
+{%%%
+\sffamily
+\centering
 
-\author{Iliiazbek~Akhmedov, Yi-Cheng~Tu~$\dagger$, Vladimir Grupcev, Joseph Fogarty, Sagar Pandit % <-this % stops a space
+{\Large 
+Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU
+}
 
-%\thanks{* These authors contributed equally to this work.}
-\thanks{$\dagger$ Author to whom all correspondence should be sent.}
-\thanks{Iliiazbek Akhmedov, Vladimir Grupcev, and Yi-Cheng Tu are with the Department of Computer Science and Engineering, University of South Florida, 4202 E.
-Fowler Ave., ENB 118, Tampa, FL 33620, U.S.A.
-Emails:
-akhmedovi@mail.usf.edu, ytu@cse.usf.edu}
-\thanks{Joseph Fogarty and Sagar Pandit are with the Department of Physics,
-University of South Florida, 4202 E. Fowler Ave., ISA 2019, Tampa,
-FL 33620, U.S.A. Emails: jcfogart@mail.usf.edu, pandit@usf.edu} }
+\vspace{1.5cm}
 
-%\pubid{0000--0000/00\$00.00~\copyright~2002 IEEE}
+{
+by
+}
+\vspace{.5cm}
 
-% use only for invited papers
-%\specialpapernotice{(Invited Paper)}
+{ 
+Iliiazbek Akhmedov
+}
 
-% make the title area
-\maketitle
+\vspace{3.5cm}
+
+A thesis submitted in partial fulfillment
+
+of the requirements for the degree of
+
+Master of Science in Computer Science
+
+Department of Computer Science and Engineering
+
+College of Engineering
+
+University of South Florida
+
+\vspace{3.5cm}
+Major Professor:  Yicheng Tu, Ph.D.
+\vspace{\fill}
+
+\vspace{.5cm}
+Date of Approval:
+
+N/A
+\vspace{\fill}
+
+Copyright \textcopyright\space 2016, Iliiazbek Akhmedov
+
+%%%
+}%%%
+
+\clearpage
+
+\newgeometry{left=20mm,right=20mm,top=35mm,bottom=35mm}
+
+\pagestyle{plain}
+\centering
+
+\textbf{DEDICATION}
+\vspace{0.5cm}
+
+\begin{flushleft}
+\hspace{3em} I dedicate my thesis work to my family who have always been supporting me throughout my life.
+I also dedicate it to my closest friends who have shared my passion and encouraged me to explore and try my ideas in life.
+\end{flushleft}
+
+\clearpage
+
+\pagestyle{plain}
+\centering
+\textbf{ACKNOWLEDGMENTS}
+\vspace{0.5cm}
+
+\begin{flushleft}
+\hspace{3em} I would like to thank my major professor throughout my master thesis Dr.
+Yicheng Tu for his guidance and knowledge that I received through all the time I worked with him.
+I really appreciate his contribution and very happy that I got a chance to work with him.
+I as well want to thank the committee members for their support and encouragement.
+\end{flushleft}
+
+\clearpage
+\restoregeometry
+
+\tableofcontents
+
+\clearpage
 
 \begin{abstract}
-
-Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
+\justify
+\hspace{3em} Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
 Traditional database management systems follow principle of pulling the needed data, processing it, and then returning the results.
 This approach is then optimized by means of caching, storing in different structures, or doing some sacrifices on precision of the results to make it faster.
-When it comes to the point of doing various queries that require analysis  of the whole data,  this design has the following disadvantages: considerable overhead on random I/O while reading from the simulation output files and low throughput of the data that consequently results in long latency, and, if there was any indexing to optimize selections, overhead of storing those becomes too big, too.
+When it comes to the point of doing various queries that require analysis  of the whole data,  this design has the following disadvantages: considerable overhead on traditional disk random I/O framework while reading from the simulation output files and low throughput of the data that consequently results in long latency, and, if there was any indexing to optimize selections, overhead of storing those becomes too big, too.
+Besides it, indexing will also cause delay during write operations and since most of the queries work with the entire data sets, indexing loses its point.
 
-There is a new approach to this problem presented in the previous paper -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and its primary steps are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filter the data and collect all the needed information which increases efficiency and data throughput.
+There is a new approach to this problem -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and its primary steps are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filters the data and collect all the needed information which increases efficiency and data throughput.
 It has a considerable advantage in analysis of molecular simulation data, because it normally involves all the data sets to be processed by the queries.
@@
 Its major difference with the previous design is usage of GPU for the actual processing part of the data flow.
-Using the same scan-based I/O framework the data is pushed through the network of queries which are processed by GPU, and due to the nature of science simulation data, this gives a big advantage for processing it faster and easier.
+Using the same scan-based I/O framework the data is pushed through the network of queries which are processed by GPU, and due to the nature of science simulation data, this gives a big advantage for processing it faster and easier (it will be explained more in later sections).
 In the old approach there were some custom data structures such as quad-tree for calculation of histograms to make the processing faster and those involved loss of data and some expectations from the data nature, too.
-In the new approach due to high performance of GPU processing and its nature, custom data structures were not even needed much though it didn't bear any loss in precision and performance.
-
+In the new approach due to high performance of GPU processing and its nature, custom data structures were not even needed much, though it didn't bear any loss in precision and performance.
 \end{abstract}
 
-\begin{IEEEkeywords}
-Push-based system, molecular simulation, scientific databases,
-spatial distance histogram, GPU, parallel processing, CUDA.
-\end{IEEEkeywords}
-%\vspace{-5mm}
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Introduction}
 
-\section{Introduction}\label{sc:intro}
-
+\justify
 In various sciences simulation systems take big place and often times they may be the clue for results.
@@
 \end{enumerate}
-}
+
 \vspace{3mm}
 The purpose of this paper is primarily focused on the third step, which basically involves the entire simulation data to be processed by the analysis software.
+It is important to underline scale of data being discussed.
+Molecular simulations represent systems starting from hundreds thousands of atoms and more which are tracked by frames as snapshot of specific time and can last for months.
+Thus, the size of data starts from couple of gigabytes up to terabytes and petabytes.
 
@@
 
-As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks. .
+As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks.
 "In order for a system to be able to perform analytical examination of the data produced in such streaming media, the system should have the capability of fast data access.
-The reason, the millions of data records(tweets) produced
-every second. Moreover, these tweets may have
-different geographical origin, introducing different languages and
-forms and often times containing unsolicited messages, errors,
+The reason is millions of data records (tweets) produced every second.
+Moreover, these tweets may have different geographical origin, introducing different languages and forms and often times containing unsolicited messages, errors,
 malicious content, etc. Therefore, some low level data uniformity
 and cleaning on top of the data access and management issues should
-be considered and possibly incorporated in the process of analytical
-investigation in order to achieve relevant result. "\cite{mainPaper, nature_bigdata08,nature_bigdata12,science_social10,jcs_twitter11}
+be considered and possibly incorporated in the process of analytic
+investigation in order to achieve relevant result".\cite{mainPaper, nature_bigdata08,nature_bigdata12,science_social10,jcs_twitter11} Another good example would be data scale of Facebook Inc.
+Being the largest social network, it has huge data coming in daily \cite{fbStats}.
+The data warehouse of Facebook is measured to be 300PB in 2014, and the daily growth is 600TB, most of the raw data is logs, which, most of the times, are used for analytics.
+Analytical queries involve entire data sets that will be handy for any future strategical planning in big companies, being able to observe data behavior and much more.
 
 The primary focus and problem in this paper is scientific data analysis.
+Examples brought up above are real life problems that are very similar to the scientific problem that we are focusing on in terms of data sizes.
 Particles simulation is one of the most popular methods of analyzing certain chemical reactions, physical processes, or other behavior of different materials.
 Molecular simulations (Molecular Dynamics) are applied in different fields and represent a method of analyzing physical movements of particles, atoms, and molecules in a fixed space with a given period of time, apparently with a possibility of giving initial state for each item that is involved in the process and can affect the system.
-This system is an N-body simulation.
+These systems are most of the time N-body simulations.
 The number of atoms in simulations vary in hundred of thousands, particularly, we may observe two simulation systems of a collagen fiber structure and dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system consisting of 890,000 and 402,400 atoms respectively on Figure \ref{fg:collagen_dipal}.
@@
 Since storage of the simulation data is very expensive, it might come to the point of analyzing the data on the fly (meaning running the simulation and analyzing it at the same time in a streaming manner), which leads to a problem of optimizing the processing part of the design proposed in the previous paper, because time spent on generating data should be tried to conceal the time spent on the processing part by means of overlapping or simply running it in a quick manner.
+We will talk about this in details as well in further sections.
 
 \begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/sample_snapshot_of_simulation.eps} }
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/sample_snapshot_of_simulation.eps} }
  \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle}}
@@
 The approach proposed in the previous paper is very good in terms of performance in comparison with the original pull-based system\cite{mainPaper}.
+As you can see on Figure \ref{fg:sample_estimation_old_paper}, it gets up to over 100 times of speedup which is quite crucial, if we think about days of computations.
 The problem is still that some of the queries processed by those means are still improvable, especially taking into consideration the fact that in the used previous works for calculation of 2-body functions, which take the biggest time for processing~\cite{ytu:icde09, EDBT12}, we might have some error bounds, which might be unacceptable for certain simulation analysis where sacrifice on loss of data is unbearable.
 Besides it, although we might have huge performance boost on specific data structure as density map, we still have certain expectation from data nature (such as its uniformity), thus, it makes sense to have desire to simplify the processing and still having pure computation based on any values, considering that the queries should be available to be pre-programmed by user in a separate module of the code.
+All the benchmarks will be done in comparison with already push-based system versus push-based system with GPU improvements.
+We think that it is  not fair to compare pull-based against GPU push-based approach, since the goal of the paper is to demonstrate capabilities of GPU approach that can beat simple sequential push-based system.
 
 \begin{figure}
- \centerline{ \includegraphics[width=1\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
  \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainPaper}}
@@
 
-\subsection{Our approach with improvement}
+\section{Our approach with improvement}
 
+Reviewing the approach we are coming up with and what it was before helps us analyzing the weak parts and things that we can focus on to eliminate old bottlenecks.
+
+Original non-improvement standard approach with pull-based querying is probably one of the most flexible, because nobody has to think about improvements and complexities, it does it for every chunk of data and pulls it whenever it is needed.
+As it has already been mentioned, at this point, the biggest challenge is that supposing we have $M$ queries, for $N$ atoms we would have to load every chunk $M$ times for $N$ chunks of data (frames), and apparently it is very time consuming of I/O and non-linear computations, and that is only for linear functions.
+Improvements could be done on caching level, but again the worst case would be the same complexity, because most of the time there is nothing really to cache, and if there is, it is too big to store it in memory, so, it just has to be re-read from file.
+Using accumulative data structures and trying to optimize by means of them doesn't help either, because, first of all, it is too big data to store in memory, and, secondly, most of the time it is based on data distribution expectation of some kind, which is not the best solution, because the experiments can be different, thus data distribution is sometimes not predictable (boiling down to a specific physics problem will destroy the abstractness and applicability of the system to variety of other problems and will narrow it down making it less useful).
+In summary, the bottlenecks in this approach are I/O time consumption and slow data processing (especially for non-linear computation).
+
+After all, push-based system for solution of the given problem is tremendously fast in comparison with pull-based system.
+General idea is to push a chunk of data (frame) through the pre-defined network of queries that will be needed to run on the data sets, which basically means chunk of data will be read from file and loaded into memory only once in order to go through all the queries.
+As you can observe \cite{mainPaper}, it can speed up for more than 100 times (depending on workload, number of atoms, and functions being used).
+It is clear that the biggest challenges in this approach were related to implementation of non-linear functions like SDH, since once the data is pushed through the queries, it is almost not coming back (if it is coming back, the point of optimization is lost), and the idea is to come up with approach to eliminate this overhead and weakness.
+In this case, it is also important what memory exactly is being used, because trying not to use used parts of the same chunk also matters, because L-level caches of CPU might play a huge role in big data sets.
+One of the solutions is building quad-tree based data structure that will help optimize the calculation, and it has some expectation from data distribution which can speed up depending on dimensions (2D and 3D).
+This is a mutual problem of both data loading (memory usage) and its processing, and push-based system is a great approach, because it still remains as an abstract concept applicable to different queries and probably wide variety of other problems of similar kind in different fields.
+
+Finally, we've analyzed some main ideas, advantages, and disadvantages of our approach in this paper which is basically push-based system with optimization on data processing part by means of GPU computation (in other words, parallelism).
 Reading the data from files is a huge overhead, thus, speedup of push-based system in general is significant over pull-based system in our given case.
@@
 Since the nature of the data that comes with simulation is basically physical properties of atoms, there is a lot of computation that involves independent primitive mathematical operations, which makes it a perfect problem for parallelism.
+For example, image processing computations and graphics processing are done on GPUs, because they have a lot of minor primitive independent pixel-based computations.
 
-Although the original idea of push-based system for molecular simulation data analysis was focused on optimizing throughput and usage of loaded data, there were some extra approaches specifically for 2-body functions.
+As it has already been mentioned, although the original idea of push-based system for molecular simulation data analysis was focused on optimizing throughput and usage of loaded data, there were some extra approaches specifically for 2-body functions.
 For example, Density Map for Spacial Distance Histogram was used in order to avoid additional memory allocation and latency reduction with proven error bounds.
@@
 
-\subsection{Contribution and roadmap of the paper} 
-The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis even faster now.
-Taking an advantage of GPU devices nature and incorporating parallelism and streaming for different types of queries, we have come up with a good speed up over sequential processing, which already had a good speed up in terms of performance in comparison with original pull-based approach.
-Since in this paper we don't expect certain form of data, we have come up with a tool that generates mock data for any number of atoms and frames, which can be used for performance tests that were done consequently, too.
+\section{Contribution and roadmap of the paper}
+
+In collaboration with Physics Department, the original motivation came from the need of improving analytic query executions performance over big data being generated by simulation software.
+The idea was to take input stream coming from simulation software (ideally live) and run network of queries very efficiently, because sometimes they might take days and more to compute.
+The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis fast having an ability to expand network of queries and customize computation functions to get the desired results.
+Our system primarily represents a system that works in several modules that were developed in C, CUDA, Python programming languages that, in combination, give us the desired result to perform efficient computations for the network of queries developed in the previous paper \cite{mainPaper} as well as to prove the concept of optimization of push-based system by means of parallelism with no data loss.
+We've developed the following modules:
+
+\begin{itemize}
+\item[1.] Mock data streamer and generator - tool developed in Python programming language that helps to run performance tests (could also be of other simplest tests like regression) on any kind of data and can be used to simulate random values for physical properties of atoms.
+The simplest usage of the tool is running a python script with arguments (number of atoms and number of frames).
+\item[2.] Main execution script - the main module developed in C and CUDA for the actual computation of the network of queries.
+Network of queries comes predefined and has to be compiled in the script (though later can be abstracted as a configuration file or argument for interpreting general functions), along with computation implementation function that interact with a given GPU.
+\item[3.] Data conversion - this module comes as a part of the code in Main execution script for the users to be able to manipulate data conversion before loading it to GPU.
+This is for advanced users having capability of optimizing computations and queries on any level of memory hierarchy (CPU and GPU).
+It is an abstract part which has to be used upon specific cases to optimize specific functions, since there is no way to predict what functions need to be run and how they can be optimized.
+\item[4.] Sequential CPU based Push-based system - we re-created original push-based system with the same network of queries as a benchmark for estimation of speedup results of the new approach
+
+\end{itemize}
+
+Taking an advantage of GPU devices nature and incorporating parallelism and streaming for different types of queries, we have come up with a good speed up over sequential processing which gets up to 60 times faster, which is just for sampling with no deep improvements on individual queries (since we believe that it should represent average user usage).
+Original push-based system already had a good speed up in terms of performance in comparison with original pull-based approach and floats around more than 100 times faster depending on number of atoms, frames, and workload.
+The module for generation of mock data in Python  was used for benchmarks, and it represents realistic and pessimistic approach of estimation, since there is no data distribution for physical properties values.
+We believe that it might be quite important, because experiments may vary widely.
 This mock data has exactly the same information that would come with Gromacs simulation files, and it has exactly the same format as in the previous paper \cite{mainPaper} right before loading it to memory.
@@
 In general, major contribution of this paper is the framework of push-based system with an incorporated parallelism based on CUDA programming language.
-Besides it, there was developed a tool which generates mockup data of the same format as of the real MS simulations and represent a python script with arguments given as number of frames and atoms.
-
+This tool can be integrated into open source project of Gromacs as a custom plugin that can be used for optimized analytic queries to be run over data sets live streaming as well as from files.
+This will be discussed in more details in further chapters.
 The structure of the framework which will be described later is developed in such a way that one can easily add new queries based on their complexity or modify existing ones adding appropriate selections and filters.
@@
 
-\vspace{3mm}
-\section{Related Work}\label{sc:relatedwork}
-\vspace{3mm}
+\chapter{Related Work}\label{sc:relatedwork}
+
+Push-based System for molecular simulation data analysis with its improved GPU features has couple of main focuses in terms of technical implementation and challenges related to those that have already been tried to be solved by other different approaches and software.
+Here is the review of other works.
+
+\section{Scientific software}
+
+First of all, this system is a tool for scientific computation and in this case particularly for molecular simulation data analysis.
+PeriScope is a project and initiative by University of Michigan which is a system that will optimize data management for scientific explorations.
+Relational Database Management Systems were created to optimize business applications and it was revolutionary.
+Those couldn't be built without these systems with abstract languages like SQL, since it manipulates and automatically takes use of indexes, etc.
+PeriScope is a project that tries to come up with same approach against scientific data and computation with respectively larger scale.\cite{PeriScope} It is kind of difficult to compare our proposed system with PeriScope, but the original motivation is very similar: optimize scientific computation.
+PeriScope is a larger project directed towards more abstract solution like RDBMs that would work for any type of data.
+Another existing tool for scientific data and optimized processing is SciDB.
+It was created to make scientific data storage and querying better and faster.
+It is very similar to traditional RDBMS systems in terms of querying language and represent data storage in tables, but states to be faster than Postgres in 2 orders of magnitude.
+Some of the main features is distributed processing with atomicity and durability (the latter is quite difficult for large scale data), while the disadvantages are no-overwriting (only append mode), single value columns only (no arrays, etc), where records are like simple tuples with join function on querying multiple tables \cite{SciDB}.
+One of the major differences between Push-based Molecular simulation data analysis tool and the projects above is that our proposed design doesn't do anything with storing the data, because we are mainly concentrating it as a side plugin for processing and querying data in an optimized way, while the actual data generation and storage will be done with the actual simulation software.
+
+\section{Stream processing}
+Another major modern issue is live data streaming and live processing, which is also possible with our system.
+Basically, streaming function is a big challenge and need nowadays, because such networks as Twitter and Facebook are able to keep track of live trends and events, including people's reaction and feedback that have to be reflected real time for tremendous amount of users across the globe.
+Borealis Stream Processing Software is a tool being developed by MIT, Brown, and Brandeis Universities.
+It's primary goal is a distributed stream processing and based on Medusa and Aurora.
+Some of the major functions of the software are highly available distributed large streams processing live, dynamic and highly-scalable with possibility of changing queries live and dynamic vision of query results.
+\cite{borealis} There is also a similar need in scientific world, since being able to keep track of results in real time might be crucial to be able to make certain explorations.
+
+\section{Big Data Frameworks}
+
+We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data.
+Nowadays there are bunch of frameworks that try to solve this problem.
+Some of these frameworks are Hadoop \cite{hadoop}, Hadoop with GPU extension \cite{hadoopgpu}, Spark \cite{spark}, Flink \cite{flink} (this is more of a big data stream processing framework) and other ones that are possibly just custom modifications and distributions of these.
+The mentioned tools help dealing with big data giving an ability to process them in batches on clusters using underlying HDFS (Highly Distributed File System), and normally go along with other utilities that help to apply MapReduce concept.
+In fact, the frameworks listed above could even be use for our problem on top of our system, because, as we discuss in section \ref{sc:conclusion}, it could be extended to a distributed system for analyzing independent frames in parallel.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\vspace{3mm}
-\section{Network of Queries}\label{sc:querynetwork}
-\vspace{3mm}
-
+\chapter{Network of Queries}\label{sc:querynetwork}
 Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
@@
 
+\section{One body functions}
 \emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
 
+\section{Multiple body functions}
 \emph{\textbf{Multiple body functions}} such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
@@
 \vspace{3mm}
-\section{The system design}\label{sc:system}
+\chapter{The system design}\label{sc:system}
 \vspace{3mm}
@@
 
-\subsection{Data input}
+\section{Data input}
 Since the size of the data in the problem is huge, organization of the data flow is crucial.
@@
 
-\subsection{Data flow architecture}
+\section{Data flow architecture}
 
@@
 
-\subsection{Network of queries organization}
+\section{Network of queries organization}
 
@@
 \begin{figure}
- \centerline{ \includegraphics[width=1.0\columnwidth]{images/threadload} }
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/threadload} }
  \caption{SDH thread loading}
@@
 
-\subsection{Computation work flow}
+\ section{Computation work flow}
 
@@
 Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-\subsubsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
+\subsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
 Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g.
@@
 First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
-\subsubsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
+\subsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
 As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity.
@@
 
-\section{Experiments Results}\label{sc:experiments}
+\chapter{Experiments Results}\label{sc:experiments}
 
@@
 
-\subsection{Results}
+\section{Results}
 Results are represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
@@
 \begin{figure}
- \centerline{ \includegraphics[width=1.0\columnwidth]{images/chart_1000}}
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/chart_1000}}
  \caption{Speedup for 1000 atoms}
@@
 \begin{figure}
- \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_10000}}
+ \centerline{\includegraphics[width=0.5\columnwidth]{images/chart_10000}}
  \caption{Speedup for 10000 atoms}
@@
 \begin{figure}
- \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_max}}
+ \centerline{\includegraphics[width=0.5\columnwidth]{images/chart_max}}
  \caption{Speedup for larger number of atoms}
@@
 
-\section{Conclusions and Future Work}\label{sc:conclusion}
+\chapter{Conclusions and Future Work}\label{sc:conclusion}
 The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+%\bibliographystyle{plain}
 \bibliographystyle{IEEEtran}
 \bibliography{references.bib}
-
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendices
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2016-05-24 thesis_paper/source.tex
changes to thesis paper

@@
 
-\linespread{1.5}
+\linespread{2}
 
-\titleformat{\chapter}
-{\normalfont\LARGE\bfseries}{\thechapter.}{1em}{}
+\titleformat{\chapter}[display]
+        {\normalfont\Large\bfseries\centering}{\chaptertitlename\ \thechapter}{0pt}{\Large\uppercase}
+
+\renewcommand{\contentsname}{CONTENTS}
+\renewcommand{\listfigurename}{LIST OF FIGURES}
+\renewcommand{\listtablename}{LIST OF TABLES}
+\renewcommand{\abstractname}{ABSTRACT}
+
+% Make chapter titles uppercase is a pain.
+% The internal chapter macros must be redefined.
+\makeatletter
+\renewcommand\chapter{\if@openright\cleardoublepage\else\clearpage\fi
+                    \thispagestyle{plain}%
+                    \global\@topnum\z@
+                    \@afterindentfalse
+                    \secdef\@chapter\@schapter}
+\def\@chapter[#1]#2{\ifnum \c@secnumdepth >\m@ne
+                       \if@mainmatter
+                         \refstepcounter{chapter}%
+                         \typeout{\@chapapp\space\thechapter.}%
+                         \addcontentsline{toc}{chapter}%
+                         {\protect\numberline{\thechapter}\texorpdfstring{\MakeUppercase{#1}}{#1}}%
+                       \else
+                       \addcontentsline{toc}{chapter}{\texorpdfstring{\MakeUppercase{#1}}{#1}}%
+                       \fi
+                    \else
+                    \addcontentsline{toc}{chapter}{\texorpdfstring{\MakeUppercase{#1}}{#1}}%
+                    \fi
+                    \chaptermark{#1}%
+                    \addtocontents{lof}{\protect\addvspace{10\p@}}%
+                    \addtocontents{lot}{\protect\addvspace{10\p@}}%
+                    \if@twocolumn
+                      \@topnewpage[\@makechapterhead{#2}]%
+                    \else
+                      \@makechapterhead{#2}%
+                      \@afterheading
+                    \fi}
+\makeatother
 
@@
 
+\linespread{1}
+
 \tableofcontents
 
+\addcontentsline{toc}{chapter}{\listfigurename}
+\listoffigures
+
+\addcontentsline{toc}{chapter}{\listtablename} 
+\listoftables
+
+\linespread{2}
 \clearpage
 
+\addcontentsline{toc}{chapter}{\abstractname} 
 \begin{abstract}
@@
  \centerline{ \includegraphics[width=0.5\columnwidth]{images/sample_snapshot_of_simulation.eps} }
- \caption{ Snapshots of two MS systems: a collagen fiber structure with 890,000 atoms (top) and a dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system with 402,400 atoms (bottom) \cite{mainArticle}}
+ \caption{ Snapshots of two MS systems with 890,000 atoms with 402,400 atoms}
  \label{fg:collagen_dipal}
@@
  \centerline{ \includegraphics[width=0.5\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
- \caption{ Speed up over different levels of atom selection for 838,400 atoms \cite{mainPaper}}
+ \caption{ Speed up over different levels of atom selection for 838,400 atoms}
  \label{fg:sample_estimation_old_paper}
@@
 
+As a plugin for optimizing computation of scientific queries against large data sets being produced, Push-based System for Molecular Simulation Data Analysis with GPU improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
+
 \section{Big Data Frameworks}
@@
 We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data.
+Traditional software tools are not able to process Big Data apparently because of huge data sets, thus, special frameworks were developed to handle them.
 Nowadays there are bunch of frameworks that try to solve this problem.
@@
 The mentioned tools help dealing with big data giving an ability to process them in batches on clusters using underlying HDFS (Highly Distributed File System), and normally go along with other utilities that help to apply MapReduce concept.
-In fact, the frameworks listed above could even be use for our problem on top of our system, because, as we discuss in section \ref{sc:conclusion}, it could be extended to a distributed system for analyzing independent frames in parallel.
+The relationship between our system and this concept along with its implementations is in data sets sizes.
+Even though we are not using or introducing any non-traditional file systems, in certain scale and problems it could be required, too.
+In fact, the frameworks listed above could even be used for our problem on top of our system, because, as we discuss in section \ref{sc:conclusion}, it could be extended to a distributed system for analyzing independent frames in parallel.
+
+\section{GPU and Big Data Processing}
+GPU is becoming more popular being used by not only graphics processing, but machine learning, deep learning, search algorithms, sorting, and so on.
+Having an ability to optimize in tremendous speedups it is being involved in video classifications, natural language processing.
+Particularly about distributed processing and GPU applications along with it, there has been made some overview and research on what is the current state for common solutions in industry.
+"There were a lot of custom GPU-Hadoop frameworks after the launch of Mars project.
+These include C-MR, SteamMR, GPMR, Grex, Panda, Shredder and many others.
+However, most of these frameworks are no longer supported and were developed for some particular scientific projects.
+Therefore, a Monte Carlo simulation framework can be hardly applied for a bioinformatics project (say) based on the other algorithms.
+Moreover, processor technologies are evolving very rapidly.
+Many new revolutionary architectures have been developed for Sony PlayStation 4, Mali GPU by ARM, Adapteva Multicore Microprocessor etc.
+Both Mali GPU and Adapteva will be compatible with OpenCL.
+Xeon Phi co-processor works with OpenCL too, which was launched by INTEL.
+It has an x86-like architecture and is a 60 core co-processor that supports the PCI Express standard.
+It consumes just 300watts of energy and provides by giving a performance of 1 TFLOPS in double precision.
+Tianhe-2 which is the most powerful supercomputer till date implements this co-processor.
+Though, it is very difficult determine which Processor Computing framework architecture will more robust and high
+performance." \cite{hadoopgpu}
 
@@
 In our case, it is primarily a single data source, since it is based on the simulation, though in the future the simulation software systems might introduce distributed computing, too.
-It is also important to mention such projects as PeriScope\cite{PeriScope}, SciDB\cite{SciDB}, and other\cite{SDSS_SIGMOD02,QBISM_ICDE94}, since they target scientific data, which is a lot more different, than usual industry customer oriented data.
 
@@
 
-\chapter{Network of Queries}\label{sc:querynetwork}
+\chapter{Benchmark scientific queries}\label{sc:querynetwork}
 Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
 Often times, the queries that need to be run over the data look like selection and some kind of accumulation, if it is not more complicated.
-To be more exact, in this section we will try to summarize common set of queries we will be improving.
+To be more exact, in this section we will try to summarize common set of queries that were already introduced in the previous work of Push-Based System for Molecular Simulation Data Analysis \cite{mainPaper}.
 
@@
 Basically, these are the queries that need to be improved with our new approach of parallelism.
-Just to mention, $n$, $r_i$, $m_i$, $c_i$ and $q_i$ denote number of particles, coordinates (vector form), mass, charge, and number of electrons of a particle $i$.
+Respectively, $n$, $r_i$, $m_i$, $c_i$ and $q_i$ denote number of particles, coordinates (vector form), mass, charge, and number of electrons of a particle $i$.
 
 \begin{table}[h]
-\resizebox{\columnwidth}{!}{\begin{minipage}{\columnwidth}
+\centering
+\resizebox{\columnwidth}{!}
+{
+\begin{minipage}{\columnwidth}
+\centering
 \renewcommand*{\arraystretch}{1.5}
@@
 \hline %inserts horizontal line
-Function Name & Equation/Description \\[0.5ex] \hline
+Function Name & Equation/Description \\[0.5ex] 
+\hline
 Moment of Inertia & $\begin{array} {lcl} I & = & \sum\limits_{i=1}^n m_ir_i \end{array}$ \\[0.5ex]
@@
 \end{tabular}
-\caption[Table caption text]{Popular analytical queries in MS\cite{mainPaper} }
+\caption[Popular analytical queries in MS]{Popular analytical queries in MS }
 \label{tb:queries}
-\end{minipage} }
+
+\end{minipage}
+
+}
 \end{table}
 
+The queries that we consider are related to computations limited within one frame of a molecular simulation.
 In general, with the given queries we have divided them into two categories: one body functions and multiple body functions.
@@
 \emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
+
+As you can see in the given table, most of them are basically based on primitive summation or other basic mathematical operations, in our experiments in further sections we will be running result on single precision values or to be more explicit we use float type in C programming language.
 
@@
 
-The might be some dependencies and preliminary computation possibilities.
+For computation of histogram it is not enough to visit every value for each atom once, since we have to do pairwise computations, thus it increases complexity of the problem not only computation-wise, but memory-wise as well, which is a bigger concern in our case, since the overhead will be huge and non-linearly relative to the size of data.
+For this purpose we have come up with a way to optimize the way we use references to memory for each atom, so that we minimize overhead of pulling data all the way from DRAM (GPU).
+
+There might be some dependencies and preliminary computation possibilities.
 For example, some queries need total mass, and we could compute it before we push it further, but this can be easily done by user based on the complexity and dependency.
@@
 In the future, the same script for generating the mock data could be used for regression and performance tests in a distributed implementation of this design as a single source.
+Potentially input can be technically anything including simple IO sockets that could simply the way the system communicates with other components (f.e.
+over network) which would also contribute to expanding it to distributed cluster processing (though we believe that it would require additional components such as scheduler or job queues, etc.).
+To be more exact, the input format of the data and the data represents nothing but a simple array of tuples where each element represents certain descriptor of the given atom in the current frame.
+We believe that this simple approach makes it a good generalization of solution which would be user friendly for any scientist that wants to run any custom analysis on his data sets.
 
@@
 MS data roughly can be represented as a list of atoms with physical properties aggregated as frames.
-Thus, we consider a single frame as a chunk which will be loaded into the memory\footnote{Memory in this design might be interpreted differently, since we have a separate device memory on GPUs, but it will be explained better later} and pushed through our network of query.
+Thus, we consider a single frame as a chunk which will be loaded into the memory\footnote{Memory in this design might be interpreted differently, since we have a separate device memory on GPUs, but it will be explained better later} and pushed through our network of queries.
 In this case, it is very convenient, because there are almost no dependencies between frames, rather than list of atoms in the same frame, which potentially again could create additional disk I/O overhead.
@@
 
+To summarize, the data goes through several levels of memory hierarchy in our system.
+Coming as an input from either streaming tool, flat file (our case), or other type of input (f.e.
+socket) which is either disk memory or RAM it goes to RAM of our system process.
+After it is loaded to RAM, we continue further and load it to Global Memory of GPU, meanwhile while it is loading the given chunk of data to GPU Global Memory (which happens asynchronously), we as well load the next chunk of data to RAM.
+In this case, it helps us avoid minor overhead of transferring data between CPU to GPU.
+Further, we take use of different types of memories in GPU to optimize different queries which will be described in further sections.
+
 \section{Network of queries organization}
@@
 \emph{\textbf{GPU execution}}.
-From hardware standpoint, it is important to understand that every GPU device consists of SM (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
+From hardware standpoint, it is important to understand that every GPU device consists of SMs (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
 Having this in mind, since this is a pure parallel processing power, the goal is to increase hardware throughput and keep highest utilization to exploit every register given.
@@
 In CUDA we have grids, blocks, and threads, where grid is 3 dimensional unit where each cell is a block, and the same relationship applies between blocks and threads\footnote{This explanation doesn't include memory setup in GPU devices, but it is covered in further sections}.
+Abstract dimensions given by CUDA are not real entities of physical GPU memory and its organization, but it is just a very convenient way of organizing data and tasks to help GPU scheduler pick and execute the instructions.
 Originally, this mapping was perfect for solving image processing problems or other graphics computation, since dimensions and such division (e.g.
@@
  \centerline{ \includegraphics[width=0.65\columnwidth]{images/reduction} }
- \caption{Array reduction example\cite{gpureduction}}
+ \caption{GPU array reduction algorithm sample representation}
  \label{fg:reduction}
@@
 
-The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed C++.
+The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
 Experiments were run on operating systems which didn't have any additional workload besides usual installation, and the operating systems particularly were Amazon Linux, which doesn't have a lot of modifications except for some additional features for management with Amazon EC-2 Console.
@@
 
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \label{tab:cpu_hardware}
+  \begin{tabular}{| l | r |}
+   \hline
+Architecture & x86\_64 \\  \hline
+CPU op-mode(s) & 32-bit, 64-bit \\   \hline
+Byte Order & Little Endian \\   \hline
+CPU(s) & 8 \\   \hline
+On-line CPU(s) list & 0-7 \\   \hline
+Thread(s) per core & 2 \\   \hline
+Core(s) per socket & 4 \\   \hline
+CPU family & 6 \\   \hline
+Model & 45 \\   \hline
+Model name & Intel Xeon E5-2670 2.60GHz \\   \hline
+Stepping & 7 \\   \hline
+CPU MHz & 2593.814 \\  \hline  
+L1d cache & 32K \\   \hline
+L1i cache & 32K \\   \hline
+L2 cache & 256K \\   \hline
+L3 cache & 20480K \\   \hline
+NUMA node0 CPU(s) & 0-7 \\  \hline
+  \end{tabular}
+    \caption{Hardware stack for CPU version of Push-based system benchmark}
+\end{table}
 \emph{\textbf{Hardware stack.}}
@@
 
-\begin{center}
-  \begin{tabular}{ l  l }
-
-Architecture & x86\_64 \\  
-CPU op-mode(s) & 32-bit, 64-bit \\  
-Byte Order & Little Endian \\  
-CPU(s) & 8 \\  
-On-line CPU(s) list & 0-7 \\  
-Thread(s) per core & 2 \\  
-Core(s) per socket & 4 \\  
-CPU family & 6 \\  
-Model & 45 \\  
-Model name & Intel Xeon E5-2670 2.60GHz \\  
-Stepping & 7 \\  
-CPU MHz & 2593.814 \\  
-L1d cache & 32K \\  
-L1i cache & 32K \\  
-L2 cache & 256K \\  
-L3 cache & 20480K \\  
-NUMA node0 CPU(s) & 0-7 \\
-
-  \end{tabular}
-\end{center}
-
 The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
@@
 Consequently, we also have GPU device with the following parameters:
-
-\begin{center}
-  \begin{tabular}{ l l }
-Name & Grid 34C K520 \\
-Memory & 4095Mib \\ 
-Total Nvidia CUDA Cores & 1536 \\ 
-Memory Size & 8 GB GDDR5 (4 GB/GPU) \\
-Total GPUs & 2 GK104 GPUs \\
-  \end{tabular}
-\end{center}
 
@@
 This possibility will be elaborated in more details in Future Work section.
+
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \label{tab:gpu_hardware}
+  \begin{tabular}{| l | r |}
+   \hline
+	Name & Grid 34C K520 \\ \hline
+	Memory & 4095Mib \\ \hline
+	Total Nvidia CUDA Cores & 1536 \\ \hline
+	Memory Size & 8 GB GDDR5 (4 GB/GPU) \\ \hline
+	Total GPUs & 2 GK104 GPUs \\ \hline
+  \end{tabular}
+    \caption{Hardware stack for GPU version of Push-based system}
+\end{table}
 

### 2016-05-26 thesis_paper/source.tex
adding some stuff to thesis paper

@@
 \usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
-\usepackage{algorithmic,amsfonts,amssymb,amsmath,calc,calrsfs,cleveref}
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{amsfonts,amssymb,calc,calrsfs,cleveref}
 \usepackage[justification=centering]{caption}
 \usepackage{url}
+
+\makeatletter
+\def\BState{\State\hskip-\ALG@thistlm}
+\makeatother
 
@@
 Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-\subsection{One-body kernel} Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
+\subsection{One-body kernel}
+
+As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
+Our implementation of one-body functions are relatively easy to understand.
+
+\subsubsection{One-body functions implementation in CPU version}
+
+Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
+Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.
+
+\vspace{1cm}
+
+\begin{algorithm}
+\caption{One body functions implementation in CPU version}
+\label{alg:onebodycpu}
+\begin{algorithmic}[1]
+\Procedure{OneBodyFuncsCalc}{atoms, atomsCount}
+\State \Comment{ $atoms$ is vector of atoms}
+\State \Comment{ $atomsCount$ is number of atoms in the frame}
+
+\State $\textit{results} \gets \textit{default initial results value}$
+
+\BState \emph{top}:
+\State $\textit{i} \gets \textit{0}$
+
+\While{$i\not=atomsCount$} \Comment{Loop through all the atoms}
+
+\State{$\textit{currentAtom} \gets \textit{atoms[i]}$}
+\State{$\textit{calcTotalMass(results, currentAtom)}$}
+\State{$\textit{calcTotalCharge(results, currentAtom)}$}
+\State{$\textit{calcInertiaX(results, currentAtom)}$}
+\State{$\textit{calcInertiaY(results, currentAtom)}$}
+\State{$\textit{calcInertiaZ(results, currentAtom)}$}
+\State{$\textit{calcDepoleMoment(results, currentAtom)}$}
+
+\State{$\textit{//Here we are adding any other one-body functions}$}
+
+%increment
+
+\State $\textit{i} \gets \textit{i + 1}$ \Comment{Increment i}
+
+\EndWhile
+
+\State \textbf{return} $results$ \Comment{Return result for the frame}
+\EndProcedure
+
+\Procedure{calcTotalMass}{results, currentAtom} \Comment{Implementation of total mass function}
+	\State $results \gets \textit{results.totalmass + currentAtom.mass}$ 
+\EndProcedure
+
+\Comment{All other functions}
+
+\end{algorithmic}
+\end{algorithm}
+
+\vspace{1cm}
+
+On Algorithm Pseudocode~\ref{alg:onebodycpu} you can see more detailed explanation of CPU version approach for calculating one-body functions.
+
+\subsubsection{One-body functions implementation in GPU version}
+
+Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
 Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g.
@@
 First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
+
 \subsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.

### 2016-05-29 thesis_paper/source.tex
paper stuff

@@
 
-\ section{Computation work flow}
+\section{Computation work flow}
 
@@
 Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-\subsection{One-body kernel}
+
+\subsection{One-body functions implementation}
 
 As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
-Our implementation of one-body functions are relatively easy to understand.
-
-\subsubsection{One-body functions implementation in CPU version}
-
-Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
-Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.
-
-\vspace{1cm}
+Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.
 
@@
 \State $\textit{i} \gets \textit{0}$
-
+\State
 \While{$i\not=atomsCount$} \Comment{Loop through all the atoms}
@@
 \State{$\textit{currentAtom} \gets \textit{atoms[i]}$}
+\State
 \State{$\textit{calcTotalMass(results, currentAtom)}$}
@@
 
-\State{$\textit{//Here we are adding any other one-body functions}$}
+\State{$\textit{// Here we are adding any other one-body functions}$}
 
-%increment
-
+\State
 \State $\textit{i} \gets \textit{i + 1}$ \Comment{Increment i}
@@
 
+\State
 \State \textbf{return} $results$ \Comment{Return result for the frame}
@@
 
+\State
 \Procedure{calcTotalMass}{results, currentAtom} \Comment{Implementation of total mass function}
@@
 
-\vspace{1cm}
+\subsubsection{One-body functions implementation in CPU version}
 
-On Algorithm Pseudocode~\ref{alg:onebodycpu} you can see more detailed explanation of CPU version approach for calculating one-body functions.
+Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
+Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.
+
+We have provided Pseudocode \ref{alg:onebodycpu} of the implementation of one body functions for CPU, which has detailed step-by-step description of the CPU sequential version of push-based system for calculating one-body functions.
 
@@
 
-Every thread of execution in one-body query writes into shared memory, before accumulating the data into global memory result in order to avoid massive read and write blocking conflicts which will eliminate any kind of performance advantages.
-Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (e.g.
-generic CUDA function atomicAdd).
-We run the first stream with one-body queries and wait till it gets finished.
-The reason is that we expect certain behavior from shared memory from each of the streams, thus, just to make sure that they don't overlap, we run them separately.
-While the computation is running, there is a possibility to use a third stream to load another chunk of data, too.
-First stream and kernel execute simple vector reduction type of computation accumulating all the needed data and keep it in the GPU.
+The approach in GPU version of one-body computation code is a lot different, because there are two major changes: the way we deal with the data in memory and mapping of GPU thread kernels to atoms.
+Both of these approaches drastically change the way we perceive the solution implementation.
+Basically in the kernel function we assume we've pre-loaded all the needed atoms for current calculations (loading into memory and higher level parts will be described in later sections).
 
-\subsection{Two-body kernel} Once the first kernel is finished we run the two-body kernel with its own stream.
+As you can see from the given algorithm (Pseudocode \ref{alg:onebodygpu}), in the thread kernel function it starts with allocation of memory in the block.
+To be more explicit, every block has its own level of memory called shared memory.
+To understand the importance of shared memory, it is on-chip memory which can get latency roughly 100x lower than global memory.
+Further down, we calculate the current index of the atom we need to work with, and this computation takes two lines of code, because we have a grid-based dimensions that have to be flattened back to one-dimension array indexing which was explained in previous sections.
+Once we are setup, we start accumulating results into the shared memory atomically, because threads within same block might have race condition on write operation.
+This is one of the reasons why we are doing it locally, because it would also take much more time and overhead to have locks in global memory that would increase race conditions and locking times.
+After all, we synchronize all threads to make sure that all the records from atoms have been collected to appropriate local results in shared memory.
+Finally, we collect all the results from blocks' shared memory into global memory in which case it will be much faster, because number of block is not that big.
+Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (we use native CUDA function atomicAdd which guarantees atomicity).
+
+\vspace{1cm}
+
+\begin{algorithm}
+\caption{Kernel function implementation for one-body queries in GPU}
+\label{alg:onebodygpu}
+\begin{algorithmic}[1]
+\Procedure{gpuonebodyfunctionskernel}{atomsCount, atoms, results}
+\State \Comment{ $atoms$ is vector of atoms kept in global gpu memory}
+\State \Comment{ $atomsCount$ is number of atoms in the frame kept in global gpu memory}
+\State \Comment{ $results$ is number of atoms in the frame kept in global gpu memory}
+
+\BState \emph{top}:
+
+\State $\textit{sdata} \gets \textit{allocate shared memory in each GPU block for fast write access}$ 
+\State \Comment{We will use this further to accumulate data across blocks}
+
+\State $\textit{indexX} \gets \textit{blockIdsx.x * blockDim.x + threadIdx.x}$ 
+\State \Comment{Identify position of the kernel in the grid scale for coordinate X}
+\State $\textit{indexY} \gets \textit{blockIdsx.y * blockDim.y + threadIdx.y}$ 
+\State \Comment{Identify position of the kernel in the grid scale for coordinate Y}
+
+\State $\textit{gridWidth} \gets \textit{gridDim.x * blockDim.x}$
+\State \Comment{Calculate grid size in kernel units}
+\State $\textit{i} \gets \textit{indexY * gridWidth + indexX}$
+\State \Comment{Finally get the index of current atom mapped to this kernel}
+
+\State $\textit{sharedResult} \gets \textit{pointer to shared memory head (in this case declared element size as float)}$
+
+\State $\textit{currentAtom} \gets \textit{atoms[i]}$
+\State 
+\State $\textit{atomicAdd(\&sharedResult[0], currentAtom.mass);}$
+\State 
+\State $\textit{atomicAdd(\&sharedResult[1], currentAtom.charge);}$
+\State $\textit{atomicAdd(\&sharedResult[2], currentAtom.mass * currentAtom.x);}$
+\State $\textit{atomicAdd(\&sharedResult[3], currentAtom.mass * currentAtom.y);}$
+\State $\textit{atomicAdd(\&sharedResult[3], currentAtom.mass * currentAtom.z);}$
+\State $\textit{// other function steps...}$
+
+\State
+
+\State$\textit{// make sure all kernel threads get here and wait for each other}$
+\State $\textit{\_\_syncthreads();}$
+
+\State
+\State $\textit{// Here we will starting collecting information from each block which will be done by first thread of each block}$
+\If {$\textit{threadIdx.x} = 0$}
+\State  $\textit{atomicAdd(\&results->mass, sharedResult[0]);}$ \Comment{All atomicAdd functions here are GPU atomic}
+\State  $\textit{atomicAdd(\&results->charge, sharedResult[1]);}$
+\State  $\textit{atomicAdd(\&results->inertiaX, sharedResult[2]);}$
+\State  $\textit{atomicAdd(\&results->inertiaY, sharedResult[3]);}$
+\State  $\textit{atomicAdd(\&results->inertiaZ, sharedResult[4]);}$
+\State  $\textit{// other functions to accumulate from block shared memory...}$
+\EndIf
+
+\State
+
+\EndProcedure
+
+\end{algorithmic}
+\end{algorithm}
+
+\vspace{1cm}
+
+\subsection{Two-body functions implementation}
+
+Since one of the physical queries is RDF which has underlying SDH calculation, we will describe our approach with SDH for CPU and GPU versions.
+
+\subsubsection{Two-body functions implementation in CPU Version}
+
+Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
+Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them, though results don't change a lot for sequential computation.
+
+\begin{algorithm}
+\caption{SDH in CPU version}
+\label{alg:sdhcpu}
+\begin{algorithmic}[1]
+\Procedure{SDHCPU}{atoms, atomsCount, histogram}
+
+\State $\textit{i} \gets 0$
+\While{i < atomsCount}
+
+\State $\textit{j} \gets i + 1$
+\While{j < atomsCount}
+\State $\textit{dist} \gets calcDistance(atoms[i], atoms[j])$
+\State $\textit{Save dist in histogram bucket}$
+\State
+\State $\textit{j} \gets j + 1$
+\EndWhile
+
+\State $\textit{i} \gets i + 1$
+\EndWhile
+
+\EndProcedure
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Two-body functions implementation in GPU Version} Once the first kernel is finished we run the two-body kernel with its own stream.
 As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity.

### 2016-05-29 thesis_paper/source.tex
paper stuff
paper

@@
 
-\vspace{1cm}
-
 \subsection{Two-body functions implementation}
@@
 
-\subsubsection{Two-body functions implementation in GPU Version} Once the first kernel is finished we run the two-body kernel with its own stream.
-As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity.
-In this particular case it is $O(n^2)$.
+\subsubsection{Two-body functions implementation in GPU Version}
+
+The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same.
+We do brute force  $O(\frac{N^2}{2})$ complex solution which is pretty much looking at distance between every pair of atoms.
+In Pseudocode \ref{alg:sdhgpu} we have tried to provide main logic flow that explains how exactly we implemented kernel for SDH computation.
+Once the first kernel is finished we run the two-body kernel with its own stream.
+As it has already been mentioned, the bigger workload is in this kernel, because it has higher complexity.
 GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence.
@@
 
-For SDH we the following strategy
-
+Here are the main steps of the algorithm:
 \begin{itemize}
-\item[1.] Find current index (for target atom)
-\item[2.] Find range of atoms it has to interact with in an equally loaded fashion
-\item[3.] Load the range of atoms into shared memory
-\item[4.] Iterate through the range of atoms from shared memory and atomically write the distances counts into buckets pre-allocated in share memory
-\item[5.] Synchronize with other threads in the block (not in the entire grid)
-\item[6.] If it is the first thread in block, add all bucket counts from shared memory into global memory
+\item[1.] Get mapped atom
+\item[2.] Load all block related atoms to shared memory
+\item[3.] Go pairwise with atoms for SDH in a equal balance manner
+\item[4.] Synchronize threads
+\item[5.] Get everything block shared memory to global memory
 \end{itemize}
+
+\begin{figure}
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/sharedMem}}
+ \caption{Representation of example of shared memory usage}
+ \label{fg:sharedMem}
+\end{figure}
+
+As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that user's wouldn't do it and having more or less realistic speedup results.
+Figure \ref{fg:sharedMem} shows swimlanes as execution thread for blocks.
+Every row is an execution block for an atom, where first element is the mapped atom.
+First of all, you can observe that we can clearly see that the execution threads are almost equally balanced (exceptions on odd number of atoms).
+Every swimlane (bordered with black) represent within block execution.
+Colored numbers are the atoms that we loaded into shared memory.
+Of course, this is about half of improvement, and we can happily add improvement of adding second part too, but, at this point, we go with this average improvement approach.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%      SDH GPU      %%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{algorithm}
+\caption{Kernel function implementation for SDH in GPU}
+\label{alg:sdhgpu}
+\begin{algorithmic}[1]
+\Procedure{gpusdhkernel}{atomsCount, pdhRes, histogram, atoms, numOfBuckets}
+\State \Comment{ $atoms$ is vector of atoms kept in global gpu memory}
+\State \Comment{ $atomsCount$ is number of atoms in the frame kept in global gpu memory}
+\State \Comment{ $histogram$ is buckets with counters for the SDH}
+\State \Comment{ $pdhRes$ is resolution of histogram}
+\State \Comment{ $numOfBuckets$ is number of buckets}
+
+\BState \emph{top}:
+
+\State $\textit{smem} \gets \textit{allocate shared memory in each GPU block for fast write access of histogram buckets}$
+
+\State $\textit{sharedHisto} \gets \textit{pointer to the head of smem}$
+\State $\textit{sharedAtoms} \gets \textit{pointer to the tail of sharedHisto}$
+\State $\textit{indexX and indexY} \gets \textit{calculate from grid sizes}$ 
+\State $\textit{index} \gets \textit{get atom mapped to this kernel}$
+\State \Comment{Finally get the index of current atom mapped to this kernel}
+
+\If {$\textit{threadIdx.x} = 0$}
+
+\State $\textit{sharedHisto} \gets \textit{Set all values to 0}$
+
+\State $\textit{start} \gets \textit{index}$
+\State $\textit{k} \gets \textit{0}$
+\State $\textit{i} \gets \textit{index}$
+\For {$i < start + blockDim.x$ and $i < atomsCount$}
+
+	\State $\textit{sharedAtoms[k]} \gets \textit{atoms[i]}$
+	\Comment{Load local atoms into shared memory}
+
+	\State $\textit{k} \gets \textit{k + 1}$
+	\State $\textit{i} \gets \textit{i + 1}$
+\EndFor
+
+\EndIf
+
+\State
+\State $\textit{\_\_syncthreads()}$
+\State $\textit{threadLoad} = \textit{(atomsCount + 1) / 2}$
+\State $\textit{start} \gets \textit{i+1}$
+\State $\textit{end} \gets \textit{i + threadLoad}$
+
+\State $\textit{end} \gets \textit{increment by one if it is odd and total count is even}$
+
+\State $\textit{bi} \gets \textit{blockDim.x * blockIdx.x}$
+\State $\textit{ei} \gets \textit{bi + blockDim.x}$
+
+\State $\textit{ind1} \gets \textit{threadIdx.x}$
+
+\State $\textit{j} \gets \textit{start}$
+
+\For {$\textit{j} < \textit{end}$}
+
+\State $\textit{ind2} \gets \textit{j mod atomsCount}$
+\State $\textit{atom1} \gets \textit{load from shared memory}$
+
+\If{ $\textit{atom2 can be loaded from shared memory}$}
+\State $\textit{atom2} \gets \textit{load from local memory}$
+\Else
+\State $\textit{atom2} \gets \textit{load from global memory}$
+\EndIf
+
+\State $\textit{calc distance and atomically write to local histogram}$
+
+\EndFor
+\State
+\State $\textit{synchronize all threads}$
+\State $\textit{gather all histogram from shared into global memory}$
+\EndProcedure
+
+\end{algorithmic}
+\end{algorithm}
 

### 2016-05-30 thesis_paper/source.tex
thesis

@@
 \usepackage{algorithm}
-\usepackage{algorithmicx}
 \usepackage{algpseudocode}
@@
 
+\vspace{1cm}
+
 \begin{algorithm}
@@
 \item[4.] Synchronize threads
-\item[5.] Get everything block shared memory to global memory
+\item[5.] Get everything for results from shared memory to global memory
 \end{itemize}
@@
 Every row is an execution block for an atom, where first element is the mapped atom.
-First of all, you can observe that we can clearly see that the execution threads are almost equally balanced (exceptions on odd number of atoms).
+First of all, you can observe that we can clearly see that the execution threads are almost equally balanced (exceptions on odd number of atoms) which means that there is no code divergence and low utilization of kernels (no awaiting redundant threads).
+In this particular example, we have block size 5 with total number of atoms to be 15 (of course, these are just simple examples, realistically it would be millions).
 Every swimlane (bordered with black) represent within block execution.
-Colored numbers are the atoms that we loaded into shared memory.
-Of course, this is about half of improvement, and we can happily add improvement of adding second part too, but, at this point, we go with this average improvement approach.
-
+Secondly, colored numbers are the atoms that we loaded into shared memory.
+Of course, this is a small improvement, and we can happily add improvement of adding more shared memory and loading as it moves synchronizing threads in the future.
+
+\vspace{1cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \Procedure{gpusdhkernel}{atomsCount, pdhRes, histogram, atoms, numOfBuckets}
-\State \Comment{ $atoms$ is vector of atoms kept in global gpu memory}
-\State \Comment{ $atomsCount$ is number of atoms in the frame kept in global gpu memory}
-\State \Comment{ $histogram$ is buckets with counters for the SDH}
-\State \Comment{ $pdhRes$ is resolution of histogram}
-\State \Comment{ $numOfBuckets$ is number of buckets}
-
-\BState \emph{top}:
-
-\State $\textit{smem} \gets \textit{allocate shared memory in each GPU block for fast write access of histogram buckets}$
+
+\State $\textit{smem} \gets \textit{pointer to head of SM}$
 
@@
 \State $\textit{sharedAtoms} \gets \textit{pointer to the tail of sharedHisto}$
-\State $\textit{indexX and indexY} \gets \textit{calculate from grid sizes}$ 
-\State $\textit{index} \gets \textit{get atom mapped to this kernel}$
-\State \Comment{Finally get the index of current atom mapped to this kernel}
-
+\State $\textit{index} \gets \textit{calculate index of kernel mapped to this atom}$
+
+\State
 \If {$\textit{threadIdx.x} = 0$}
@@
 \State $\textit{sharedHisto} \gets \textit{Set all values to 0}$
-
-\State $\textit{start} \gets \textit{index}$
-\State $\textit{k} \gets \textit{0}$
-\State $\textit{i} \gets \textit{index}$
+\State $\textit{start} \gets \textit{index, }
+			 \textit{i} \gets \textit{index, }
+			 \textit{k} \gets \textit{0,}$
+
 \For {$i < start + blockDim.x$ and $i < atomsCount$}
-
 	\State $\textit{sharedAtoms[k]} \gets \textit{atoms[i]}$
 	\Comment{Load local atoms into shared memory}
-
-	\State $\textit{k} \gets \textit{k + 1}$
-	\State $\textit{i} \gets \textit{i + 1}$
+	\State $\textit{i} \gets \textit{i + 1, } \textit{k} \gets \textit{k + 1}$
 \EndFor
@@
 \State $\textit{start} \gets \textit{i+1}$
-\State $\textit{end} \gets \textit{i + threadLoad}$
-
-\State $\textit{end} \gets \textit{increment by one if it is odd and total count is even}$
-
-\State $\textit{bi} \gets \textit{blockDim.x * blockIdx.x}$
-\State $\textit{ei} \gets \textit{bi + blockDim.x}$
+\State $\textit{end} \gets \textit{i + threadLoad, }
+\textit{increment end by one if it is odd and total count is even}$
+
+\State $\textit{bi} \gets \textit{blockDim.x * blockIdx.x, } \textit{ei} \gets \textit{bi + blockDim.x}$
 
@@
 \State $\textit{calc distance and atomically write to local histogram}$
-
 \EndFor
+
 \State
@@
 \end{algorithm}
+
+\subsection{Data flow implementation in CPU}
+
+As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
+
+\vspace{1cm}
+%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%      CPU FLOW %%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{algorithm}
+\caption{Main algorithm in CPU for managing data and pushing to GPU}
+\label{alg:cpuflow}
+\begin{algorithmic}[1]
+\Procedure{runsinglekernel}{atomsCount, atoms, workload}
+
+\State $\textit{stream1} \gets \textit{create CUDA stream with cudaStreamCreate}$
+\State $\textit{stream2} \gets \textit{create CUDA stream with cudaStreamCreate}$
+\State
+\State $\textit{gpuResults} \gets \textit{allocate gpu global memory}$
+\State $\textit{gpuAtoms} \gets \textit{allocate gpu global memory}$
+\State $\textit{gpuHistogram} \gets \textit{allocate gpu global memory}$
+\State
+\State $\textit{gpuAtoms} \gets \textit{copy atoms from CPU to GPU with cudaMemcpy}$
+
+\State $\textit{gridSize} \gets \textit{compute grid size depending on atom count}$
+
+\State
+
+\For {$workload$}
+
+\State $\textit{gpukernel1(stream1, sharedMemorySize, gpuAtoms, ...)}$
+\Comment{One body queries kernel}
+\State $\textit{gpukernel2(stream2, sharedMemorySize, gpuAtoms, ...)}$
+\Comment{Two body queries kernel}
+
+\State
+\State $\textit{cudaStreamSynchronize(stream1)}$
+\State $\textit{cudaStreamSynchronize(stream2)}$
+
+\EndFor
+
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\vspace{1cm}
+
+In the Pseudocode \ref{alg:cpuflow} you can see rough logic of our CPU mapping function that primarily allocates all the memory in GPU, triggers kernel function and manipulates streams.
+This is a prototype which definitely is not the last version, because we can as well avoid overhead of copying data into shared memory while synchronously it will be computing data already contained there.
+The main steps here are the following.
+we create to instances of CUDA streams that are again abstract data types that are scheduled by GPU in a special way.
+We as well allocate memory in GPU for atoms list, histogram results and common results for one-body queries (in this case it is a data structure with fields that represent some accumulated result).
+Once we are setup with memory management on GPU, we trigger kernel calls one by one each running on a separate CUDA stream.
+We don't expect them to run really concurrently, because our two kernels are quite heavily loaded, though streaming can definitely be use for memory management optimization.
+
+\vspace{1cm}
+%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%    CPU FLOW %%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%
+\begin{algorithm}
+\caption{Main algorithm in CPU for managing data and pushing to GPU}
+\label{alg:cpuflowfuture}
+\begin{algorithmic}[1]
+\Procedure{runsinglekernel}{atomsCount, atoms, nextAtoms, workload}
+
+\State $\textit{stream1, stream2} \gets \textit{create CUDA stream with cudaStreamCreate}$
+
+\State
+\State $\textit{The same routing with allocating memory for now for nextAtoms too}$
+
+\State
+\State $\textit{Here we start loading nextAtoms to GPU on stream2 }$
+\State
+
+\For {$workload$}
+
+\State $\textit{gpukernel1(stream1, sharedMemorySize, gpuAtoms, ...)}$
+\Comment{One body queries kernel}
+\State $\textit{gpukernel2(stream1, sharedMemorySize, gpuAtoms, ...)}$
+\Comment{Two body queries kernel}
+
+\State
+\State $\textit{cudaStreamSynchronize(stream1)}$
+\State $\textit{cudaStreamSynchronize(stream2)}$
+
+\EndFor
+
+\EndProcedure
+\end{algorithmic}
+\end{algorithm}
+
+\vspace{1cm}
+
+The above Pseudocode \ref{alg:cpuflowfuture} shows what memory concurrent could look like.
+As you can see it is pretty much overlapping $nextAtoms$ loading with computation of $atoms$, which reduces this tiny overhead of copy data over.
 

### 2016-06-01 thesis_paper/source.tex
adding some experiment results

@@
 \usepackage{url}
+\usepackage{subfigure}
+\usepackage{cleveref}
 
@@
 
+\section{Hardware stack}
+
+There are primarily several parts that are really important in this hardware stack.
+Since this is primarily big data processing, we care about processing power and memory, Of course, things like PCI bus and all the transportation are important as well, but for speedup benchmarks against CPU vs GPU those stay the same for both sides.
+RAM on the node that we were running has 16GBs.
+
+\subsection{CPU}
+We have summarized main CPU parameters for CPU version and it is described in Table \ref{tb:cpuhardware}. As you may notice, it is a regular server node CPU with 2.60GHz and 3 levels of L cache.
+
 \begin{table}[h!]
@@
 	\centering
-  \label{tab:cpu_hardware}
   \begin{tabular}{| l | r |}
@@
     \caption{Hardware stack for CPU version of Push-based system benchmark}
+      \label{tb:cpuhardware}
 \end{table}
-\emph{\textbf{Hardware stack.}}
-There are primarily several parts that are really important in this hardware stack.
-Since this is primarily big data processing, we care about processing power and memory.
-RAM on the node that we were running has 16GB.
-The server allocated from Amazon has the following CPU parameters:
+
+\subsection{GPU}
 
 The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
-Just to mentioned, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load.
-Consequently, we also have GPU device with the following parameters:
-
-\emph{\textbf{Data sets.}} As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
-It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format.
-We have generated number of mock data sets in range of $1K-8M$ atoms with $10, 100, 1000$ frames each.
-For each data set we run workloads with the values: $1, 5, 25, 50, 100$ .
-Number of atoms and frames are obviously up to the scientist and depend on simulations, but workloads are a bit trickier.
-First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests.
-This possibility will be elaborated in more details in Future Work section.
+So, for CPU in GPU improved version refer to the same Table \ref{tb:cpuhardware}.
+For recall, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load.
+Consequently, we also have GPU device.
+Hardware summary for GPU device is described in Table \ref{tb:gpuhardware}
 
@@
   \centering
-  \label{tab:gpu_hardware}
   \begin{tabular}{| l | r |}
@@
     \caption{Hardware stack for GPU version of Push-based system}
+    	  \label{tb:gpuhardware}
 \end{table}
+\section{Data Sets}
+
+As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
+It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format.
+We have generated number of mock data sets in range of $1K-8M$ atoms with $10, 100, 1000$ frames each.
+For each data set we run workloads with the values: $1, 5, 25, 50, 100$ .
+Number of atoms and frames are obviously up to the scientist and depend on simulations, but workloads are a bit trickier.
+First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests.
+This possibility will be elaborated in more details in Future Work section.
 
@@
 \section{Results}
-Results are represented, in general, on Figures \ref{fg:chart_1000}, \ref{fg:chart_10000} \ref{fg:chart_max}.
+
+We have run number of experiments with different number of atoms and frames.
+Results vary as we choose different data sets, but we can see some tendency towards growth and, fortunately, it is very positive for our improved version of push-based system.
+
+\subsection{Experiment results for 1000 atoms}
+
+Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes.
+You can observe summary in Figures \ref{fg:exp_1000_10}, \ref{fg:exp_1000_100}, \ref{fg:exp_1000_1000}, \ref{fg:speedup_1000}.
+There is a minor threshold, where CPU version is faster, where workload is less than 20 on Figure \ref{fg:exp_1000_10}, but it is changed with growth of workload.
+That presumably is the minor overhead of loading data from CPU to GPU and number of kernels versus computation work actually needs to be done.
+Once there is more load on computation, strongly GPU accelerates and becomes faster.
+
+\begin{figure}[!tbp]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_1000_10}
+    \caption{Time for 1000 atoms and 10 frames}
+    \label{fg:exp_1000_10}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_1000_100}
+    \caption{Time for 1000 atoms and 100 frames}
+     \label{fg:exp_1000_100}
+  \end{minipage}
+\hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_1000_1000}
+    \caption{Time for 1000 atoms and 1000 frames}
+     \label{fg:exp_1000_1000}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/speedup_1000}
+    \caption{Speedup for 1000 atoms}
+     \label{fg:speedup_1000}
+  \end{minipage}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Experiment results for 10,000 atoms}
+
+\begin{figure}[!tbp]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_10000_10}
+    \caption{Time for 10000 atoms and 10 frames}
+    \label{fg:exp_10000_10}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_10000_100}
+    \caption{Time for 10000 atoms and 100 frames}
+     \label{fg:exp_10000_100}
+  \end{minipage}
+\hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_10000_1000}
+    \caption{Time for 10000 atoms and 1000 frames}
+     \label{fg:exp_10000_1000}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/speedup_10000}
+    \caption{Speedup for 10000 atoms}
+     \label{fg:speedup_10000}
+  \end{minipage}
+\end{figure}
+
 You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
@@
 
-\emph{\textbf{10000 atoms.}}
 As you can see, speedup bumps up to almost 30 with 10000 atoms (Figure \ref{fg:chart_10000}).
@@
 
-\begin{figure}
- \centerline{ \includegraphics[width=0.5\columnwidth]{images/chart_1000}}
- \caption{Speedup for 1000 atoms}
- \label{fg:chart_1000}
-\end{figure}
-
-\begin{figure}
- \centerline{\includegraphics[width=0.5\columnwidth]{images/chart_10000}}
- \caption{Speedup for 10000 atoms}
- \label{fg:chart_10000}
+\subsection{Experiment results for 100,000 atoms}
+
+\begin{figure}[!tbp]
+  \centering
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_100000_10}
+    \caption{Time for 100,000 atoms and 10 frames}
+    \label{fg:exp_100000_10}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/exp_100000_100}
+    \caption{Time for 100,000 atoms and 100 frames}
+     \label{fg:exp_100000_100}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.4\textwidth}
+    \includegraphics[width=\textwidth]{images/speedup_100000}
+    \caption{Speedup for 100,000 atoms}
+     \label{fg:speedup_100000}
+  \end{minipage}
 \end{figure}

### 2016-06-03 thesis_paper/source.tex
thesis

@@
 
+\vspace{1cm}
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+	Device & Context & Stream & Kernel & Warps Launched & Shared Load \\ \hline
+	GRID K520 (0) & 1 & 13 & ${gpu\_one\_body\_function}$ & 32 & 2 \\ \hline
+	GRID K520 (0) & 1 & 14 & ${gpu\_two\_body\_function}$ & 32 & 96080 \\ \hline
+
+  \end{tabular}
+    \caption{GPU profiling results for 1000 atoms}
+    	  \label{tb:gpuprofile1000}
+\end{table}
+
+On Table \ref{tb:gpuprofile1000} you can observe results of Nvidia Profiling Tool of how much memory call we did to shared memory for SDH computation.
+You can also notice that there were 32 warps launched for computing 1000 atoms which is absolutely reasonable, since as we already mentioned one warp (scheduling unit for ultimate parallel execution) has 32 threads maximum, thus ${\frac{1000}{32}}$ \textasciitilde \space$32$.
+As expected, stream IDs differ for two kernel executions, and both were run on the same GPU device, though we assume that our system can potentially be extended to using multiple GPU devices as well.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-As you can see, speedup bumps up to almost 30 with 10000 atoms (Figure \ref{fg:chart_10000}).
+As you can see, speedup bumps up to almost 30 with 10000 atoms.
 Once the overhead on memory transfer is covered by computation speedup, it keeps growing.
 It is important to pay attention to the growth of speedup over growth of workload.
-It is not that significant, because even though CPUs are much slower, they still use smart caching and internal multi-core processing, which is pretty fast, too.
+It is not that significant, because even though CPUs are much slower, they still use smart caching and internal multi-core processing, which is pretty fast, too, while GPU does pretty straightforward multicore processing, which is the main reason it stagnates.
 
@@
 
+When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good. Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
+
 Fortunately, scientific data is much bigger than above examples.

### 2016-06-05 thesis_paper/source.tex
thesis

@@
 
-\section{Network of queries organization}
+\section{Queries organization}
 
@@
 Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.
+
+\vspace{1cm}
 
@@
 
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+
+$Duration$ & $Grid Size$ & $Block Size$ & $DSMem*$ & $Size$ & $Throughput$ \\ \hline 
+$1.1830us$ & $-$ & $-$ & $-$ & $4B$ & $3.3812MB/s$ \\ \hline 
+$6.0470us$ & $-$ & $-$ & $-$ & $36.000KB$ & $5.9534GB/s$ \\ \hline 
+$1.0240us$ & $-$ & $-$ & $-$ & $36B$ & $35.156MB/s$ \\ \hline 
+$1.2160us$ & $-$ & $-$ & $-$ & $1.4080KB$ & $1.1579GB/s$ \\ \hline 
+$234.16us$ & $(1,1,1)$ & $(1024,1,1)$ & $40.960KB$ & $-$ & $-$ \\ \hline 
+$3.1640ms$ & $(1,1,1)$ & $(1024,1,1)$ & $38.272KB$ & $-$ & $-$ \\ \hline 
+$2.5280us$ & $-$ & $-$ & $-$ & $36B$ & $14.241MB/s$ \\ \hline 
+$2.4960us$ & $-$ & $-$ & $-$ & $1.4080KB$ & $564.10MB/s$ \\
+	\hline
+
+  \end{tabular}
+    \caption{Memory profiling results for 1000 atoms}
+    	  \label{tb:gpumem1000}
+\end{table}
+
 On Table \ref{tb:gpuprofile1000} you can observe results of Nvidia Profiling Tool of how much memory call we did to shared memory for SDH computation.
@@
 
+\vspace{1cm}
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+	Device & Context & Stream & Kernel & Warps Launched & Shared Load \\ \hline
+	GRID K520 (0) & 1 & 13 & ${gpu\_one\_body\_function}$ & 320 & 20 \\ \hline
+	GRID K520 (0) & 1 & 14 & ${gpu\_two\_body\_function}$ & 320 & 5,181,545 \\ \hline
+
+  \end{tabular}
+    \caption{GPU profiling results for 10,000 atoms}
+    	  \label{tb:gpuprofile10000}
+\end{table}
+
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+
+$Duration $ & $ Grid Size $ & $ Block Size $ & $ DSMem* $ & $ Size $ & $ Throughput  $ \\ \hline
+$1.1840us $ & $ - $ & $ - $ & $ - $ & $ 4B $ & $ 3.3784MB/s  $ \\ \hline
+$42.470us $ & $ - $ & $ - $ & $ - $ & $ 360.00KB $ & $ 8.4766GB/s  $ \\ \hline
+$1.0240us $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 35.156MB/s  $ \\ \hline
+$1.2480us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 1.1282GB/s  $ \\ \hline
+$456.86us $ & $ (10,1,1) $ & $ (1024,1,1) $ & $ 40.960KB $ & $ - $ & $ -  $ \\ \hline
+$59.627ms $ & $ (10,1,1) $ & $ (1024,1,1) $ & $ 38.272KB $ & $ - $ & $ -  $ \\ \hline
+$2.7520us $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 13.081MB/s  $ \\ \hline
+$2.7850us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 505.57MB/s$ \\
+
+	\hline
+
+  \end{tabular}
+    \caption{Memory profiling results for 10,000 atoms}
+    	  \label{tb:gpumem10000}
+\end{table}
+
 As you can see, speedup bumps up to almost 30 with 10000 atoms.
@@
 It is not that significant, because even though CPUs are much slower, they still use smart caching and internal multi-core processing, which is pretty fast, too, while GPU does pretty straightforward multicore processing, which is the main reason it stagnates.
+Result of another profiling on Table \ref{tb:gpuprofile100000} changed significantly in terms of warps launched, which is reasonable, and shared memory.
+It is also noticeable that growth of shared memory loading instructions is non-linear which you can observe in the results for 100K atoms.
 
 \subsection{Experiment results for 100,000 atoms}
+
+When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good. Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
 
@@
 
-When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good. Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+
+ $Duration $ & $ Grid Size $ & $ Block Size $ & $ DSMem* $ & $ Size $ & $ Throughput $ \\ \hline 
+ $1.1840us $ & $ - $ & $ - $ & $ - $ & $ 4B $ & $ 3.3784MB/s $ \\ \hline 
+ $729.81us $ & $ - $ & $ - $ & $ - $ & $ 3.6000MB $ & $ 4.9328GB/s $ \\ \hline 
+ $992ns $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 36.290MB/s $ \\ \hline 
+ $1.2160us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 1.1579GB/s $ \\ \hline 
+ $3.0014ms $ & $ (98,1,1) $ & $ (1024,1,1) $ & $ 40.960KB $ & $ - $ & $ - $ \\ \hline 
+ $3.86896s $ & $ (98,1,1) $ & $ (1024,1,1) $ & $ 38.272KB $ & $ - $ & $ - $ \\ \hline 
+ $3.0720us $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 11.719MB/s $ \\ \hline 
+ $2.9760us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 473.12MB/s $ \\ \hline
+
+	\hline
+  \end{tabular}
+    \caption{Memory profiling results for 100,000 atoms}
+    	  \label{tb:gpumem100000}
+\end{table}
 
@@
 
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+	Device & Context & Stream & Kernel & Warps Launched & Shared Load \\ \hline
+	GRID K520 (0) & 1 & 13 & ${gpu\_one\_body\_function}$ & 3136 & 196 \\ \hline
+	GRID K520 (0) & 1 & 14 & ${gpu\_two\_body\_function}$ & 3136 & 473,692,099 \\ \hline
+
+  \end{tabular}
+    \caption{GPU profiling results for 100,000 atoms}
+    	  \label{tb:gpuprofile100000}
+\end{table}
+
+Finally, for 100,000 atoms the warps launched for execution seems to be growing reasonable along with number of atoms (Table \ref{tb:gpuprofile100000}), while on the other hand usage of shared load for SDH computation is extremely high.
+
+\subsection{Overall performance}
+
+It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
+It is because we've used static approach of computing grid dimensions.
+Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
+
 \begin{figure}
- \centerline{\includegraphics[width=0.5\columnwidth]{images/chart_max}}
+ \centerline{\includegraphics[width=0.7\columnwidth]{images/chart_max}}
  \caption{Speedup for larger number of atoms}
@@
 
-It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
-It is because we've used static approach of computing grid dimensions.
-Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
-
 \chapter{Conclusions and Future Work}\label{sc:conclusion}
 The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
-As you could see from result on charts (Figure \ref{fg:chart_1000}, \ref{fg:chart_max}), the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time.
+As you could see from result on charts, the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time.
 The results have been gathered with a single snapshot of the framework tracked in control version system which can be improved with means of data converter module and dynamic grid dimensions computation module.

### 2016-06-05 gpu_single.cu
Trying to utilize shared memory in SDH

@@ -75,6 +75,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
     unsigned long long* shared_histo = smem;
     atom* sharedAtoms = (atom*) &shared_histo[num_buckets];
+    atom* sharedAtoms1 = (atom*) &sharedAtoms[blockDim.x];
 
     long index_x = blockIdx.x * blockDim.x + threadIdx.x;
     long index_y = blockIdx.y * blockDim.y + threadIdx.y;
@@ -101,6 +102,16 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         for(i = start; i < start + blockDim.x && i < PDH_acnt; i++, k++) {
             sharedAtoms[k] = at_list[i];
         }
+
+        k = 0;
+        for(i = start; i < start + blockDim.x * 2 && i < PDH_acnt; i++, k++) {
+
+            if(i < start + blockDim.x) {
+                sharedAtoms1[k] = sharedAtoms[k];
+            } else {
+                sharedAtoms1[k] = at_list[i];    
+            }
+        }
     }
 
     __syncthreads();
@@ -116,8 +127,8 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         end++;
     }
 
-    int bi = blockDim.x * blockIdx.x;   // block start
-    int ei = bi + blockDim.x;           // block end
+    int bi = blockDim.x * blockIdx.x;   // shared block start
+    int ei = bi + blockDim.x * 2;           // shared block end
 
     int ind1 = threadIdx.x;             // in this block from sharedAtoms
     int ind2;
@@ -132,16 +143,31 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         double z1 = sharedAtoms[ind1].z;
 
         double x2, y2, z2;
-        
-        if(bi <= ind2 && ind2 < ei) {
-            x2 = sharedAtoms[ind2 - bi].x;
-            y2 = sharedAtoms[ind2 - bi].y;
-            z2 = sharedAtoms[ind2 - bi].z;
-        } else {
-            x2 = at_list[ind2].x;
-            y2 = at_list[ind2].y;
-            z2 = at_list[ind2].z;
+
+        __syncthreads();
+
[…]

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -146,7 +146,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
         __syncthreads();
 
-        if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei) { //not finding in shared memory
+        if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
             bi += blockDim.x * 2;
             ei += blockDim.x * 2;
             for(i = bi - blockDim.x; i < ei; i++, k++) {
@@ -164,7 +164,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         z2 = sharedAtoms1[ind2 - bi].z;
 
         __syncthreads();
-        
+
         // x2 = at_list[ind2].x;
         // y2 = at_list[ind2].y;
         // z2 = at_list[ind2].z;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -149,6 +149,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
             bi += blockDim.x * 2;
             ei += blockDim.x * 2;
+            int k = 0;
             for(i = bi - blockDim.x; i < ei; i++, k++) {
                 
                 if(i < bi) {

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -134,6 +134,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
     int ind2;
 
     int j;
+    int k = 0;
     for(j = start; j < end; j++) {
 
         ind2 = j % PDH_acnt;
@@ -149,7 +150,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
             bi += blockDim.x * 2;
             ei += blockDim.x * 2;
-            int k = 0;
+            k = 0;
             for(i = bi - blockDim.x; i < ei; i++, k++) {
                 
                 if(i < bi) {

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -288,7 +288,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + block_size.x * sizeof(atom);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -101,16 +101,12 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         int k = 0;
         for(i = start; i < start + blockDim.x && i < PDH_acnt; i++, k++) {
             sharedAtoms[k] = at_list[i];
+            sharedAtoms1[k] = sharedAtoms[k];
         }
 
-        k = 0;
-        for(i = start; i < start + blockDim.x * 2 && i < PDH_acnt; i++, k++) {
-
-            if(i < start + blockDim.x) {
-                sharedAtoms1[k] = sharedAtoms[k];
-            } else {
-                sharedAtoms1[k] = at_list[i];    
-            }
+        // load one more block
+        for(; i < start + blockDim.x * 2; i++, k++) {
+            sharedAtoms1[k] = at_list[i % PDH_acnt];
         }
     }
 
@@ -288,7 +284,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(atom);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -71,114 +71,114 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
 __global__
 void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double PDH_res) {
 
-    extern __shared__ unsigned long long smem[];
+    // extern __shared__ unsigned long long smem[];
 
-    unsigned long long* shared_histo = smem;
-    atom* sharedAtoms = (atom*) &shared_histo[num_buckets];
-    atom* sharedAtoms1 = (atom*) &sharedAtoms[blockDim.x];
+    // unsigned long long* shared_histo = smem;
+    // atom* sharedAtoms = (atom*) &shared_histo[num_buckets];
+    // atom* sharedAtoms1 = (atom*) &sharedAtoms[blockDim.x];
 
-    long index_x = blockIdx.x * blockDim.x + threadIdx.x;
-    long index_y = blockIdx.y * blockDim.y + threadIdx.y;
+    // long index_x = blockIdx.x * blockDim.x + threadIdx.x;
+    // long index_y = blockIdx.y * blockDim.y + threadIdx.y;
 
-    // map the two 2D indices to a single linear, 1D index
-    long grid_width = gridDim.x * blockDim.x;
-    long index = index_y * grid_width + index_x;
+    // // map the two 2D indices to a single linear, 1D index
+    // long grid_width = gridDim.x * blockDim.x;
+    // long index = index_y * grid_width + index_x;
 
-    long i = index;
+    // long i = index;
 
-    //check the bound
-    if(i >= PDH_acnt) {
-        return;
-    }
+    // //check the bound
+    // if(i >= PDH_acnt) {
+    //     return;
+    // }
 
-    //for every first thread of the block
-    if(threadIdx.x == 0) {
-        for(i = 0; i < num_buckets; i++) {
-            shared_histo[i] = 0;
-        }
+    // //for every first thread of the block
+    // if(threadIdx.x == 0) {
+    //     for(i = 0; i < num_buckets; i++) {
+    //         shared_histo[i] = 0;
+    //     }
 
-        int start = index;
-        int k = 0;
-        for(i = start; i < start + blockDim.x && i < PDH_acnt; i++, k++) {
-            sharedAtoms[k] = at_list[i];
[…]

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -284,7 +284,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + block_size.x * sizeof(atom);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -284,7 +284,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + 1 * block_size.x * sizeof(atom);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -254,6 +254,8 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     //int gridSize = ceil(atomsCnt / (float)blockSize) + 1;
     //int stripe = 1024 / ;
 
+    printf("Histogram size: %d\n", num_buckets * sizeof(unsigned long long));
+
     int i, w;
     for(w = 0; w < workload; w++) {
 
@@ -284,7 +286,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + 1 * block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + 2 * block_size.x * sizeof(atom);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -254,8 +254,6 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     //int gridSize = ceil(atomsCnt / (float)blockSize) + 1;
     //int stripe = 1024 / ;
 
-    printf("Histogram size: %d\n", num_buckets * sizeof(unsigned long long));
-
     int i, w;
     for(w = 0; w < workload; w++) {
 
@@ -286,7 +284,8 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         gpu_one_body_functions_kernel<<<grid_size, block_size, smem1, streamComp1 >>>(g_s_atomsCnt, g_s_atom_list, g_s_res);
 
         //----------------------------------2 BODY KERNEL---------------------------------------------------
-        int smem2 = num_buckets * sizeof(unsigned long long) + 2 * block_size.x * sizeof(atom);
+        int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(atom);
+        printf("SMEM size: %d\n", smem2);
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -286,6 +286,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         //----------------------------------2 BODY KERNEL---------------------------------------------------
         int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(atom);
         printf("SMEM size: %d\n", smem2);
+        printf("Float size: %d\n", sizeof(float));
         gpu_two_body_functions_kernel<<<grid_size, block_size, smem2, streamComp2 >>>(g_s_atom_list, atomsCnt, d_histogram, num_buckets, PDH_res);
         
         cudaStreamSynchronize(streamComp1);

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -71,114 +71,123 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
 __global__
 void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double PDH_res) {
 
-    // extern __shared__ unsigned long long smem[];
+    extern __shared__ unsigned long long smem[];
 
-    // unsigned long long* shared_histo = smem;
-    // atom* sharedAtoms = (atom*) &shared_histo[num_buckets];
-    // atom* sharedAtoms1 = (atom*) &sharedAtoms[blockDim.x];
+    unsigned long long* shared_histo = smem;
+    coordinates* sharedAtoms = (coordinates*) &shared_histo[num_buckets];
+    coordinates* sharedAtoms1 = (coordinates*) &sharedAtoms[blockDim.x * 3];
 
-    // long index_x = blockIdx.x * blockDim.x + threadIdx.x;
-    // long index_y = blockIdx.y * blockDim.y + threadIdx.y;
+    long index_x = blockIdx.x * blockDim.x + threadIdx.x;
+    long index_y = blockIdx.y * blockDim.y + threadIdx.y;
 
-    // // map the two 2D indices to a single linear, 1D index
-    // long grid_width = gridDim.x * blockDim.x;
-    // long index = index_y * grid_width + index_x;
+    // map the two 2D indices to a single linear, 1D index
+    long grid_width = gridDim.x * blockDim.x;
+    long index = index_y * grid_width + index_x;
 
-    // long i = index;
+    long i = index;
 
-    // //check the bound
-    // if(i >= PDH_acnt) {
-    //     return;
-    // }
+    //check the bound
+    if(i >= PDH_acnt) {
+        return;
+    }
 
-    // //for every first thread of the block
-    // if(threadIdx.x == 0) {
-    //     for(i = 0; i < num_buckets; i++) {
-    //         shared_histo[i] = 0;
-    //     }
+    //for every first thread of the block
+    if(threadIdx.x == 0) {
+        for(i = 0; i < num_buckets; i++) {
+            shared_histo[i] = 0;
+        }
 
-    //     int start = index;
-    //     int k = 0;
-    //     for(i = start; i < start + blockDim.x && i < PDH_acnt; i++, k++) {
[…]

### 2016-06-05 types.h

@@ -11,6 +11,10 @@ typedef struct atom {
     float mass;
 } atom;
 
+typedef struct coordinates {
+    float x, y, z; //coordinates
+} coordinates;
+
 typedef struct query_results {
 	
 	float mass;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -75,7 +75,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
     unsigned long long* shared_histo = smem;
     coordinates* sharedAtoms = (coordinates*) &shared_histo[num_buckets];
-    coordinates* sharedAtoms1 = (coordinates*) &sharedAtoms[blockDim.x * 3];
+    coordinates* sharedAtoms1 = (coordinates*) &sharedAtoms[blockDim.x];
 
     long index_x = blockIdx.x * blockDim.x + threadIdx.x;
     long index_y = blockIdx.y * blockDim.y + threadIdx.y;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -109,9 +109,9 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
         // load one more block
         for(; i < start + blockDim.x * 2; i++, k++) {
-            sharedAtoms1[k].x = at_list[i].x;
-            sharedAtoms1[k].y = at_list[i].y;
-            sharedAtoms1[k].z = at_list[i].z;
+            sharedAtoms1[k].x = at_list[i % PDH_res].x;
+            sharedAtoms1[k].y = at_list[i % PDH_res].y;
+            sharedAtoms1[k].z = at_list[i % PDH_res].z;
         }
     }
 

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -109,9 +109,9 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
         // load one more block
         for(; i < start + blockDim.x * 2; i++, k++) {
-            sharedAtoms1[k].x = at_list[i % PDH_res].x;
-            sharedAtoms1[k].y = at_list[i % PDH_res].y;
-            sharedAtoms1[k].z = at_list[i % PDH_res].z;
+            sharedAtoms1[k].x = at_list[i % PDH_acnt].x;
+            sharedAtoms1[k].y = at_list[i % PDH_acnt].y;
+            sharedAtoms1[k].z = at_list[i % PDH_acnt].z;
         }
     }
 

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -115,79 +115,79 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         }
     }
 
-    __syncthreads();
+    // __syncthreads();
 
-    i = index;
+    // i = index;
     
-    int threadLoad = (PDH_acnt + 1) / 2;
+    // int threadLoad = (PDH_acnt + 1) / 2;
 
-    int start = i + 1;
-    int end = i + threadLoad;
+    // int start = i + 1;
+    // int end = i + threadLoad;
 
-    if(PDH_acnt % 2 == 0 && i < PDH_acnt / 2) {
-        end++;
-    }
+    // if(PDH_acnt % 2 == 0 && i < PDH_acnt / 2) {
+    //     end++;
+    // }
 
-    int bi = blockDim.x * blockIdx.x;   // shared block start
-    int ei = bi + blockDim.x * 2;           // shared block end
+    // int bi = blockDim.x * blockIdx.x;   // shared block start
+    // int ei = bi + blockDim.x * 2;           // shared block end
 
-    int ind1 = threadIdx.x;             // in this block from sharedAtoms
-    int ind2;
+    // int ind1 = threadIdx.x;             // in this block from sharedAtoms
+    // int ind2;
 
-    int j;
-    int k = 0;
-    for(j = start; j < end; j++) {
+    // int j;
+    // int k = 0;
+    // for(j = start; j < end; j++) {
 
-        ind2 = j % PDH_acnt;
+    //     ind2 = j % PDH_acnt;
 
-        double x1 = sharedAtoms[ind1].x;
-        double y1 = sharedAtoms[ind1].y;
-        double z1 = sharedAtoms[ind1].z;
+    //     double x1 = sharedAtoms[ind1].x;
+    //     double y1 = sharedAtoms[ind1].y;
+    //     double z1 = sharedAtoms[ind1].z;
 
-        double x2, y2, z2;
+    //     double x2, y2, z2;
 
-        __syncthreads();
+    //     __syncthreads();
 
-        if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
-            bi += blockDim.x * 2;
-            ei += blockDim.x * 2;
-            k = 0;
-            for(i = bi - blockDim.x; i < ei; i++, k++) {
[…]

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -115,79 +115,79 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         }
     }
 
-    // __syncthreads();
+    __syncthreads();
 
-    // i = index;
+    i = index;
     
-    // int threadLoad = (PDH_acnt + 1) / 2;
+    int threadLoad = (PDH_acnt + 1) / 2;
 
-    // int start = i + 1;
-    // int end = i + threadLoad;
+    int start = i + 1;
+    int end = i + threadLoad;
 
-    // if(PDH_acnt % 2 == 0 && i < PDH_acnt / 2) {
-    //     end++;
-    // }
+    if(PDH_acnt % 2 == 0 && i < PDH_acnt / 2) {
+        end++;
+    }
 
-    // int bi = blockDim.x * blockIdx.x;   // shared block start
-    // int ei = bi + blockDim.x * 2;           // shared block end
+    int bi = blockDim.x * blockIdx.x;   // shared block start
+    int ei = bi + blockDim.x * 2;           // shared block end
 
-    // int ind1 = threadIdx.x;             // in this block from sharedAtoms
-    // int ind2;
+    int ind1 = threadIdx.x;             // in this block from sharedAtoms
+    int ind2;
 
-    // int j;
-    // int k = 0;
-    // for(j = start; j < end; j++) {
+    int j;
+    int k = 0;
+    for(j = start; j < end; j++) {
 
-    //     ind2 = j % PDH_acnt;
+        ind2 = j % PDH_acnt;
 
-    //     double x1 = sharedAtoms[ind1].x;
-    //     double y1 = sharedAtoms[ind1].y;
-    //     double z1 = sharedAtoms[ind1].z;
+        double x1 = sharedAtoms[ind1].x;
+        double y1 = sharedAtoms[ind1].y;
+        double z1 = sharedAtoms[ind1].z;
 
-    //     double x2, y2, z2;
+        double x2, y2, z2;
 
-    //     __syncthreads();
+        __syncthreads();
 
-    //     if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
-    //         bi += blockDim.x * 2;
-    //         ei += blockDim.x * 2;
-    //         k = 0;
-    //         for(i = bi - blockDim.x; i < ei; i++, k++) {
[…]

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -152,9 +152,9 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
             bi += blockDim.x;
             ei += blockDim.x;
             k = 0;
-            for(i = bi - blockDim.x; i < ei; i++, k++) {
+            for(i = bi; i < ei; i++, k++) {
                 
-                if(i < bi) {
+                if(i < bi + blockDim.x) {
                     sharedAtoms1[k].x = sharedAtoms1[k + blockDim.x].x;
                     sharedAtoms1[k].y = sharedAtoms1[k + blockDim.x].y;
                     sharedAtoms1[k].z = sharedAtoms1[k + blockDim.x].z;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -166,11 +166,11 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
             }
         }
 
+        __syncthreads();
+        
         x2 = sharedAtoms1[ind2 - bi].x;
         y2 = sharedAtoms1[ind2 - bi].y;
         z2 = sharedAtoms1[ind2 - bi].z;
-
-        __syncthreads();
 
         // x2 = at_list[ind2].x;
         // y2 = at_list[ind2].y;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -138,8 +138,6 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
     int k = 0;
     for(j = start; j < end; j++) {
 
-        ind2 = j % PDH_acnt;
-
         double x1 = sharedAtoms[ind1].x;
         double y1 = sharedAtoms[ind1].y;
         double z1 = sharedAtoms[ind1].z;
@@ -153,7 +151,6 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
             ei += blockDim.x;
             k = 0;
             for(i = bi; i < ei; i++, k++) {
-                
                 if(i < bi + blockDim.x) {
                     sharedAtoms1[k].x = sharedAtoms1[k + blockDim.x].x;
                     sharedAtoms1[k].y = sharedAtoms1[k + blockDim.x].y;
@@ -167,7 +164,8 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         }
 
         __syncthreads();
-        
+        //ind2 = j % PDH_acnt;
+
         x2 = sharedAtoms1[ind2 - bi].x;
         y2 = sharedAtoms1[ind2 - bi].y;
         z2 = sharedAtoms1[ind2 - bi].z;

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -144,8 +144,6 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
         double x2, y2, z2;
 
-        __syncthreads();
-
         if(threadIdx.x == 0 && !(bi <= ind2 && ind2 < ei)) { //not finding in shared memory
             bi += blockDim.x;
             ei += blockDim.x;
@@ -164,19 +162,17 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         }
 
         __syncthreads();
-        //ind2 = j % PDH_acnt;
 
-        x2 = sharedAtoms1[ind2 - bi].x;
-        y2 = sharedAtoms1[ind2 - bi].y;
-        z2 = sharedAtoms1[ind2 - bi].z;
+        int load = 0;
+        while(load < blockDim.x) {
+            x2 = sharedAtoms1[ind2 - bi].x;
+            y2 = sharedAtoms1[ind2 - bi].y;
+            z2 = sharedAtoms1[ind2 - bi].z;
 
-        // x2 = at_list[ind2].x;
-        // y2 = at_list[ind2].y;
-        // z2 = at_list[ind2].z;
-
-        double dist = sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2) + (z1 - z2) * (z1 - z2));
-        int h_pos = (int) (dist / PDH_res);
-        atomicAdd(&shared_histo[h_pos], 1);
+            double dist = sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2) + (z1 - z2) * (z1 - z2));
+            int h_pos = (int) (dist / PDH_res);
+            atomicAdd(&shared_histo[h_pos], 1);
+        }
     }
 
     __syncthreads();

### 2016-06-05 gpu_single.cu
automatic triggered changes

@@ -136,7 +136,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
 
     int j;
     int k = 0;
-    for(j = start; j < end; j++) {
+    for(j = start; j < end; j += blockDim.x) {
 
         double x1 = sharedAtoms[ind1].x;
         double y1 = sharedAtoms[ind1].y;
@@ -164,7 +164,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
         __syncthreads();
 
         int load = 0;
-        while(load < blockDim.x) {
+        while(load < blockDim.x && j < end) {
             x2 = sharedAtoms1[ind2 - bi].x;
             y2 = sharedAtoms1[ind2 - bi].y;
             z2 = sharedAtoms1[ind2 - bi].z;
@@ -172,6 +172,8 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
             double dist = sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2) + (z1 - z2) * (z1 - z2));
             int h_pos = (int) (dist / PDH_res);
             atomicAdd(&shared_histo[h_pos], 1);
+            load++;
+            j++;
         }
     }
 

### 2016-06-07 gpu_single.cu
automatic triggered changes

@@ -312,7 +312,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
     printf("%-40s %.3f\n", "Inertia Z Axis: ", res->inertiaZ);
     printf("%-40s %.3f\n", "Depole Moment Z Axis: ", res->depoleMoment);
     printf("%-40s %.3fmillis\n", "Running time: ", elapsed);
-    //output_histogram(histogram, num_buckets);
+    output_histogram(histogram, num_buckets);
 
     /**
     * MEM FREE

### 2016-06-07 gpu_single.cu
automatic triggered changes

@@ -69,7 +69,7 @@ void gpu_one_body_functions_kernel(int* g_s_atomsCnt, atom* g_s_atom_list, query
 
 //2 body functions (SDH or POINT DISTANCE HISTOGRAM)
 __global__
-void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double PDH_res) {
+void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double bucket_width) {
 
     extern __shared__ unsigned long long smem[];
 
@@ -170,7 +170,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
             z2 = sharedAtoms1[ind2 - bi].z;
 
             double dist = sqrt((x1 - x2) * (x1 - x2) + (y1 - y2) * (y1 - y2) + (z1 - z2) * (z1 - z2));
-            int h_pos = (int) (dist / PDH_res);
+            int h_pos = (int) (dist / bucket_width);
             atomicAdd(&shared_histo[h_pos], 1);
             load++;
             j++;
@@ -186,7 +186,7 @@ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, in
     }
 }
 
-void output_histogram(bucket* hist, int num_buckets){
+void output_histogram(bucket* hist, int num_buckets) {
     int i; 
     unsigned long long total_cnt = 0;
     for(i = 0; i < num_buckets; i++) {
@@ -202,12 +202,9 @@ void output_histogram(bucket* hist, int num_buckets){
 }
 
 
-void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
+void run_single_kernel(int atomsCnt, atom* atomList, int workload, float bucket_width, float space) {
 
-    int BOX_SIZE = 175;
-    int PDH_res = 1;
-
-    int num_buckets = BOX_SIZE + 1;
+    int num_buckets = space / bucket_width + 1;
 
     query_results* res = (query_results*) malloc(sizeof(query_results));
     bucket* histogram = (bucket *)malloc(sizeof(bucket) * num_buckets); 
@@ -292,7 +289,7 @@ void run_single_kernel(int atomsCnt, atom* atomList, int workload) {
         int smem2 = num_buckets * sizeof(unsigned long long) + 3 * block_size.x * sizeof(coordinates);
[…]

### 2016-06-07 gpu_single.h

@@ -6,6 +6,6 @@
 __global__ void gpu_one_body_functions_kernel(int* g_s_atoms_cnt, atom* g_s_atom_list, query_results* g_s_res);
 __global__ void gpu_two_body_functions_kernel(atom* at_list, int PDH_acnt, bucket* hist, int num_buckets, double PDH_res);
 
-void run_single_kernel(int atomsCnt, atom* atomList, int workload);
+void run_single_kernel(int atomsCnt, atom* atomList, int workload, float bucket_width, float space);
 
 #endif

### 2016-06-07 main.cu

@@ -43,6 +43,13 @@ int main(int argc, char *argv[]) {
         workload = atoi(argv[3]);
     }
 
+    float bucket_width, space;
+    if(argc < 6) {
+        printf("Usage: ./main atomCount file workload bucket-width space-width\n");
+        bucket_width = atof(argv[4]);
+        space = atof(argv[5]);
+    }
+
     std::ifstream stream(fileName.c_str());
     std::cout << "Reading file: " << fileName << std::endl;
 
@@ -72,7 +79,7 @@ int main(int argc, char *argv[]) {
             if(atomCount > 0) {
                 std::cout << "**********************Frame #" << heads << "*****************" << std::endl;
                 std::cout << atomCount << " atoms read." << std::endl;
-                run_single_kernel(atomsCnt, atomsList, workload);
+                run_single_kernel(atomsCnt, atomsList, workload, bucket_width, space);
             }
             
             // check for error
@@ -105,9 +112,9 @@ int main(int argc, char *argv[]) {
         lineStream >> atomsList[atomCount].charge;
         lineStream >> atomsList[atomCount].mass;
 
-        atomsList[atomCount].x = rand() % 100;
-        atomsList[atomCount].y = rand() % 100;
-        atomsList[atomCount].z = rand() % 100;
+        atomsList[atomCount].x = (((float)rand() % 100) / 100.0f) * space;
+        atomsList[atomCount].y = (((float)rand() % 100) / 100.0f) * space;
+        atomsList[atomCount].z = (((float)rand() % 100) / 100.0f) * space;
 
         atomCount++;
     }

### 2016-06-07 main.cu
automatic triggered changes

@@ -112,9 +112,9 @@ int main(int argc, char *argv[]) {
         lineStream >> atomsList[atomCount].charge;
         lineStream >> atomsList[atomCount].mass;
 
-        atomsList[atomCount].x = (((float)rand() % 100) / 100.0f) * space;
-        atomsList[atomCount].y = (((float)rand() % 100) / 100.0f) * space;
-        atomsList[atomCount].z = (((float)rand() % 100) / 100.0f) * space;
+        atomsList[atomCount].x = ((rand() % 100) / 100.0f) * space;
+        atomsList[atomCount].y = ((rand() % 100) / 100.0f) * space;
+        atomsList[atomCount].z = ((rand() % 100) / 100.0f) * space;
 
         atomCount++;
     }

### 2016-06-07 main.cu
automatic triggered changes

@@ -46,9 +46,11 @@ int main(int argc, char *argv[]) {
     float bucket_width, space;
     if(argc < 6) {
         printf("Usage: ./main atomCount file workload bucket-width space-width\n");
-        bucket_width = atof(argv[4]);
-        space = atof(argv[5]);
+        exit(1);
     }
+
+    bucket_width = atof(argv[4]);
+    space = atof(argv[5]);
 
     std::ifstream stream(fileName.c_str());
     std::cout << "Reading file: " << fileName << std::endl;

### 2016-06-09 thesis_paper/source.tex
thesis stuff

@@
 
+\subsection{Advanced SDH implementation}
+
+We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.
+
+\begin{figure}
+ \centerline{ \includegraphics[width=0.3\columnwidth]{images/improved_sdh}}
+ \caption{New approach of using shared memory}
+ \label{fg:improved_sdh}
+\end{figure}
+
+Basically, as you can see from Figure \ref{fg:improved_sdh}, now we allocate more shared memory.
+Yellow sections are the atoms that are stored in shared memory throughout the whole thread execution of the entire block, since those atoms are needed at every step, while the green ones are the working zones.
+We load twice size of the block into shared memory, because the very first one thread in the block is going to communicate starting from second one, while the last one will start with the next block, which means that it will have to be twice size.
+Fortunately, the first part of the threads in the block will consequently use the next block which is already loaded, thus, at each step we load one more block size simply moving window and re-using the ones that are already loaded for first part.
+
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+
+$Duration $ & $ Grid Size $ & $ Block Size $ & $ DSMem* $ & $ Size $ & $ Throughput $ \\ \hline 
+$1.4080us $ & $ - $ & $ - $ & $ - $ & $ 4B $ & $ 2.8409MB/s $ \\ \hline 
+$1.0769ms $ & $ - $ & $ - $ & $ - $ & $ 3.6000MB $ & $ 3.3429GB/s $ \\ \hline 
+$1.0880us $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 33.088MB/s $ \\ \hline 
+$1.2800us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 1.1000GB/s $ \\ \hline 
+$3.0340ms $ & $ (98, 1, 1) $ & $ (1024, 1, 1) $ & $ 40.960KB $ & $ - $  & $-$ \\ \hline 
+$24.5691s $ & $ (98, 1, 1) $ & $  (1024, 1, 1) $ & $ 38.272KB $ & $ - $ & $ - $ \\ \hline 
+$2.7520us $ & $ - $ & $ - $ & $ - $ & $ 36B $ & $ 13.081MB/s $ \\ \hline 
+$2.7830us $ & $ - $ & $ - $ & $ - $ & $ 1.4080KB $ & $ 505.93MB/s $    
+ \\ \hline
+
+	\hline
+  \end{tabular}
+    \caption{Memory profiling results for 100,000 atoms in improved SDH}
+    	  \label{tb:gpumem100000_improved}
+\end{table}
+
+\begin{table}[h!]
+	\renewcommand*{\arraystretch}{1.5}
+	\tabcolsep=0.12cm
+  \centering
+  \begin{tabular}{| c | c | c | c | c | c |}
+   \hline
+	Device & Context & Stream & Kernel & Warps Launched & Shared Load \\ \hline
+	GRID K520 (0) & 1 & 13 & ${gpu\_one\_body\_function}$ & 3136 & 196 \\ \hline
+	GRID K520 (0) & 1 & 14 & ${gpu\_two\_body\_function}$ & 3136 & 246,323,337 \\ \hline
+
+  \end{tabular}
+    \caption{GPU profiling results for 100,000 atoms for improved SDH}
+    	  \label{tb:gpuprofile100000}
+\end{table}
+
+\begin{figure}
+ \centerline{ \includegraphics[width=0.5\columnwidth]{images/sdhvsimproved}}
+ \caption{Comparison of GPU SDH with GPU SDH Improved}
+ \label{fg:gpusdhsdh}
+\end{figure}
+
+As you may have noticed, even though we increased utilization of shared memory, number of loads is around the same.
+The trick is that we replaced atoms in shared memory with simpler data structure called $coordinates$ that would contains 3 float type variables which saves huge amount of memory.
+The results that we have achieved are extremely great, in general all the run times get twice less in comparison with simple SDH approach.
+You may observe run times for some sample data sets on Figure \ref{fg:gpusdhsdh}.
+In general, it is quite visible that it is twice less.
+This is a sample improvement which proves that a little more elaboration on a particular query can bring much better results which could end up saving scientists days and months of time of computation.
+
 \subsection{Overall performance}
@@
 \begin{figure}
- \centerline{\includegraphics[width=0.7\columnwidth]{images/chart_max}}
+ \centerline{\includegraphics[width=0.8\columnwidth]{images/overall_speedup}}
  \caption{Speedup for larger number of atoms}
@@
 \end{figure}
+
+With improved version of SDH speedup gets up to more than 75 times (Figure \ref{fg:chart_max}).
+We believe that these numbers could get even better depending on queries and elaboration of improvement of particular cases.
 

### 2016-08-10 thesis_paper/source.tex
Modified bunch of things regarding styling of the paper

@@
 
-\linespread{2}
-
-\titleformat{\chapter}[display]
-        {\normalfont\Large\bfseries\centering}{\chaptertitlename\ \thechapter}{0pt}{\Large\uppercase}
+\titleformat{\chapter}[display]{\normalfont\Large\bfseries\centering}{\chaptertitlename\ \thechapter}{0pt}{\Large\uppercase}
 
@@
 
-%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
-
 \begin{document}
-
-%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
-%   TITLEPAGE
-%
-%   due to variety of titlepage schemes it is probably better to make titlepage manually
-%
-%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
+{
 \thispagestyle{empty}
 
-{%%%
-\sffamily
-\centering
-
-{\Large 
-Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU
-}
-
-\vspace{1.5cm}
-
-{
-by
-}
-\vspace{.5cm}
-
-{ 
-Iliiazbek Akhmedov
-}
-
-\vspace{3.5cm}
+\linespread{1}\selectfont
+\centering
+\vspace*{5\baselineskip}
+Parallelization of Push-based System for
+
+\vspace*{1\baselineskip}
+
+Molecular Simulation Data Analysis with GPU 
+\vspace*{3\baselineskip}
+
+%{ by }
+\vspace*{3\baselineskip}
+
+{ Iliiazbek Akhmedov }
+
+\vspace*{5\baselineskip}
 
@@
 
-Department of Computer Science and Engineering
-
-College of Engineering
+Department of Computer Science and Engineering College of Engineering
 
@@
 
-\vspace{3.5cm}
+\vspace*{3\baselineskip}
 Major Professor:  Yicheng Tu, Ph.D.
-\vspace{\fill}
-
+
+\vspace*{2\baselineskip}
 \vspace{.5cm}
@@
 N/A
-\vspace{\fill}
-
+
+\vspace*{3\baselineskip}
+Keywords: data processing, cuda optimization, big data,
+
+streaming
+
+\vspace*{1\baselineskip}
 Copyright \textcopyright\space 2016, Iliiazbek Akhmedov
 
-%%%
-}%%%
+}
 
 \clearpage
-
-\newgeometry{left=20mm,right=20mm,top=35mm,bottom=35mm}
-
 \pagestyle{plain}
@@
 
+\vspace*{5\baselineskip}
+\linespread{2}\selectfont
 \textbf{DEDICATION}
-\vspace{0.5cm}
+
+\vspace*{1\baselineskip}
 
@@
 \centering
+
+\linespread{1}\selectfont
+\vspace*{5\baselineskip}
+
 \textbf{ACKNOWLEDGMENTS}
-\vspace{0.5cm}
+
+\vspace*{1\baselineskip}
+
+\linespread{2}\selectfont
 
@@
 
-\linespread{1}
-
+\linespread{1}\selectfont
 \tableofcontents
 
+\linespread{2}\selectfont
 \addcontentsline{toc}{chapter}{\listfigurename}
@@
 
-\linespread{2}
-\clearpage
+\linespread{1}\selectfont
+\vspace*{5\baselineskip}
 
 \addcontentsline{toc}{chapter}{\abstractname}
+
+\vspace*{1\baselineskip}
+
+\linespread{2}\selectfont
+
 \begin{abstract}
@@
 
+\subsection{SDH resolution performance impact}
+
 \subsection{Overall performance}
@@
 It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
-It is because we've used static approach of computing grid dimensions.
+One of possible reasons could the fact that we've used static approach of computing grid dimensions.
 Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.

### 2016-08-24 thesis_paper/source.tex
Some paper modifications
Draft version of thesis paper

@@
 
-\titleformat{\chapter}[display]{\normalfont\Large\bfseries\centering}{\chaptertitlename\ \thechapter}{0pt}{\Large\uppercase}
+\titleformat{\chapter}[hang] 
+{\centering\normalfont\bfseries}{\MakeUppercase{\chaptertitlename}\ \thechapter:}{0.5em}{\uppercase}
+
+\titleformat{\section}
+  {\normalfont\fontsize{12}{15}\bfseries}{\thesection}{0.5em}{}
+
+\makeatletter
+
+\newcommand\frontmatter{%
+    \cleardoublepage
+  %\@mainmatterfalse
+  \pagenumbering{roman}}
+
+\newcommand\mainmatter{%
+    \cleardoublepage
+ % \@mainmattertrue
+  \pagenumbering{arabic}}
+
+\newcommand\backmatter{%
+  \if@openright
+    \cleardoublepage
+  \else
+    \clearpage
+  \fi
+ % \@mainmatterfalse
+   }
+
+\makeatother
 
@@
 
-% Make chapter titles uppercase is a pain.
-% The internal chapter macros must be redefined.
-\makeatletter
-\renewcommand\chapter{\if@openright\cleardoublepage\else\clearpage\fi
-                    \thispagestyle{plain}%
-                    \global\@topnum\z@
-                    \@afterindentfalse
-                    \secdef\@chapter\@schapter}
-\def\@chapter[#1]#2{\ifnum \c@secnumdepth >\m@ne
-                       \if@mainmatter
-                         \refstepcounter{chapter}%
-                         \typeout{\@chapapp\space\thechapter.}%
-                         \addcontentsline{toc}{chapter}%
-                         {\protect\numberline{\thechapter}\texorpdfstring{\MakeUppercase{#1}}{#1}}%
-                       \else
-                       \addcontentsline{toc}{chapter}{\texorpdfstring{\MakeUppercase{#1}}{#1}}%
-                       \fi
-                    \else
-                    \addcontentsline{toc}{chapter}{\texorpdfstring{\MakeUppercase{#1}}{#1}}%
-                    \fi
-                    \chaptermark{#1}%
-                    \addtocontents{lof}{\protect\addvspace{10\p@}}%
-                    \addtocontents{lot}{\protect\addvspace{10\p@}}%
-                    \if@twocolumn
-                      \@topnewpage[\@makechapterhead{#2}]%
-                    \else
-                      \@makechapterhead{#2}%
-                      \@afterheading
-                    \fi}
-\makeatother
-
 \begin{document}
 {
+\frontmatter
+
 \thispagestyle{empty}
@@
 
-\addcontentsline{toc}{chapter}{\abstractname}
+\addcontentsline{toc}{chapter}{ABSTRACT}
 
@@
 
-\begin{abstract}
+\chapter*{Abstract}
 \justify
@@
 In the new approach due to high performance of GPU processing and its nature, custom data structures were not even needed much, though it didn't bear any loss in precision and performance.
-\end{abstract}
 
@@
 
-\chapter{Introduction}
-
-\justify
-In various sciences simulation systems take big place and often times they may be the clue for results.
+\mainmatter
+\linespread{1}\selectfont
+
+\chapter{Introduction} \linespread{2}\selectfont \justify
+
+\hspace{3em} In various sciences simulation systems take big place and often times they may be the clue for results.
 One of such sciences, which is primarily related to this paper, is physics.
 In this case, from computer engineering point of view, simulations have the following flow:
-
-\vspace{5mm}
 
@@
 
-\vspace{3mm}
 The purpose of this paper is primarily focused on the third step, which basically involves the entire simulation data to be processed by the analysis software.
@@
 These systems are most of the time N-body simulations.
-The number of atoms in simulations vary in hundred of thousands, particularly, we may observe two simulation systems of a collagen fiber structure and dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system consisting of 890,000 and 402,400 atoms respectively on Figure \ref{fg:collagen_dipal}.
+The number of atoms in simulations vary in hundred of thousands, particularly, we may observe two simulation systems of a collagen fiber structure and dipalmitoylphosphatidylcholine (DPPC) bi-layer lipid system consisting of 890,000 and 402,400 atoms as mentioned and illustrated in previous work of the system \cite{mainPaper}.
 Simulation data represents number of records of physical properties such as mass, charge, velocity, coordinates, and forces for each item aggregated as frames, where each frame represents a snapshot of time, placed with a fixed time interval which may also vary depending on the simulation itself and simulation precision requirement.
@@
 
-\begin{figure}
- \centerline{ \includegraphics[width=0.5\columnwidth]{images/sample_snapshot_of_simulation.eps} }
- \caption{ Snapshots of two MS systems with 890,000 atoms with 402,400 atoms}
- \label{fg:collagen_dipal}
-\end{figure}
-
 \section{Problem Statement}
-Simulation software systems, in general, follow the same methodology of running and storing simulation data.
+\hspace{3em} Simulation software systems, in general, follow the same methodology of running and storing simulation data.
 The simulation software system examples are: Gromacs \cite{GROMACS4}, VMD\cite{VMD}, MDAnalysis~\cite{MDAnalysis}, Wordom~\cite{wordom}, MD-TRACKS~\cite{MDtracks}, SimulaidOne~\cite{Simulaid}, Charmm~\cite{CHARMM}.
@@
 
-\begin{figure}
- \centerline{ \includegraphics[width=0.5\columnwidth]{images/speedup838K-pomalo-za-4.eps} }
- \caption{ Speed up over different levels of atom selection for 838,400 atoms}
- \label{fg:sample_estimation_old_paper}
-\end{figure}
-
 \section{Our approach with improvement}
 
-Reviewing the approach we are coming up with and what it was before helps us analyzing the weak parts and things that we can focus on to eliminate old bottlenecks.
+\hspace{3em} Reviewing the approach we are coming up with and what it was before helps us analyzing the weak parts and things that we can focus on to eliminate old bottlenecks.
 
@@
 Since we have to load the data every time there is a different query versus push-based system loads a chunk of data once, it is quite noticeable how the framework contributes to efficiency of analysis of simulation data.
-You can observe some estimation samples in Figure \ref{fg:sample_estimation_old_paper}.
 Even though the memory loading and pushing steps are the key features of the proposed design, nevertheless, having general knowledge about the network of queries, in this paper, we will try to focus on optimizing the actual execution of them.
@@
 
-In collaboration with Physics Department, the original motivation came from the need of improving analytic query executions performance over big data being generated by simulation software.
+\hspace{3em} In collaboration with Physics Department, the original motivation came from the need of improving analytic query executions performance over big data being generated by simulation software.
 The idea was to take input stream coming from simulation software (ideally live) and run network of queries very efficiently, because sometimes they might take days and more to compute.
@@
 
-Push-based System for molecular simulation data analysis with its improved GPU features has couple of main focuses in terms of technical implementation and challenges related to those that have already been tried to be solved by other different approaches and software.
+\hspace{3em} Push-based System for molecular simulation data analysis with its improved GPU features has couple of main focuses in terms of technical implementation and challenges related to those that have already been tried to be solved by other different approaches and software.
 Here is the review of other works.
@@
 
-First of all, this system is a tool for scientific computation and in this case particularly for molecular simulation data analysis.
+\hspace{3em} First of all, this system is a tool for scientific computation and in this case particularly for molecular simulation data analysis.
 PeriScope is a project and initiative by University of Michigan which is a system that will optimize data management for scientific explorations.
@@
 \section{Stream processing}
-Another major modern issue is live data streaming and live processing, which is also possible with our system.
+\hspace{3em} Another major modern issue is live data streaming and live processing, which is also possible with our system.
 Basically, streaming function is a big challenge and need nowadays, because such networks as Twitter and Facebook are able to keep track of live trends and events, including people's reaction and feedback that have to be reflected real time for tremendous amount of users across the globe.
@@
 
-We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data.
+\hspace{3em} We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data.
 Traditional software tools are not able to process Big Data apparently because of huge data sets, thus, special frameworks were developed to handle them.
@@
 \section{GPU and Big Data Processing}
-GPU is becoming more popular being used by not only graphics processing, but machine learning, deep learning, search algorithms, sorting, and so on.
+\hspace{3em} GPU is becoming more popular being used by not only graphics processing, but machine learning, deep learning, search algorithms, sorting, and so on.
 Having an ability to optimize in tremendous speedups it is being involved in video classifications, natural language processing.
@@
 \chapter{Benchmark scientific queries}\label{sc:querynetwork}
-Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
+\hspace{3em} Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
 Often times, the queries that need to be run over the data look like selection and some kind of accumulation, if it is not more complicated.
@@
 
+\vspace{3em}
+
 \begin{table}[h]
@@
 \hline %inserts horizontal line
-Function Name & Equation/Description \\[0.5ex] 
+Function Name & Equation\\[0.5ex] 
 \hline
@@
 \hline
-Moment of Inertia on z axis & $\begin{array} {lcl} I_z & = & \sum\limits_{i=1}^n m_ir_{zi} \end{array}$ \\[0.5ex]
-\hline
 Sum of masses & $\begin{array} {lcl} M & = & \sum\limits_{i=1}^n m_i \end{array}$ \\[0.5ex]
@@
 \hline
-Electron Density & $\begin{array} {lcl} ED & = & \frac{\sum\limits_{i=1}^n (e_i-q_i)}{dz \cdot x \cdot y} \end{array}$ \\[0.5ex]
-\hline
-Heat Capacity & $\begin{array} {lcl} HC & = & \frac{3000 \cdot \sqrt{T} \cdot boltz}{2 \cdot \sqrt{T}-n \cdot df \cdot VarT} \end{array}$ \\[0.5ex]
-\hline
-Isothermal Compressibility & $\begin{array} {lcl} I & = & \frac{VarV}{V_{avg} \cdot boltz \cdot T \cdot PresFac} \end{array}$ \\[0.5ex]
-\hline
-Mean Square Displacement& $\begin{array} {lcl} msd & = & \langle (r_{t+\Delta_t}-r_t)^2 \rangle \end{array}$ \\[0.5ex]
-\hline
-Diffusion Constant & $\begin{array} {lcl} D_t & = & \frac{6 \cdot msd(t)}{t} \end{array}$ \\[0.5ex]
-\hline Velocity Autocorrelation & $\begin{array} {lcl} V_{acor} & =&
-\langle (V_{t+\Delta_t}\cdot V_t) \rangle \end{array}$ \\[0.5ex]
-\hline
-Force Autocorrelation & $\begin{array} {lcl} F_{acor} & = & \langle (F_{t+\Delta_t} \cdot F_t) \rangle \end{array}$ \\[0.5ex]
-\hline
 Density Function & Histogram of atom counts \\[0.5ex]
@@
 \end{tabular}
-\caption[Popular analytical queries in MS]{Popular analytical queries in MS }
+\caption[Some of the popular analytical queries in MS]{Popular analytical queries in MS }
 \label{tb:queries}
@@
 \section{One body functions}
-\emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
+\hspace{3em} \emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
 
@@
 
-The proposed design has a pretty straightforward structure which is going to be explained in this section.
+\hspace{3em} The proposed design has a pretty straightforward structure which is going to be explained in this section.
 Shortly, the system starts with original input files (or data) and gets consumed in a sequential manner by simple disk IO framework into random access memory by chunks, then, chunk gets uploaded to GPU device and processed.
@@
 \section{Data input}
-Since the size of the data in the problem is huge, organization of the data flow is crucial.
+\hspace{3em} Since the size of the data in the problem is huge, organization of the data flow is crucial.
 As mentioned before, original data is retrieved from simulation software output files and in our case it is Gromacs Software's trajectory files\footnote{The file types, their structure, and their decompression has been well explained in the previous paper in section "Data orgranization in main memory"}.
@@
 
-MS data roughly can be represented as a list of atoms with physical properties aggregated as frames.
+\hspace{3em} MS data roughly can be represented as a list of atoms with physical properties aggregated as frames.
 Thus, we consider a single frame as a chunk which will be loaded into the memory\footnote{Memory in this design might be interpreted differently, since we have a separate device memory on GPUs, but it will be explained better later} and pushed through our network of queries.
@@
 
-Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
+\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
 Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
@@
 
-As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
+\hspace{3em} As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
 Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.
@@
 
-Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
+\hspace{3em} Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
 Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.
@@
 
-The approach in GPU version of one-body computation code is a lot different, because there are two major changes: the way we deal with the data in memory and mapping of GPU thread kernels to atoms.
+\hspace{3em} The approach in GPU version of one-body computation code is a lot different, because there are two major changes: the way we deal with the data in memory and mapping of GPU thread kernels to atoms.
 Both of these approaches drastically change the way we perceive the solution implementation.
@@
 
-Since one of the physical queries is RDF which has underlying SDH calculation, we will describe our approach with SDH for CPU and GPU versions.
+\hspace{3em} Since one of the physical queries is RDF which has underlying SDH calculation, we will describe our approach with SDH for CPU and GPU versions.
 
@@
 
-Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
+\hspace{3em} Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
 Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them, though results don't change a lot for sequential computation.
@@
 
-The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same.
+\hspace{3em} The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same.
 We do brute force  $O(\frac{N^2}{2})$ complex solution which is pretty much looking at distance between every pair of atoms.
@@
 
-As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
+\hspace{3em} As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
 
@@
 
-The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
+\hspace{3em} The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
 Experiments were run on operating systems which didn't have any additional workload besides usual installation, and the operating systems particularly were Amazon Linux, which doesn't have a lot of modifications except for some additional features for management with Amazon EC-2 Console.
@@
 
-There are primarily several parts that are really important in this hardware stack.
+\hspace{3em} There are primarily several parts that are really important in this hardware stack.
 Since this is primarily big data processing, we care about processing power and memory, Of course, things like PCI bus and all the transportation are important as well, but for speedup benchmarks against CPU vs GPU those stay the same for both sides.
@@
 \subsection{CPU}
-We have summarized main CPU parameters for CPU version and it is described in Table \ref{tb:cpuhardware}. As you may notice, it is a regular server node CPU with 2.60GHz and 3 levels of L cache.
+\hspace{3em} We have summarized main CPU parameters for CPU version and it is described in Table \ref{tb:cpuhardware}.
+As you may notice, it is a regular server node CPU with 2.60GHz and 3 levels of L cache.
 
@@
 
-The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
+\hspace{3em} The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
 So, for CPU in GPU improved version refer to the same Table \ref{tb:cpuhardware}.
@@
 
-As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
+\hspace{3em} As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments.
 It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format.
@@
 
-We have run number of experiments with different number of atoms and frames.
+\hspace{3em} We have run number of experiments with different number of atoms and frames.
 Results vary as we choose different data sets, but we can see some tendency towards growth and, fortunately, it is very positive for our improved version of push-based system.
@@
 
-Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes.
+\hspace{3em} Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes.
 You can observe summary in Figures \ref{fg:exp_1000_10}, \ref{fg:exp_1000_100}, \ref{fg:exp_1000_1000}, \ref{fg:speedup_1000}.
@@
 
-You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
+\hspace{3em} You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
 There is a small overhead we have to consider before computing part: in order to let GPU process the data, we have to transfer the data to Global Memory, and upon completion it is needed to load the results back to RAM to show it to the user.
@@
 
-When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good. Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
+\hspace{3em} When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good.
+Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
 
@@
 
-We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.
+\hspace{3em} We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.
 
 \begin{figure}
- \centerline{ \includegraphics[width=0.3\columnwidth]{images/improved_sdh}}
+ \centerline{ \includegraphics[width=0.3\columnwidth]{images/sharedMem}}
  \caption{New approach of using shared memory}
@@
 
-\subsection{SDH resolution performance impact}
-
 \subsection{Overall performance}
 
-It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
+\hspace{3em} It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
 One of possible reasons could the fact that we've used static approach of computing grid dimensions.
@@
 \chapter{Conclusions and Future Work}\label{sc:conclusion}
-The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
+\hspace{3em} The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
 As you could see from result on charts, the proposed improved version of the system could save a lot of time for scientists which is up to 60-70 times less of time.

### 2016-09-16 thesis_paper/source.tex
styling
Last known good state

@@
-\documentclass[11pt,a4paper]{report}
+\documentclass[12pt,letterpaper]{report}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{geometry}
-% PDF VIEW
-\geometry{total={210mm,297mm},
-left=25mm,right=25mm,%
-bindingoffset=0mm, top=25mm,bottom=25mm}
-%PRINT
-\geometry{total={210mm,297mm},
-left=20mm,right=20mm,
-bindingoffset=0mm, top=15mm,bottom=35mm}
+\geometry{left=1in,right=1in,top=1in,bottom=1in}
 
@@
     \cleardoublepage
-  %\@mainmatterfalse
   \pagenumbering{roman}}
@@
     \cleardoublepage
- % \@mainmattertrue
   \pagenumbering{arabic}}
@@
   \fi
- % \@mainmatterfalse
    }
@@
 
-\renewcommand{\contentsname}{CONTENTS}
+\renewcommand{\contentsname}{TABLE OF CONTENTS}
 \renewcommand{\listfigurename}{LIST OF FIGURES}
@@
 {
-\frontmatter
-
-\thispagestyle{empty}
-
-\linespread{1}\selectfont
-\centering
-\vspace*{5\baselineskip}
+\frontmatter\thispagestyle{empty}\linespread{1}\selectfont\centering\vspace*{4\baselineskip}
 Parallelization of Push-based System for
@@
 
-%{ by }
+{ by }
+
 \vspace*{3\baselineskip}
@@
 \vspace*{2\baselineskip}
-\vspace{.5cm}
 Date of Approval:
 
-N/A
+August 23, 2016
 
@@
 
-\clearpage
-\pagestyle{plain}
-\centering
-
-\vspace*{5\baselineskip}
+\linespread{1}\selectfont\clearpage
+\thispagestyle{empty}
+\centering
+\linespread{1}\selectfont
+\vspace*{3\baselineskip}
+\textbf{DEDICATION}
 \linespread{2}\selectfont
-\textbf{DEDICATION}
-
-\vspace*{1\baselineskip}
-
 \begin{flushleft}
@@
 
-\pagestyle{plain}
-\centering
-
+\linespread{1}\selectfont\clearpage
+\thispagestyle{empty}
+\centering
 \linespread{1}\selectfont
-\vspace*{5\baselineskip}
-
+\vspace*{3\baselineskip}
 \textbf{ACKNOWLEDGMENTS}
-
-\vspace*{1\baselineskip}
-
 \linespread{2}\selectfont
-
 \begin{flushleft}
@@
 \linespread{1}\selectfont
+\frontmatter
 \tableofcontents
@@
 \linespread{1}\selectfont
-
 \chapter{Introduction} \linespread{2}\selectfont \justify

### 2016-10-08 thesis_paper/source.tex
Last known good state
Last known good state goooood
automatic triggered changes
Formatting corrections

@@
 \usepackage{titlesec}
-\usepackage{wrapfig,subfigure,setspace,verbatim,cite,url,graphicx,float}
+\usepackage{wrapfig,setspace,verbatim,cite,url,graphicx,float}
 \usepackage{amsmath}
@@
 \usepackage{url}
-\usepackage{subfigure}
 \usepackage{cleveref}
@@
 \usepackage{geometry}
-\geometry{left=1in,right=1in,top=1in,bottom=1in}
-
-\titleformat{\chapter}[hang] 
-{\centering\normalfont\bfseries}{\MakeUppercase{\chaptertitlename}\ \thechapter:}{0.5em}{\uppercase}
+\geometry{left=1in,right=1in,top=1in,bottom=1.2in}
+
+\titleformat{\chapter}[hang]{\normalfont\normalsize\filcenter\bfseries}{\centering\MakeUppercase{\chaptertitlename}\ \thechapter:}{0.5em}{\uppercase}
 
@@
   {\normalfont\fontsize{12}{15}\bfseries}{\thesection}{0.5em}{}
+\titleformat{\subsection}
+  {\normalfont\fontsize{12}{15}\bfseries}{\thesection}{0.5em}{} \titleformat{\subsubsection}
+  {\normalfont\fontsize{12}{15}\bfseries}{\thesection}{0.5em}{}
+%toc
+
+\titlespacing*{\chapter} {0pt}{50pt}{15pt}
+\titlespacing*{\section}{0pt}{0\baselineskip}{0\baselineskip}
+\titlespacing*{\subsection}{0pt}{0\baselineskip}{0\baselineskip}
+\titlespacing*{\subsubsection}{0pt}{0\baselineskip}{0\baselineskip}
 
@@
 
-\renewcommand{\contentsname}{TABLE OF CONTENTS}
-\renewcommand{\listfigurename}{LIST OF FIGURES}
-\renewcommand{\listtablename}{LIST OF TABLES}
-\renewcommand{\abstractname}{ABSTRACT}
+\renewcommand{\contentsname}{\normalsize TABLE OF CONTENTS}
+\renewcommand{\listfigurename}{\normalsize LIST OF FIGURES}
+\renewcommand{\listtablename}{\normalsize LIST OF TABLES}
+\renewcommand{\abstractname}{\normalsize ABSTRACT}
+\renewcommand{\bibname}{\normalsize REFERENCES}
+
+\usepackage{tocloft}
+\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}}
+\renewcommand{\cftpartleader}{\cftdotfill{\cftdotsep}} 
+\renewcommand{\cftchapfont}{\normalfont}
+\renewcommand{\cftchappagefont}{\MakeUppercase\normalfont}
+\newlength\mylength
+\renewcommand\cftchappresnum{\MakeUppercase{\chaptername~}}
+\renewcommand\cftchapaftersnum{:}
+
+\setlength\cftaftertoctitleskip{20pt}
+\setlength\cftafterlottitleskip{15pt}
+\setlength\cftafterloftitleskip{15pt}
+
+\newcommand*{\noaddvspace}{\renewcommand*{\addvspace}[1]{}}
+\addtocontents{lof}{\protect\noaddvspace}
+\addtocontents{lot}{\protect\noaddvspace}
+
+\renewcommand{\cftfigpresnum}{Figure\ }
+\renewcommand{\cfttabpresnum}{Table\ }
+
+\cftsetindents{section}{0.5in}{0.3in}
+\cftsetindents{subsection}{1in}{0.4in}
+\cftsetindents{subsubsection}{1.5in}{0.5in}
+
+\setlength{\cfttabindent}{0in}
+\setlength{\cftfigindent}{0in}
+
+\newlength{\mylenf}
+\settowidth{\mylenf}{\cftfigpresnum}
+\setlength{\cftfignumwidth}{\dimexpr\mylenf+2em}
+\setlength{\cfttabnumwidth}{\dimexpr\mylenf+2em}
+
+\settowidth\mylength{\cftchappresnum\cftchapaftersnum}
+\addtolength\cftchapnumwidth{\mylength-0.5em}
 
@@
 
-Department of Computer Science and Engineering College of Engineering
+Department of Computer Science and Engineering
+
+College of Engineering
 
@@
 
+Srinivas Katkoori, Ph. D.
+
+Sameer Varma, Ph. D.
+
 \vspace*{2\baselineskip}
@@
 
-August 23, 2016
+October 10, 2016
 
@@
 \linespread{2}\selectfont
-\begin{flushleft}
+\justify
 \hspace{3em} I dedicate my thesis work to my family who have always been supporting me throughout my life.
 I also dedicate it to my closest friends who have shared my passion and encouraged me to explore and try my ideas in life.
-\end{flushleft}
 
@@
 \linespread{2}\selectfont
-\begin{flushleft}
+\justify
 \hspace{3em} I would like to thank my major professor throughout my master thesis Dr.
@@
 I as well want to thank the committee members for their support and encouragement.
-\end{flushleft}
+
+\linespread{1}\selectfont
+\vspace{1in}
+
+\frontmatter
+\centering
+\tableofcontents
 
 \clearpage
-\restoregeometry
-
-\linespread{1}\selectfont
-\frontmatter
-\tableofcontents
-
 \linespread{2}\selectfont
+\addcontentsline{toc}{chapter}{\listtablename} 
+\listoftables
+
+\clearpage
 \addcontentsline{toc}{chapter}{\listfigurename}
@@
 
-\addcontentsline{toc}{chapter}{\listtablename} 
-\listoftables
-
 \linespread{1}\selectfont
@@
 \justify
-\hspace{3em} Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
+\hspace{3em}Modern simulation systems generate big amount of data, which consequently has to be analyzed in a timely fashion.
 Traditional database management systems follow principle of pulling the needed data, processing it, and then returning the results.
@@
 When it comes to the point of doing various queries that require analysis  of the whole data,  this design has the following disadvantages: considerable overhead on traditional disk random I/O framework while reading from the simulation output files and low throughput of the data that consequently results in long latency, and, if there was any indexing to optimize selections, overhead of storing those becomes too big, too.
-Besides it, indexing will also cause delay during write operations and since most of the queries work with the entire data sets, indexing loses its point.
-
-There is a new approach to this problem -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and its primary steps are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filters the data and collect all the needed information which increases efficiency and data throughput.
+Beside it, indexing will also cause delay during write operations and since most of the queries work with the entire data sets, indexing loses its point.
+
+\noindent\hspace{3em}There is a new approach to this problem -- Push-based System for Molecular Simulation Data Analysis for processing network of queries proposed in the previous paper and its primary steps are: i) it uses traditional scan-based I/O framework to load the data from files to the main memory and then ii) the data is pushed through a network of queries which consequently filters the data and collect all the needed information which increases efficiency and data throughput.
 It has a considerable advantage in analysis of molecular simulation data, because it normally involves all the data sets to be processed by the queries.
 
-In this paper, we propose improved version of Push-based System for Molecular Simulation Data Analysis.
+\noindent\hspace{3em}In this paper, we propose improved version of Push-based System for Molecular Simulation Data Analysis.
 Its major difference with the previous design is usage of GPU for the actual processing part of the data flow.
@@
 
-%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
-%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%---%%%
-
 \mainmatter
 \linespread{1}\selectfont
-\chapter{Introduction} \linespread{2}\selectfont \justify
-
+\chapter{INTRODUCTION} 
+\linespread{2}\selectfont\justify
 \hspace{3em} In various sciences simulation systems take big place and often times they may be the clue for results.
@@
 
-The purpose of this paper is primarily focused on the third step, which basically involves the entire simulation data to be processed by the analysis software.
+\noindent\hspace{3em}The purpose of this paper is primarily focused on the third step, which basically involves the entire simulation data to be processed by the analysis software.
 It is important to underline scale of data being discussed.
@@
 
-Big data processing is becoming one of the key issues with the amount of data being generated by modern systems.
+\noindent\hspace{3em}Big data processing is becoming one of the key issues with the amount of data being generated by modern systems.
 Eventually, this data is required to be processed on the fly, thus, analysis software should be able to handle massive data in a very short period of time.
@@
 For example, it can take days and weeks to analyze big enough data sets, because the data cannot be simply loaded into memory, since it can get up to terabytes, thus, it has extra overhead because of traditional random access disk I/O framework to read from disk chunks by chunks.
-Besides it, analysis of the data can get even more complicated, since going through certain parts of the data once would not be enough, which leads to low throughput and efficiency of loaded data.
+Beside it, analysis of the data can get even more complicated, since going through certain parts of the data once would not be enough, which leads to low throughput and efficiency of loaded data.
 One such example, the data analysis approach has polynomial complexity just for reading the data in order to come up with result, and since the data can't simply be saved in memory, it raises the overhead of disk I/O, too.
@@
 
-As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks.
+\noindent\hspace{3em}As it has already been mentioned in the previous paper \cite{mainPaper}, one of the modern issues of analyzing massive data on the fly is social networks.
 "In order for a system to be able to perform analytical examination of the data produced in such streaming media, the system should have the capability of fast data access.
@@
 
-The primary focus and problem in this paper is scientific data analysis.
+\noindent\hspace{3em}The primary focus and problem in this paper is scientific data analysis.
 Examples brought up above are real life problems that are very similar to the scientific problem that we are focusing on in terms of data sizes.
@@
 
-Scientist gives the properties to simulation software (for example, Gromacs), runs the simulations, and finally get the output file.
+\noindent\hspace{3em}Scientist gives the properties to simulation software (for example, Gromacs), runs the simulations, and finally get the output file.
 The output file must then be analyzed to produce certain results which may help him come up with certain consensus on original theoretical model that resulted in the molecular simulation system\cite{Frenkel:api01}.
@@
 
-In this paper, we are incorporating parallelization into the processing part of the proposed design by means of CUDA programming language for GPU.
+\noindent\hspace{3em}In this paper, we are incorporating parallelization into the processing part of the proposed design by means of CUDA programming language for GPU.
 Since storage of the simulation data is very expensive, it might come to the point of analyzing the data on the fly (meaning running the simulation and analyzing it at the same time in a streaming manner), which leads to a problem of optimizing the processing part of the design proposed in the previous paper, because time spent on generating data should be tried to conceal the time spent on the processing part by means of overlapping or simply running it in a quick manner.
@@
 The problem is still that some of the queries processed by those means are still improvable, especially taking into consideration the fact that in the used previous works for calculation of 2-body functions, which take the biggest time for processing~\cite{ytu:icde09, EDBT12}, we might have some error bounds, which might be unacceptable for certain simulation analysis where sacrifice on loss of data is unbearable.
-Besides it, although we might have huge performance boost on specific data structure as density map, we still have certain expectation from data nature (such as its uniformity), thus, it makes sense to have desire to simplify the processing and still having pure computation based on any values, considering that the queries should be available to be pre-programmed by user in a separate module of the code.
+Beside it, although we might have huge performance boost on specific data structure as density map, we still have certain expectation from data nature (such as its uniformity), thus, it makes sense to have desire to simplify the processing and still having pure computation based on any values, considering that the queries should be available to be pre-programmed by user in a separate module of the code.
 All the benchmarks will be done in comparison with already push-based system versus push-based system with GPU improvements.
@@
 
-\section{Our approach with improvement}
+\section{Our Approach With Improvement}
 
@@
 
-Original non-improvement standard approach with pull-based querying is probably one of the most flexible, because nobody has to think about improvements and complexities, it does it for every chunk of data and pulls it whenever it is needed.
+\noindent\hspace{3em}Original non-improvement standard approach with pull-based querying is probably one of the most flexible, because nobody has to think about improvements and complexities, it does it for every chunk of data and pulls it whenever it is needed.
 As it has already been mentioned, at this point, the biggest challenge is that supposing we have $M$ queries, for $N$ atoms we would have to load every chunk $M$ times for $N$ chunks of data (frames), and apparently it is very time consuming of I/O and non-linear computations, and that is only for linear functions.
@@
 
-After all, push-based system for solution of the given problem is tremendously fast in comparison with pull-based system.
+\noindent\hspace{3em}After all, push-based system for solution of the given problem is tremendously fast in comparison with pull-based system.
 General idea is to push a chunk of data (frame) through the pre-defined network of queries that will be needed to run on the data sets, which basically means chunk of data will be read from file and loaded into memory only once in order to go through all the queries.
@@
 
-Finally, we've analyzed some main ideas, advantages, and disadvantages of our approach in this paper which is basically push-based system with optimization on data processing part by means of GPU computation (in other words, parallelism).
+\noindent\hspace{3em}Finally, we've analyzed some main ideas, advantages, and disadvantages of our approach in this paper which is basically push-based system with optimization on data processing part by means of GPU computation (in other words, parallelism).
 Reading the data from files is a huge overhead, thus, speedup of push-based system in general is significant over pull-based system in our given case.
@@
 
-Some of the queries may be very slow due to their nature on a sequential type of computation,  especially, 2-body functions.
+\noindent\hspace{3em}Some of the queries may be very slow due to their nature on a sequential type of computation,  especially, 2-body functions.
 Since the nature of the data that comes with simulation is basically physical properties of atoms, there is a lot of computation that involves independent primitive mathematical operations, which makes it a perfect problem for parallelism.
@@
 
-As it has already been mentioned, although the original idea of push-based system for molecular simulation data analysis was focused on optimizing throughput and usage of loaded data, there were some extra approaches specifically for 2-body functions.
+\noindent\hspace{3em}As it has already been mentioned, although the original idea of push-based system for molecular simulation data analysis was focused on optimizing throughput and usage of loaded data, there were some extra approaches specifically for 2-body functions.
 For example, Density Map for Spacial Distance Histogram was used in order to avoid additional memory allocation and latency reduction with proven error bounds.
@@
 
-We believe that having this nature of computational problems, the proposed GPU improved version of push-based system will significantly change in terms of performance by incorporating parallelism with CUDA.
+\noindent\hspace{3em}We believe that having this nature of computational problems, the proposed GPU improved version of push-based system will significantly change in terms of performance by incorporating parallelism with CUDA.
 
@@
 
-\section{Contribution and roadmap of the paper}
+\section{Contribution And Roadmap Of The Paper}
 
@@
 
-Taking an advantage of GPU devices nature and incorporating parallelism and streaming for different types of queries, we have come up with a good speed up over sequential processing which gets up to 60 times faster, which is just for sampling with no deep improvements on individual queries (since we believe that it should represent average user usage).
+\noindent\hspace{3em}Taking an advantage of GPU devices nature and incorporating parallelism and streaming for different types of queries, we have come up with a good speed up over sequential processing which gets up to 60 times faster, which is just for sampling with no deep improvements on individual queries (since we believe that it should represent average user usage).
 Original push-based system already had a good speed up in terms of performance in comparison with original pull-based approach and floats around more than 100 times faster depending on number of atoms, frames, and workload.
@@
 
-\chapter{Related Work}\label{sc:relatedwork}
+\chapter{RELATED WORK}\label{sc:relatedwork}
 
@@
 
-\section{Scientific software}
+\section{Scientific Software}
 
@@
 
-\section{Stream processing}
+\section{Stream Processing}
 \hspace{3em} Another major modern issue is live data streaming and live processing, which is also possible with our system.
@@
 
-As a plugin for optimizing computation of scientific queries against large data sets being produced, Push-based System for Molecular Simulation Data Analysis with GPU improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
+\noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets being produced, Push-based System for Molecular Simulation Data Analysis with GPU improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
 
@@
 
-\section{GPU and Big Data Processing}
+\section{GPU And Big Data Processing}
 \hspace{3em} GPU is becoming more popular being used by not only graphics processing, but machine learning, deep learning, search algorithms, sorting, and so on.
@@
 
-Push-based system for MS data analysis is primarily proposing an idea of pushing chunks of data through a network of queries.
+\noindent\hspace{3em}Push-based system for MS data analysis is primarily proposing an idea of pushing chunks of data through a network of queries.
 Essentially, this has also been proposed, as mentioned already, by other frameworks \cite{DataPath,Volcano,Qpipe}.
@@
 
-Since dealing with data is very delicate in this kind of engines, there are couple of features and requirements that this architecture is applicable to, in order to avoid inappropriate understanding or application of this approach in systems that were covered in the lectures\cite{DataStreamManagement}.
+\noindent\hspace{3em}Since dealing with data is very delicate in this kind of engines, there are couple of features and requirements that this architecture is applicable to, in order to avoid inappropriate understanding or application of this approach in systems that were covered in the lectures\cite{DataStreamManagement}.
 The data is volatile and not persistent anymore because of obvious sizes of data.
@@
 
-This version of Push-based System of MS data analysis is primarily about taking an advantage of GPU processing of the data.
+\noindent\hspace{3em}This version of Push-based System of MS data analysis is primarily about taking an advantage of GPU processing of the data.
 Hadoop is a well-known and well-used framework by many companies nowadays\cite{hadoop}.
@@
 
-Gromacs generate output files with physical properties description for each atom in each frame.
+\noindent\hspace{3em}Gromacs generate output files with physical properties description for each atom in each frame.
 Modern CPUs do great job in terms of caching and computing, but obviously in this particular problem massive monolithic computation is needed, thus, GPU devices are perfect candidates for such a problem following SIMD type of computation.
@@
 
-\chapter{Benchmark scientific queries}\label{sc:querynetwork}
+\chapter{BENCHMARK SCIENTIFIC QUERIES}\label{sc:querynetwork}
 \hspace{3em} Having generated big amount of data in large files, now scientists need to get useful information out of it by means of analysis utilities.
@@
 
-In previous work, there has already been developed a network of queries widely used by scientists for MS systems.
+\noindent\hspace{3em}In previous work, there has already been developed a network of queries widely used by scientists for MS systems.
 In Table~\ref{tb:queries} you may observe the common set of queries that has been developed.
@@
 
-The queries that we consider are related to computations limited within one frame of a molecular simulation.
+\noindent\hspace{3em}The queries that we consider are related to computations limited within one frame of a molecular simulation.
 In general, with the given queries we have divided them into two categories: one body functions and multiple body functions.
@@
 
-\section{One body functions}
+\section{One Body Functions}
 \hspace{3em} \emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
 
-As you can see in the given table, most of them are basically based on primitive summation or other basic mathematical operations, in our experiments in further sections we will be running result on single precision values or to be more explicit we use float type in C programming language.
-
-\section{Multiple body functions}
-\emph{\textbf{Multiple body functions}} such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
+\noindent\hspace{3em}As you can see in the given table, most of them are basically based on primitive summation or other basic mathematical operations, in our experiments in further sections we will be running result on single precision values or to be more explicit we use float type in C programming language.
+
+\section{Multiple Body Functions}
+\noindent\hspace{3em}\emph{\textbf{Multiple body functions}} such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
 This is generally combination of two across N data: 	$C\binom{N}{2}$.
@@
 
-For computation of histogram it is not enough to visit every value for each atom once, since we have to do pairwise computations, thus it increases complexity of the problem not only computation-wise, but memory-wise as well, which is a bigger concern in our case, since the overhead will be huge and non-linearly relative to the size of data.
+\noindent\hspace{3em}For computation of histogram it is not enough to visit every value for each atom once, since we have to do pairwise computations, thus it increases complexity of the problem not only computation-wise, but memory-wise as well, which is a bigger concern in our case, since the overhead will be huge and non-linearly relative to the size of data.
 For this purpose we have come up with a way to optimize the way we use references to memory for each atom, so that we minimize overhead of pulling data all the way from DRAM (GPU).
 
-There might be some dependencies and preliminary computation possibilities.
+\noindent\hspace{3em}There might be some dependencies and preliminary computation possibilities.
 For example, some queries need total mass, and we could compute it before we push it further, but this can be easily done by user based on the complexity and dependency.
@@
 
-\vspace{3mm}
-\chapter{The system design}\label{sc:system}
-\vspace{3mm}
+\chapter{THE SYSTEM DESIGN}\label{sc:system}
 
@@
 
-\section{Data input}
+\section{Data Input}
 \hspace{3em} Since the size of the data in the problem is huge, organization of the data flow is crucial.
@@
 
-In order to experiment with different input cases, the input files have been generated with a tool that we have built in Python language.
+\noindent\hspace{3em}In order to experiment with different input cases, the input files have been generated with a tool that we have built in Python language.
 Thus, we are able to generate mock data files with any number of frames and atoms, and the values are random and can be adjusted with needs of the experiment (for example, for cases when we want to analyze bucket widths in a limited space, the coordinates are generated in range of $0 - 1$, etc.).
@@
 
-\section{Data flow architecture}
+\section{Data Flow Architecture}
 
@@
 
-Having defined the chunk, we load it from flat format file to RAM.
+\noindent\hspace{3em}Having defined the chunk, we load it from flat format file to RAM.
 A single atom data structure has all the attributes such as charge, mass, velocity, type of atom, coordinates, and so on.
@@
 
-To summarize, the data goes through several levels of memory hierarchy in our system.
+\noindent\hspace{3em}To summarize, the data goes through several levels of memory hierarchy in our system.
 Coming as an input from either streaming tool, flat file (our case), or other type of input (f.e.
@@
 
-\section{Queries organization}
-
-Since we assume that our framework should be easily used by a scientist doing MS analysis, it is important to keep queries module as a separate piece of code to make it easily extendable.
+\section{Queries Organization}
+
+\noindent\hspace{3em}Since we assume that our framework should be easily used by a scientist doing MS analysis, it is important to keep queries module as a separate piece of code to make it easily extendable.
 In order to understand how queries module is organized and the reason for it, we need to understand how CUDA and GPU work.
 
-\emph{\textbf{GPU execution}}.
+\noindent\hspace{3em}\emph{\textbf{GPU execution}}.
 From hardware standpoint, it is important to understand that every GPU device consists of SMs (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
@@
 
-\emph{\textbf{One-body queries kernel}}.
+\noindent\hspace{3em}\emph{\textbf{One-body queries kernel}}.
 Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
@@
 
-\emph{\textbf{Two-body queries kernel}}.
+\noindent\hspace{3em}\emph{\textbf{Two-body queries kernel}}.
 SDH (or RDF) is the algorithm that requires special treatment in our design.
@@
 
-\emph{\textbf{Architecture overview.}}
+\noindent\hspace{3em}\emph{\textbf{Architecture overview.}}
 This section is to emphasize general architecture of the system and summarize the work flow as well as the data flow.
@@
 
-\emph{\textbf{GPU grid and block sizes.}} As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
+\noindent\hspace{3em}\emph{\textbf{GPU grid and block sizes.}} As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
 The approach we came up with is 1 execution thread per atom.
@@
 
-\section{Computation work flow}
+\section{Computation Work Flow}
 
 \hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
-Besides it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-
-\subsection{One-body functions implementation}
+Beside it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
+
+\subsection{One-body Functions Implementation}
 
@@
 Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.
-
-\vspace{1cm}
 
@@
 
-\subsubsection{One-body functions implementation in CPU version}
+\subsubsection{One-body Functions Implementation In CPU Version}
 
@@
 
-We have provided Pseudocode \ref{alg:onebodycpu} of the implementation of one body functions for CPU, which has detailed step-by-step description of the CPU sequential version of push-based system for calculating one-body functions.
-
-\subsubsection{One-body functions implementation in GPU version}
+\noindent\hspace{3em}We have provided Pseudocode \ref{alg:onebodycpu} of the implementation of one body functions for CPU, which has detailed step-by-step description of the CPU sequential version of push-based system for calculating one-body functions.
+
+\subsubsection{One-body Functions Implementation In GPU Version}
 
@@
 
-As you can see from the given algorithm (Pseudocode \ref{alg:onebodygpu}), in the thread kernel function it starts with allocation of memory in the block.
+\noindent\hspace{3em}As you can see from the given algorithm (Pseudocode \ref{alg:onebodygpu}), in the thread kernel function it starts with allocation of memory in the block.
 To be more explicit, every block has its own level of memory called shared memory.
@@
 Once a block is finished, first thread of each block will accumulate data from shared memory into global synchronizing the results, every write operation to shared and global memories is atomic (we use native CUDA function atomicAdd which guarantees atomicity).
-
-\vspace{1cm}
 
@@
 
-\subsection{Two-body functions implementation}
+\subsection{Two-body Functions Implementation}
 
@@
 
-\subsubsection{Two-body functions implementation in CPU Version}
+\subsubsection{Two-body Functions Implementation In CPU Version}
 
@@
 Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them, though results don't change a lot for sequential computation.
-
-\vspace{1cm}
 
@@
 
-\subsubsection{Two-body functions implementation in GPU Version}
+\subsubsection{Two-body Functions Implementation In GPU Version}
 
@@
 
-Here are the main steps of the algorithm:
+\noindent\hspace{3em}Here are the main steps of the algorithm:
 \begin{itemize}
@@
 
-As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that user's wouldn't do it and having more or less realistic speedup results.
+\noindent\hspace{3em}As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that user's wouldn't do it and having more or less realistic speedup results.
 Figure \ref{fg:sharedMem} shows swimlanes as execution thread for blocks.
@@
 
-\vspace{1cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Data flow implementation in CPU}
+\subsection{Data Flow Implementation In CPU}
 
 \hspace{3em} As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
-
-\vspace{1cm}
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\vspace{1cm}
-
-In the Pseudocode \ref{alg:cpuflow} you can see rough logic of our CPU mapping function that primarily allocates all the memory in GPU, triggers kernel function and manipulates streams.
+\noindent\hspace{3em}In the Pseudocode \ref{alg:cpuflow} you can see rough logic of our CPU mapping function that primarily allocates all the memory in GPU, triggers kernel function and manipulates streams.
 This is a prototype which definitely is not the last version, because we can as well avoid overhead of copying data into shared memory while synchronously it will be computing data already contained there.
@@
 We don't expect them to run really concurrently, because our two kernels are quite heavily loaded, though streaming can definitely be use for memory management optimization.
-
-\vspace{1cm}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\vspace{1cm}
-
-The above Pseudocode \ref{alg:cpuflowfuture} shows what memory concurrent could look like.
+\noindent\hspace{3em}The above Pseudocode \ref{alg:cpuflowfuture} shows what memory concurrent could look like.
 As you can see it is pretty much overlapping $nextAtoms$ loading with computation of $atoms$, which reduces this tiny overhead of copy data over.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Experiments Results}\label{sc:experiments}
-
+\chapter{EXPERIMENTS RESULTS}\label{sc:experiments}
 \hspace{3em} The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
@@
 
-\section{Hardware stack}
-
+\section{Hardware Stack}
 \hspace{3em} There are primarily several parts that are really important in this hardware stack.
@@
 
-\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
+\noindent\hspace{3em}\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
 More improvements can be done for acceleration of this approach, but in the system we have not gone too far with data converter module, because initial framework has to be simple and user friendly.
@@
 
-\subsection{Experiment results for 1000 atoms}
+\subsection{Experiment Results For 1000 Atoms}
 
@@
 
-\vspace{1cm}
 \begin{table}[h!]
@@
 
-On Table \ref{tb:gpuprofile1000} you can observe results of Nvidia Profiling Tool of how much memory call we did to shared memory for SDH computation.
+\noindent\hspace{3em}On Table \ref{tb:gpuprofile1000} you can observe results of Nvidia Profiling Tool of how much memory call we did to shared memory for SDH computation.
 You can also notice that there were 32 warps launched for computing 1000 atoms which is absolutely reasonable, since as we already mentioned one warp (scheduling unit for ultimate parallel execution) has 32 threads maximum, thus ${\frac{1000}{32}}$ \textasciitilde \space$32$.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment results for 10,000 atoms}
+\subsection{Experiment Results For 10,000 Atoms}
 
@@
 Fortunately, scientists usually work with much larger data sets, thus, this doesn't really represent general speedup for average use case.
-
-\vspace{1cm}
 \begin{table}[h!]
@@
 
-As you can see, speedup bumps up to almost 30 with 10000 atoms.
+\noindent\hspace{3em}As you can see, speedup bumps up to almost 30 with 10000 atoms.
 Once the overhead on memory transfer is covered by computation speedup, it keeps growing.
@@
 
-\subsection{Experiment results for 100,000 atoms}
+\subsection{Experiment Results For 100,000 Atoms}
 
@@
 
-Fortunately, scientific data is much bigger than above examples.
+\noindent\hspace{3em}Fortunately, scientific data is much bigger than above examples.
 You can see average speedup for bigger number of atoms on Figure \ref{fg:chart_max}.
@@
 
-Finally, for 100,000 atoms the warps launched for execution seems to be growing reasonable along with number of atoms (Table \ref{tb:gpuprofile100000}), while on the other hand usage of shared load for SDH computation is extremely high.
-
-\subsection{Advanced SDH implementation}
+\noindent\hspace{3em}Finally, for 100,000 atoms the warps launched for execution seems to be growing reasonable along with number of atoms (Table \ref{tb:gpuprofile100000}), while on the other hand usage of shared load for SDH computation is extremely high.
+
+\subsection{Advanced SDH Implementation}
 
@@
 
-Basically, as you can see from Figure \ref{fg:improved_sdh}, now we allocate more shared memory.
+\noindent\hspace{3em}Basically, as you can see from Figure \ref{fg:improved_sdh}, now we allocate more shared memory.
 Yellow sections are the atoms that are stored in shared memory throughout the whole thread execution of the entire block, since those atoms are needed at every step, while the green ones are the working zones.
@@
 
-As you may have noticed, even though we increased utilization of shared memory, number of loads is around the same.
+\noindent\hspace{3em}As you may have noticed, even though we increased utilization of shared memory, number of loads is around the same.
 The trick is that we replaced atoms in shared memory with simpler data structure called $coordinates$ that would contains 3 float type variables which saves huge amount of memory.
@@
 
-\subsection{Overall performance}
+\subsection{Overall Performance}
 
@@
 One of possible reasons could the fact that we've used static approach of computing grid dimensions.
-Besides it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
+Beside it, in the future, it can be expanded into a dynamic version that will automatically adjust the grid dimensions choosing least overhead that it has to go with.
 
@@
 
-With improved version of SDH speedup gets up to more than 75 times (Figure \ref{fg:chart_max}).
+\noindent\hspace{3em}With improved version of SDH speedup gets up to more than 75 times (Figure \ref{fg:chart_max}).
 We believe that these numbers could get even better depending on queries and elaboration of improvement of particular cases.
 
-\chapter{Conclusions and Future Work}\label{sc:conclusion}
+\chapter{CONCLUSION}\label{sc:conclusion}
 \hspace{3em} The ultimate goal of this work was to get an improved version of Push-based system for molecular simulation data analysis.
@@
 
-%\bibliographystyle{plain}
+\linespread{1}\selectfont
+\addcontentsline{toc}{chapter}{\bibname}
 \bibliographystyle{IEEEtran}

### 2016-10-09 thesis_paper/source.tex
automatic triggered changes

@@
 
+\usepackage{float}
+\floatstyle{plaintop}
+\restylefloat{table}
+
 \makeatletter
@@
 \linespread{2}\selectfont\justify
-\hspace{3em} In various sciences simulation systems take big place and often times they may be the clue for results.
+\hspace{3em} In various sciences simulation systems are very crucial and might be the key factors for results.
 One of such sciences, which is primarily related to this paper, is physics.
@@
 It is important to underline scale of data being discussed.
-Molecular simulations represent systems starting from hundreds thousands of atoms and more which are tracked by frames as snapshot of specific time and can last for months.
+Molecular simulations represent systems starting from thousands of atoms and more which are tracked by frames as snapshot of specific time and can last for months.
 Thus, the size of data starts from couple of gigabytes up to terabytes and petabytes.
@@
 Eventually, this data is required to be processed on the fly, thus, analysis software should be able to handle massive data in a very short period of time.
-When working with huge volume of data, there are non-trivial issues arise.
+When working with huge volume of data, there can be additional issues while using regular tools.
 For example, it can take days and weeks to analyze big enough data sets, because the data cannot be simply loaded into memory, since it can get up to terabytes, thus, it has extra overhead because of traditional random access disk I/O framework to read from disk chunks by chunks.
@@
 nanoscopic scale." \cite{mainPaper}
+\clearpage
 
@@
 Since storage of the simulation data is very expensive, it might come to the point of analyzing the data on the fly (meaning running the simulation and analyzing it at the same time in a streaming manner), which leads to a problem of optimizing the processing part of the design proposed in the previous paper, because time spent on generating data should be tried to conceal the time spent on the processing part by means of overlapping or simply running it in a quick manner.
-We will talk about this in details as well in further sections.
+We will talk about this in detail as well in further sections.
 
@@
 The approach proposed in the previous paper is very good in terms of performance in comparison with the original pull-based system\cite{mainPaper}.
-As you can see on Figure \ref{fg:sample_estimation_old_paper}, it gets up to over 100 times of speedup which is quite crucial, if we think about days of computations.
+It gets up to over 100 times of speedup which is quite crucial, if we think about days of computations.
 The problem is still that some of the queries processed by those means are still improvable, especially taking into consideration the fact that in the used previous works for calculation of 2-body functions, which take the biggest time for processing~\cite{ytu:icde09, EDBT12}, we might have some error bounds, which might be unacceptable for certain simulation analysis where sacrifice on loss of data is unbearable.
@@
 This tool can be integrated into open source project of Gromacs as a custom plugin that can be used for optimized analytic queries to be run over data sets live streaming as well as from files.
-This will be discussed in more details in further chapters.
+This will be discussed in more detail in further chapters.
 The structure of the framework which will be described later is developed in such a way that one can easily add new queries based on their complexity or modify existing ones adding appropriate selections and filters.
@@
 Borealis Stream Processing Software is a tool being developed by MIT, Brown, and Brandeis Universities.
-It's primary goal is a distributed stream processing and based on Medusa and Aurora.
+Its primary goal is a distributed stream processing and based on Medusa and Aurora.
 Some of the major functions of the software are highly available distributed large streams processing live, dynamic and highly-scalable with possibility of changing queries live and dynamic vision of query results.
@@
 
-\noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets being produced, Push-based System for Molecular Simulation Data Analysis with GPU improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
+\noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets, an improved Push-based System for MS improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
 
@@
 
-\vspace{3em}
-
 \begin{table}[h]
@@
 \section{One Body Functions}
-\hspace{3em} \emph{\textbf{One body functions}} such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
+\hspace{3em} One body functions such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
 
@@
 \section{Multiple Body Functions}
-\noindent\hspace{3em}\emph{\textbf{Multiple body functions}} such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
+\noindent\hspace{3em}Multiple body functions such as SDH and RDF (Spacial Distance Histogram and Radial Distribution function) require computation of distances pairwise across all the particles.
 This is generally combination of two across N data: 	$C\binom{N}{2}$.
@@
 For example, some queries need total mass, and we could compute it before we push it further, but this can be easily done by user based on the complexity and dependency.
-It is explained in more details in further sections.
+It is explained in more detail in further sections.
 
@@
 
-\noindent\hspace{3em}\emph{\textbf{GPU execution}}.
-From hardware standpoint, it is important to understand that every GPU device consists of SMs (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
+\noindent\hspace{3em} From hardware standpoint, it is important to understand that every GPU device consists of SMs (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
 Having this in mind, since this is a pure parallel processing power, the goal is to increase hardware throughput and keep highest utilization to exploit every register given.
@@
 
-\noindent\hspace{3em}\emph{\textbf{One-body queries kernel}}.
-Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
+\noindent\hspace{3em}Commonly used CUDA array reduction\cite{gpureduction} makes the implementation of all the one-body queries obvious.
 Figure \ref{fg:reduction} represents general idea of the data mapping, where each cell of the vector can be associated with a thread of execution in the kernel.
@@
 
-\noindent\hspace{3em}\emph{\textbf{Two-body queries kernel}}.
-SDH (or RDF) is the algorithm that requires special treatment in our design.
+\noindent\hspace{3em}SDH (or RDF) is the algorithm that requires special treatment in our design.
 The point, as mentioned earlier, is that this function requires more than one iteration of the given set of atoms, but rather all combinations of pairs in a given frame, thus, it is not linear complexity.
@@
 
-\noindent\hspace{3em}\emph{\textbf{Architecture overview.}}
-This section is to emphasize general architecture of the system and summarize the work flow as well as the data flow.
+\noindent\hspace{3em}This section is to emphasize general architecture of the system and summarize the work flow as well as the data flow.
 In total, we have flat files with ready flat values for each atom aggregated in frames coming into the program as data input using common disk I/O framework of an operating system.
@@
 Just for architecture clearance and possible future improvements, two kernels are executed in two different streams\footnote{In CUDA stream is another abstract layer of organizing execution threads.
-They are not guaranteed to be executed at the same time or after each other, GPU scheduling architecture is completely different from CPU having a warp to be single scheduling unit at maximum of 32 threads at a time} (You can see them on Figure\ref{fg:overallarchitecture}).
+They are not guaranteed to be executed at the same time or after each other, GPU scheduling architecture is completely different from CPU having a warp to be single scheduling unit at maximum of 32 threads at a time} (You can see them on Figure \ref{fg:overallarchitecture}).
 In each kernel we apply certain heuristics to improve memory access using such architecture features of GPU as shared memory and blocking of kernel threads.
 
-\noindent\hspace{3em}\emph{\textbf{GPU grid and block sizes.}} As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
+\noindent\hspace{3em}GPU grid and block sizes.
+As it has already been mentioned, utilization of hardware on GPU device is important, because it is the key to performance improvement, and the higher utilization is, the bigger improvement is.
 The approach we came up with is 1 execution thread per atom.
@@
 
-\section{Computation Work Flow}
-
 \hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
 Beside it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
-
-\subsection{One-body Functions Implementation}
 
@@
 
-\subsubsection{One-body Functions Implementation In CPU Version}
-
-\hspace{3em} Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
+\noindent\hspace{3em} Basically being given a vector of atoms in CPU version implementation (the one that we use for speedup benchmark) we loop through every atom and accumulate all the needed information to RAM.
 Once we have the result for one-body functions, we take it out for further usage, it can as well be used for further calculations where there are dependencies on such things as total mass, total charge, etc.
@@
 
-\subsubsection{One-body Functions Implementation In GPU Version}
-
-\hspace{3em} The approach in GPU version of one-body computation code is a lot different, because there are two major changes: the way we deal with the data in memory and mapping of GPU thread kernels to atoms.
+\noindent\hspace{3em} The approach in GPU version of one-body computation code is a lot different, because there are two major changes: the way we deal with the data in memory and mapping of GPU thread kernels to atoms.
 Both of these approaches drastically change the way we perceive the solution implementation.
@@
 
-\State $\textit{sharedResult} \gets \textit{pointer to shared memory head (in this case declared element size as float)}$
+\State $\textit{sharedResult} \gets \textit{pointer to shared memory head}$
 
@@
 \State
-\State $\textit{// Here we will starting collecting information from each block which will be done by first thread of each block}$
+\State $\textit{// Here we will starting collecting information from each block}$
 \If {$\textit{threadIdx.x} = 0$}
-\State  $\textit{atomicAdd(\&results->mass, sharedResult[0]);}$ \Comment{All atomicAdd functions here are GPU atomic}
+\State  $\textit{atomicAdd(\&results->mass, sharedResult[0]);}$ \Comment{All atomic functions here}
 \State  $\textit{atomicAdd(\&results->charge, sharedResult[1]);}$
@@
 
-\subsection{Two-body Functions Implementation}
-
-\hspace{3em} Since one of the physical queries is RDF which has underlying SDH calculation, we will describe our approach with SDH for CPU and GPU versions.
-
-\subsubsection{Two-body Functions Implementation In CPU Version}
-
-\hspace{3em} Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
+\noindent\hspace{3em} Since one of the physical queries is RDF which has underlying SDH calculation, we will describe our approach with SDH for CPU and GPU versions.
+
+\noindent\hspace{3em} Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
 Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them, though results don't change a lot for sequential computation.
@@
 
-\subsubsection{Two-body Functions Implementation In GPU Version}
-
-\hspace{3em} The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same.
+\noindent\hspace{3em} The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same.
 We do brute force  $O(\frac{N^2}{2})$ complex solution which is pretty much looking at distance between every pair of atoms.
@@
 
-\noindent\hspace{3em}As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that user's wouldn't do it and having more or less realistic speedup results.
+\noindent\hspace{3em}As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that users wouldn't do it and having more or less realistic speedup results.
 Figure \ref{fg:sharedMem} shows swimlanes as execution thread for blocks.
@@
 
-\subsection{Data Flow Implementation In CPU}
-
-\hspace{3em} As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
+\noindent\hspace{3em} As we have described in earlier sections, our CPU is primarily used for data flow management, triggering of GPU functions, data structure changes, and GPU computation stream manipulations.
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The main steps here are the following.
-we create to instances of CUDA streams that are again abstract data types that are scheduled by GPU in a special way.
+We create 2 instances of CUDA streams that are again abstract data types that are scheduled by GPU in a special way.
 We as well allocate memory in GPU for atoms list, histogram results and common results for one-body queries (in this case it is a data structure with fields that represent some accumulated result).
 Once we are setup with memory management on GPU, we trigger kernel calls one by one each running on a separate CUDA stream.
-We don't expect them to run really concurrently, because our two kernels are quite heavily loaded, though streaming can definitely be use for memory management optimization.
+We don't expect them to run really concurrently, because our two kernels are quite heavily loaded, though streaming can definitely be used for memory management optimization.
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 \chapter{EXPERIMENTS RESULTS}\label{sc:experiments}
-\hspace{3em} The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
+\noindent\hspace{3em} The framework has been developed in a programming language CUDA, and the sequential version for benchmark\cite{mainPaper} has been originally developed in C++.
 Experiments were run on operating systems which didn't have any additional workload besides usual installation, and the operating systems particularly were Amazon Linux, which doesn't have a lot of modifications except for some additional features for management with Amazon EC-2 Console.
@@
 \section{Hardware Stack}
-\hspace{3em} There are primarily several parts that are really important in this hardware stack.
+\noindent\hspace{3em} There are primarily several parts that are really important in this hardware stack.
 Since this is primarily big data processing, we care about processing power and memory, Of course, things like PCI bus and all the transportation are important as well, but for speedup benchmarks against CPU vs GPU those stay the same for both sides.
@@
 
-\subsection{CPU}
-\hspace{3em} We have summarized main CPU parameters for CPU version and it is described in Table \ref{tb:cpuhardware}.
+\noindent\hspace{3em} We have summarized main CPU parameters for CPU version and it is described in Table \ref{tb:cpuhardware}.
 As you may notice, it is a regular server node CPU with 2.60GHz and 3 levels of L cache.
@@
 
-\subsection{GPU}
-
-\hspace{3em} The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
+\noindent\hspace{3em} The same CPU has been used for sequential as well as for improved GPU version during experiments with benchmarks.
 So, for CPU in GPU improved version refer to the same Table \ref{tb:cpuhardware}.
@@
 First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests.
-This possibility will be elaborated in more details in Future Work section.
-
-\noindent\hspace{3em}\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
+This possibility will be elaborated in more detail in Future Work section.
+
+\noindent\hspace{3em}SDH implementation.
+Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.
 More improvements can be done for acceleration of this approach, but in the system we have not gone too far with data converter module, because initial framework has to be simple and user friendly.
@@
 You can see average speedup for bigger number of atoms on Figure \ref{fg:chart_max}.
-The speedup gets upto 70 times which could save huge amount of time for scientists.
+The speedup gets up to 70 times which could save huge amount of time for scientists.
 These results are not the best, as it has been already mentioned earlier, since we have not modified general structure of memory being transferred to Global Memory not taking advantage of coalesced memory access to make it easily extendable.

### 2016-10-12 thesis_paper/source.tex
Formatted some thesis paper parts

@@
 
-Srinivas Katkoori, Ph. D.
-
-Sameer Varma, Ph. D.
+Srinivas Katkoori, Ph.D.
+
+Sameer Varma, Ph.D.
 
@@
 
-October 10, 2016
+August 23, 2016
 
@@
 The proposed improved version of push-based system for molecular simulation data analysis, we believe, gives an opportunity for scientists to run their analysis fast having an ability to expand network of queries and customize computation functions to get the desired results.
-Our system primarily represents a system that works in several modules that were developed in C, CUDA, Python programming languages that, in combination, give us the desired result to perform efficient computations for the network of queries developed in the previous paper \cite{mainPaper} as well as to prove the concept of optimization of push-based system by means of parallelism with no data loss.
-We've developed the following modules:
+Our system primarily represents a system that works in several modules that were developed in C, CUDA, Python programming languages that, in combination, give us the desired result to perform efficient computations for the network of queries developed in the previous paper \cite{mainPaper} as well as to prove the concept of optimization of push-based system by means of parallelism with no data loss:
 
@@
 \noindent\hspace{3em}As a plugin for optimizing computation of scientific queries against large data sets, an improved Push-based System for MS improvement should be able to handle data streams as well, in order to get real time results and saving up on time, instead of pulling everything from files later on.
+For now it has been developed as a separate program that runs independently accepting input in a specified format which is described in the paper, but ideally it is supposed to be embedded right into Gromacs.
 
 \section{Big Data Frameworks}
-
 \hspace{3em} We are potentially dealing with data that could easily get up to terabytes and petabytes in sizes, and it will be fair to apply term of big data.
@@
 
-\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
+\noindent\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
 Beside it, we allocate some memory in GPU global memory for results, too, and keep it as an object with fields to easily pull it back when it is ready, in our case all the data sizes are static before the kernels are executed, which follows nature of GPU devices.
 
-\hspace{3em} As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
+\noindent\hspace{3em} As it has already been mentioned in previous sections, one body functions are relatively easy to implement, since they don't require iteration through data more than once with a write approach.
 Our implementation of one-body functions are relatively easy to understand which is what we tried to explain further.
@@
 For recall, CPU in improved version is used in-between memory loading from file and uploading it to GPU, printing the results to the user, which is not big load.
-Consequently, we also have GPU device.
-Hardware summary for GPU device is described in Table \ref{tb:gpuhardware}
+\clearpage
+Consequently, we also have GPU device with its own performance capacities. Hardware capacities summary for GPU device has been described in Table \ref{tb:gpuhardware}
 
@@
 
-\hspace{3em} We have run number of experiments with different number of atoms and frames.
+\noindent\hspace{3em} We have run number of experiments with different number of atoms and frames.
 Results vary as we choose different data sets, but we can see some tendency towards growth and, fortunately, it is very positive for our improved version of push-based system.
 
-\subsection{Experiment Results For 1000 Atoms}
-
-\hspace{3em} Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes.
+\noindent\hspace{3em} Even though we assume that majority of our cases involve huge amount of atoms (starting from hundred thousands), we still ran experiments on smaller amount of atoms to see difference in speedup relatively with the data set sizes.
 You can observe summary in Figures \ref{fg:exp_1000_10}, \ref{fg:exp_1000_100}, \ref{fg:exp_1000_1000}, \ref{fg:speedup_1000}.
@@
 Once there is more load on computation, strongly GPU accelerates and becomes faster.
+
+\clearpage
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Experiment Results For 10,000 Atoms}
-
 \begin{figure}[!tbp]
@@
 
-\subsection{Experiment Results For 100,000 Atoms}
-
 \hspace{3em} When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good.
@@
 
-\subsection{Advanced SDH Implementation}
-
-\hspace{3em} We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.
+\noindent\hspace{3em} We implemented more advanced version of SDH algorithm taking better use of shared memory just to demonstrate what improvement and speedup could be achieved if we focus on particular query implementation.
 
@@
  \centerline{ \includegraphics[width=0.5\columnwidth]{images/sdhvsimproved}}
- \caption{Comparison of GPU SDH with GPU SDH Improved}
+ \caption{Comparison of GPU SDH with GPU SDH improved}
  \label{fg:gpusdhsdh}
@@
 
-\subsection{Overall Performance}
-
-\hspace{3em} It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
+\noindent\hspace{3em} It is noticable that growth of the speedup is not stable and with certain number of atoms it drops dramatically.
 One of possible reasons could the fact that we've used static approach of computing grid dimensions.

### 2016-10-19 thesis_paper/source.tex
Adjusting thesis

@@
 \section{One Body Functions}
-\hspace{3em} One body functions such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough, and taking into consideration the fact that we are going with push-based system, it is done at the same time for all of them.
+\hspace{3em} One body functions such as sum of masses, center of mass, or simple counting with selection  are of a complexity $O(n)$, thus, in order to get the results for them going through every atom (essentially over the entire data) only once is enough..
 
@@
 SDH is a very expensive computation, and we have used some of the existing work of ours\cite{sdhgpu} to incorporate in this design.
-
+\clearpage
 \noindent\hspace{3em}For computation of histogram it is not enough to visit every value for each atom once, since we have to do pairwise computations, thus it increases complexity of the problem not only computation-wise, but memory-wise as well, which is a bigger concern in our case, since the overhead will be huge and non-linearly relative to the size of data.
@@
 In order to understand how queries module is organized and the reason for it, we need to understand how CUDA and GPU work.
-
+\clearpage
 \noindent\hspace{3em} From hardware standpoint, it is important to understand that every GPU device consists of SMs (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\clearpage
 \noindent\hspace{3em} Once the data is loaded into global memory of GPU, we have two kernels for two types of queries setup for two different streams.
@@
 \noindent\hspace{3em} Implementation for CPU version of SDH is very simple, we do 2 nested loops that calculate histogram.
-Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them, though results don't change a lot for sequential computation.
+Since SDH criteria and parameters might vary by bucket width and box size for the simulation space, we've used one sample size for most of them.
 
@@
 Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.
-
+\clearpage
 \noindent\hspace{3em}Here are the main steps of the algorithm:
@@
 
+\clearpage
 \noindent\hspace{3em}The above Pseudocode \ref{alg:cpuflowfuture} shows what memory concurrent could look like.
@@
 
-\hspace{3em} You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
+\noindent\hspace{3em} You can instantly notice that speedup for 1000 atoms is not very impressive, especially for $10$ frames, and the reason for that is nature of GPU and memory transfer.
 There is a small overhead we have to consider before computing part: in order to let GPU process the data, we have to transfer the data to Global Memory, and upon completion it is needed to load the results back to RAM to show it to the user.
@@
 
-\hspace{3em} When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good.
+\noindent\hspace{3em}When we finally get to the number of atoms that are realistic to the ones scientists use, happily, speedup gets pretty good.
 Speedup gets up to almost 40 times (Figure \ref{fg:speedup_100000}).
